### 2015-09-24 paper.Rmd
foreign direct investment article

---
title: Characterization of foreign direct investment in Mozambique and implications
  for public health
author: "Joe Brew and Elisa Sicuri"
csl: health-economics-policy-and-law.csl
output:
  pdf_document:
    includes:
      in_header: style.sty
  html_document: default
highlight: zenburn
latex_engine: xelatex
header-includes: \usepackage{longtable}
bibliography: ../fdi_moz.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE }
# No scientific notation
options(scipen=999)

#####
# PACKAGES
#####
library(dplyr)
library(xtable)
library(ggplot2)
library(knitr)
library(png)
library(grid)
library(extrafont)
library(raster)
library(rgdal)
loadfonts(device = 'postscript', quiet = TRUE)

# BASIC KNITR OPTIONS
opts_chunk$set(comment = NA, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
options(xtable.comment = FALSE)
```

```{r}
#####
# GET MAPS OF MOZAMBIQUE
#####
setwd('data')
# This will get three different administrative
# levels of Mozambican shapefiles from the GADM database
# using the raster library
for (i in 0:3){
  assign(paste0('moz', i),
         getData('GADM', country = 'MOZ', level = i))
}
setwd('..')
```

```{r, fig.height = 3, fig.align='center'}
plot(moz0)
```

## Introduction 
Mozabique's recent economic growth has been facilitated by the government's open policy towards foreign investment, a plentiful supply of natural resources (particularly minerals and hydrocarbons) [@Rogers] and relatively inexpensive labor. Improvements in health have accompanied economic expansion, but Mozambique still lags behind in basic health outcomes - particularly those related to malaria - even by regional standards[@whostat]. Though strong evidence shows that improved health yields substantial economic benefits [@World1999, @Bloom2008], private foreign investment has played only a secondary role in malaria prevention and control in recent years.

Most of the malaria-related private investment in recent years has come in large, one-off projects which are well-publicized but difficult-to-reproduce [@rollback] or whose primary aim is social impact (not profit) [@Han]. However, the magnitude and kinds of recent foreign direct investment suggest the potential for a shift from a large-but-few to a small-but-numerous approach.  This shift, if carried out effectively, could be a "win-win" for business and public health.

Given the rapidly changing economic and epidemiologic context in Mozambique, a comprehensive and current understanding of both (a) the landscape of foreign direct investment in Mozambique and (b) public health issues which directly affect those investments is sorely needed. Such an understanding may foster greater private interest in for-profit investment in public health measures. Likewise, it can facilitate better public sector understanding of potential industry partners and stakeholders, a prerequisite to greater public-private collaborations.

This paper gives an overview of trends in foreign investment in Mozambique and their related public health issues. It is by no means comprehensive, but offers a consolidated starting point for understanding where the interests and incentives of the public and private sectors may converge.

## Macroeconomic overview

## Breakdown by industry

## Current private investments in health

## Potential for scaling-up collaboration

Bla bla bla[@Unctad2012].  Bla bla bla.

---
nocite: | 
  @Unctad2012
...

### 2015-09-24 junk.R

@@ -1,0 +1,10 @@
+
+
+---
+  nocite: | 
+  @Unctad2012
+...
+
+
+\bibliographystyle{plain}
+\bibliography{/home/joebrew/Documents/fdi_moz.bib}

### 2015-09-24 paper.Rmd
double columns!

@@
 latex_engine: xelatex
-header-includes: \usepackage{longtable}
 bibliography: ../fdi_moz.bib
@@
 ## Introduction
+
+\columnsbegin
+
 Mozabique's recent economic growth has been facilitated by the government's open policy towards foreign investment, a plentiful supply of natural resources (particularly minerals and hydrocarbons) [@Rogers] and relatively inexpensive labor.
@@
 
+\columnsend
+
 ## Macroeconomic overview

### 2015-09-25 paper.Rmd
clean up and maps

@@
 library(rgdal)
+library(readxl)
+library(tidyr)
+library(readr)
+library(RColorBrewer)
 loadfonts(device = 'postscript', quiet = TRUE)
@@
 options(xtable.comment = FALSE)
+
+# HELPER FUNCTIONS
+source('helpers.R')
 ```
@@
 
-Most of the malaria-related private investment in recent years has come in large, one-off projects which are well-publicized but difficult-to-reproduce [@rollback] or whose primary aim is social impact (not profit) [@Han].
+Most of the malaria-related private investment in recent years has come in large, "mega-projects" whose malaria abatement activities are well-publicized but difficult-to-reproduce [@rollback] or whose primary aim is social impact (not profit) [@Han].
 However, the magnitude and kinds of recent foreign direct investment suggest the potential for a shift from a large-but-few to a small-but-numerous approach.
@@
 
-\columnsend
-
 ## Macroeconomic overview
 
+Following independence (1974), Mozambique saw two decades of low and unsteady foreign investment, largely due to the civil war (which did not end until 1992).
+Thereafter, foreign investment began a steady increase but leveled off by 2000.
+However, the discovery of new sources of oil and gas set off a new spur of investments beginning in 2007.
+From 2010 to 2013, foreign direct investment grew from 1.26 to 6.70 billion USD, a more than five-fold increase.
+
+```{r}
+#####
+# READ IN WORLD BANK FOREIGN INVESTMENT DATA
+#####
+# http://data.worldbank.org/indicator/BX.KLT.DINV.CD.WD
+wb <- read_excel('data/bx.klt.dinv.cd.wd_Indicator_en_excel_v2.xls',
+                 skip = 3)
+wb <- data.frame(wb)
+names(wb[5:ncol(wb)]) <- paste0('y', names(wb[5:ncol(wb)]))
+
+# Clean up to just get mozambique
+temp <- gather(wb, year, value, X1961:X2015)
+# Clean up year column
+temp$year <- as.numeric(gsub('X', '', temp$year))
+# Subset to only include Mozambique
+temp <- temp[temp$Country.Name == 'Mozambique',]
+# Remove years without data
+temp <- temp[temp$year >= 1970 & temp$year <= 2014,]
+# temp <- temp[!is.na(temp$value),]
+
+#####
+# PLOT FOREIGN DIRECT INVESTMENT NET INFLOWS
+#####
+ggplot(data = temp, aes(x = year, y = value)) +
+  geom_point(alpha = 0.5, size = 2) +
+  geom_line(alpha = 0.5, size = 2) +
+  theme_bw() +
+  xlab('Year') +
+  ylab('Current USD (logarithmic scale)') +
+  scale_y_log10() +
+  ggtitle('Foreign direct investment: net inflows')
+```
+
+In addition to the discovery of new resources, recent growth has also been fueled by political and economic reforms which have made it easier for foreigners to do business in Mozambique.
+Of particular note, inflation has remained relatively low, and the costs to import and export have decreased dramatically, as have the costs of starting a business and registering property.
+
+```{r, comment = NA, message = FALSE, results = 'hide'}
+######
+# READ IN DOING BUSINESS DATA
+#####
+db <- read_excel('data/doing_business.xls')
+
+# Clean up columns
+db <- data.frame(db)
+macro_names <- names(db)
+for (j in 1:length(macro_names)){
+  number <- j
+  while(grepl('Var', macro_names[number])){
+    number <- number -1
+  }
+  macro_names[j] <- macro_names[number]
+}
+micro_names <- as.character(db[1,])
+combined_names <- gsub(' ', '', paste0(macro_names, '..', ifelse(is.na(micro_names), '', micro_names)))
+# Remove extra columns and replace column names
+db <- db[-1,]
+names(db) <- combined_names
+
+# Clean up year
+db$year <- as.numeric(gsub('DB', '', db$Year.., fixed = TRUE))
+
+# Get my columns of interest
+db$cost_export_container <- db[,"Trading.Across.Borders..Costtoimport(deflatedUS$percontainer)"]
+db$start_business <- db[,"Starting.a.Business..Time(days)"]
+db$cost_import_container <- db[,"Trading.Across.Borders..Costtoimport(deflatedUS$percontainer)" ]
+db$register_property <- 
+  db[,"Registering.Property..Cost(%ofpropertyvalue)"]
+
+db <- db[,c('year', 'cost_export_container', 'cost_import_container', 'start_business', 'register_property')]
+
+# Clean up
+temp <- gather(db, key, value, cost_export_container:register_property)
+temp$value <- as.numeric(temp$value)
+
+# Clean up names/labels
+temp$key <- factor(temp$key,
+                   labels = c('Cost to export container (deflated USD)',
+                              'Cost to import container (deflated USD)',
+                              'Cost (% of GDP per capita) to start business',
+                              'Cost (% of GDP per capita) to register property'))
+
+# PLOT
+g1 <- 
+  ggplot(data = temp[temp$key %in% unique(sort(temp$key))[1:2],], 
+         aes(x = year, y = value)) +
+    geom_line() +
+    facet_grid(.~key) +
+    theme_bw()
+g2 <- 
+  ggplot(data = temp[temp$key %in% unique(sort(temp$key))[3:4],], 
+         aes(x = year, y = value)) +
+    geom_line() +
+    facet_grid(.~key) +
+    theme_bw()
+multiplot(g1, g2)  
+```
+
 ## Breakdown by industry
+
+```{r}
+#####
+# MANUALLY ENTER DATA FROM INSTiTUtE OF STATISTICS
+#####
+industry <- data.frame(var = c('extractivas', 
+                               'alimentares',
+                               'madeira',
+                               'coque',
+                               'naometalicos',
+                               'metalicos',
+                               'equipamentos'),
+                       description = c('Extractive',
+                                       'Food, drink, tobacco',
+                                       'Wood, paper, printing',
+                                       'Chemical, petroluem products',
+                                       'Non-metalic minerals',
+                                       'Metalic minerals',
+                                       'Equipment (including vehicles)'))
+industry[,paste0('y', 2011:2014)] <- NA
+
+industry[industry$var == 'extractivas' ,paste0('y', 2011:2014)] <-
+  c(100, 194.4, 203.3, 230.1)
+industry[industry$var == 'alimentares' ,paste0('y', 2011:2014)] <-
+  c(100, 100.3, 147.2, 214.1)
+industry[industry$var == 'madeira' ,paste0('y', 2011:2014)] <-
+  c(100, 93.5, 103.3, 118.5)
+industry[industry$var == 'coque' ,paste0('y', 2011:2014)] <-
+  c(100, 105.6, 100.2, 79.5)
+industry[industry$var == 'naometalicos' ,paste0('y', 2011:2014)] <-
+  c(100, 102.9, 110, 131.5)
+industry[industry$var == 'metalicos' ,paste0('y', 2011:2014)] <-
+  c(100, 78.9, 83.2, 120.7)
+industry[industry$var == 'equipamentos' ,paste0('y', 2011:2014)] <-
+  c(100, 73.8, 167.5, 25.7)
+
+# Clean up a bit
+temp <- gather(industry, year, value, y2011:y2014)
+temp$year <- as.numeric(gsub('y', '', temp$year))
+temp$Industry <- temp$description
+
+ggplot(data = temp, aes(x = year, y = value, color = Industry)) +
+  geom_line(size = 2, alpha = 0.5) +
+  geom_point(aes(group = Industry)) +
+  theme_bw() +
+  geom_hline(yintercept = 100) +
+  scale_color_brewer(palette = 'Set1') +
+  xlab('Year') +
+  ylab('Percent of 2011 Volume') +
+  ggtitle('Industry volume index') +
+  theme(legend.position="bottom") +
+  guides(col = guide_legend(nrow = 3))
+
+```
+
+## Breakdown by region
+
+```{r}
+#####
+# READ IN GDP BY REGION/TIME DATA FROM
+# http://knoema.com/atlas/Mozambique/ranks/GDP-at-Constant-Prices
+#####
+gdp <- read_csv('data/gdp_by_region.csv')
+
+# Clean names to match map
+gdp$NAME[which(gdp$NAME == 'Niassa')] <- 'Nassa'
+
+# Remove city
+gdp <- gdp[gdp$NAME != 'Maputo cidade',]
+
+# Join map and data
+temp <- moz1
+temp@data <- left_join(x = temp@data, y = gdp, by = c('NAME_1' = 'NAME'))
+
+#####
+# PLOT
+#####
+cols <- colorRampPalette(brewer.pal(9, 'Blues'))(ceiling(max(temp$y2009)))
+colorfy <- function(number){
+  cols[round(number)]
+}
+
+par(mfrow = c(1,3))
+par(mar = c(4,0,4,0))
+plot(temp, col = colorfy(temp$y2000), border = NA); title(main = '2000')
+plot(temp, col = colorfy(temp$y2005), border = NA); title(main = '2005')
+plot(temp, col = colorfy(temp$y2009), border = NA); title(main = '2009')
+par(mar = c(5,4,2,2))
+
+legend_vec <- seq(6000, 24000, 6000)
+legend('right',
+       fill = cols[legend_vec],
+       bty = 'n',
+       border = NA,
+       legend = legend_vec)
+```
 
@@
 ...
+
+\columnsend
+
+******
+
+## References

### 2015-09-25 helpers.R

@@ -1,0 +1,122 @@
+
+##### SHARED LEGEND FOR MULTIPLE PLOTS GGPLOT2
+# https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs
+grid_arrange_shared_legend <- function(...) {
+  plots <- list(...)
+  g <- ggplotGrob(plots[[1]] + theme(legend.position="bottom"))$grobs
+  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
+  lheight <- sum(legend$height)
+  grid.arrange(
+    do.call(arrangeGrob, lapply(plots, function(x)
+      x + theme(legend.position="none"))),
+    legend,
+    ncol = 1,
+    heights = unit.c(unit(1, "npc") - lheight, lheight))
+}
+
+#####
+# EXTRACT JUST THE LEGEND
+# http://stackoverflow.com/questions/12539348/ggplot-separate-legend-and-plot
+g_legend<-function(a.gplot){
+  tmp <- ggplot_gtable(ggplot_build(a.gplot))
+  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
+  legend <- tmp$grobs[[leg]]
+  legend
+}
+
+##### CONFIDENCE INTERVALS ON PROPORTIONS
+# (https://aghaynes.wordpress.com/2014/04/09/calculating-confidence-intervals-for-proportions/)
+simpasym <- function(n, p, z=1.96, cc=TRUE){
+  out <- list()
+  if(cc){
+    out$lb <- p - z*sqrt((p*(1-p))/n) - 0.5/n
+    out$ub <- p + z*sqrt((p*(1-p))/n) + 0.5/n
+  } else {
+    out$lb <- p - z*sqrt((p*(1-p))/n)
+    out$ub <- p + z*sqrt((p*(1-p))/n)
+  }
+  out
+}
+
+
+capwords <- function(s, strict = FALSE) {
+  cap <- function(s) paste(toupper(substring(s, 1, 1)),
+                           {s <- substring(s, 2); if(strict) tolower(s) else s},
+                           sep = "", collapse = " " )
+  sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
+}
+
+library(reshape2)
+
+multiplot <- function(..., plotlist = NULL, file, cols = 1, layout = NULL) {
+  require(grid)
+  
[…]

### 2015-10-16 paper.Rmd
little things

@@
 
+Most of recent growth has come in the "extractive" industries, a term encompassing...  XXX
+
 ```{r}
@@
 
+The principal foreign firms operating in this space are YYY YYY YYY.
+
 ## Breakdown by region
+
+Growth has been similarly large in all districts.
+From 2000 to 2009, GDP approximately doubled, with the greatest growth occurring in Inhambane (119% growth from 2000 to 2009), and the least robust grwoth in Nampula (95%).
 
@@
 par(mfrow = c(1,3))
-par(mar = c(4,0,4,0))
+par(mar = c(4,0,2,1))
 plot(temp, col = colorfy(temp$y2000), border = NA); title(main = '2000')
@@
 
-legend_vec <- seq(6000, 24000, 6000)
+legend_vec <- seq(8000, 24000, 8000)
 legend('right',
@@
        border = NA,
-       legend = legend_vec)
+       legend = paste0(legend_vec/1000,'K'),
+       ncol = 1,
+       title = 'GDP')
 ```

### 2015-10-16 knit.R

@@ -1,0 +1,6 @@
+setwd('/home/joebrew/Documents/fdi_moz')
+require(knitr) # required for knitting from rmd to md
+require(markdown) # required for md to html 
+knit('paper.Rmd', 'paper.md') # creates md file
+markdownToHTML('paper.md', 'paper.html') # creates html file
+browseURL(paste('file://', file.path(getwd(),'paper.html'), sep='')) # open file in browser

### 2016-01-04 paper.Rmd
minor updates

@@
 #####
-cols <- colorRampPalette(brewer.pal(9, 'Blues'))(ceiling(max(temp$y2009)))
+cols <- colorRampPalette(brewer.pal(9, 'Blues'))(ceiling(max(temp$y2009, na.rm = TRUE)))
 colorfy <- function(number){

### 2016-01-27 paper.Rmd
done with first rev

--- 2015-09-25 paper.Rmd
@@
 ---
-title: Characterization of foreign direct investment in Mozambique and implications
-  for public health
+title: Foreign direct investment, corporate social responsibility, and malaria control in Mozambique
 author: "Joe Brew and Elisa Sicuri"
@@
 latex_engine: xelatex
-bibliography: ../fdi_moz.bib
+bibliography: library.bib
 ---
@@
 
-```{r, fig.height = 3, fig.align='center'}
+```{r, fig.height = 3, fig.align='center', eval = FALSE}
 plot(moz0)
 ```
+
+\begin{changemargin}{3cm}{3cm}
+
+\begin{center}
+\textbf{Abstract}
+\end{center}
+
+\emph{Foreign direct investment in Mozambique has increased rapidly in the last two decades.
+The growing interest in corporate social responsibility - combined with a recent push for malaria eradication - suggest the need for a critical examination of how firms spend money on malaria control.
+The interests of both the private and public sectors align in the need for more private investment in malaria control.
+However, given the lack of coordination between public and private sectors, an over-reliance on foreign and private initiative for funding eradication is not without risks.}
+
+\end{changemargin}
 
@@
 Mozabique's recent economic growth has been facilitated by the government's open policy towards foreign investment, a plentiful supply of natural resources (particularly minerals and hydrocarbons) [@Rogers] and relatively inexpensive labor.
-Improvements in health have accompanied economic expansion, but Mozambique still lags behind in basic health outcomes - particularly those related to malaria - even by regional standards[@whostat].
-Though strong evidence shows that improved health yields substantial economic benefits [@World1999, @Bloom2008], private foreign investment has played only a secondary role in malaria prevention and control in recent years.
+Improvements in health have accompanied economic expansion, but Mozambique still lags behind in basic health outcomes - particularly those related to malaria - even by regional standards [@whostat].
+Despite  strong evidence showing that improved health yields substantial economic benefits [@World1999, @Bloom2008], direct foreign investment (FDI) under the auspices of "corporate social responsibility" (CSR) has largely ignored malaria prevention in favor of other social good activities.
 
-Most of the malaria-related private investment in recent years has come in large, "mega-projects" whose malaria abatement activities are well-publicized but difficult-to-reproduce [@rollback] or whose primary aim is social impact (not profit) [@Han].
-However, the magnitude and kinds of recent foreign direct investment suggest the potential for a shift from a large-but-few to a small-but-numerous approach.
-This shift, if carried out effectively, could be a "win-win" for business and public health.
+The few malaria-related CSR projects in recent years has come in large, "mega-projects" whose well-publicized malaria abatement activities are profit-driven [@Mouzin2011] or whose primary aim is not profit-related [@Han].
+Though the former is often portrayed as a "win-win" for business and public health, the latter also offers tangible benefits for private industry, and should be understood as operating under the same conditions and with the same motivations.
 
-Given the rapidly changing economic and epidemiologic context in Mozambique, a comprehensive and current understanding of both (a) the landscape of foreign direct investment in Mozambique and (b) public health issues which directly affect those investments is sorely needed.
+In 2015, the flood of FDI into Mozambique (and many other subsaharan African nations) slowed to a trickle.
+The impact of this slowdown on CSR is not yet known, but it can reasonably be assumed that it will mean a reduction in CSR activities (albeit with lag).
+Given the rapidly changing economic and epidemiologic context in Mozambique, a comprehensive and current understanding of both (a) the landscape of FDI and CSR in Mozambique and (b) public health issues (namely, malaria control) which are directly affected those investments is sorely needed.
 Such an understanding may foster greater private interest in for-profit investment in public health measures.
-Likewise, it can facilitate better public sector understanding of potential industry partners and stakeholders, a prerequisite to greater public-private collaborations.
+Likewise, it can facilitate better public sector understanding of potential industry partners and stakeholders (a prerequisite to greater public-private collaborations), and guide the public sector away from an inflexible dependence on FDI for the provision of public health necessities.
 
-This paper gives an overview of trends in foreign investment in Mozambique and their related public health issues.
-It is by no means comprehensive, but offers a consolidated starting point for understanding where the interests and incentives of the public and private sectors may converge.
+This paper gives an overview of trends in FDI and CSR in Mozambique, with a focus on its impact on malaria control.
+It is by no means comprehensive, but offers a consolidated starting point for understanding where the interests and incentives of the public and private sectors converge and differ in regards to malaria control.
 
-## Macroeconomic overview
+## FDI and CSR in Mozambique
+
+### Massive increase in FDI
 
 Following independence (1974), Mozambique saw two decades of low and unsteady foreign investment, largely due to the civil war (which did not end until 1992).
-Thereafter, foreign investment began a steady increase but leveled off by 2000.
-However, the discovery of new sources of oil and gas set off a new spur of investments beginning in 2007.
-From 2010 to 2013, foreign direct investment grew from 1.26 to 6.70 billion USD, a more than five-fold increase.
+Thereafter, foreign investment began a steady increase but leveled off by the late 1990s.
+However, the discovery of novel sources of oil and gas set off a new spur of investments beginning in 2007, and continuing through last year.
+From 2010 to 2013, foreign direct investment grew from 1.26 to 6.70 billion USD, a more than five-fold increase [@wbdata].
 
@@
 temp <- temp[temp$year >= 1970 & temp$year <= 2014,]
-# temp <- temp[!is.na(temp$value),]
+temp <- temp[!is.na(temp$value),]
+temp <- temp[temp$value > 0,]
 
@@
 ggplot(data = temp, aes(x = year, y = value)) +
-  geom_point(alpha = 0.5, size = 2) +
-  geom_line(alpha = 0.5, size = 2) +
+    scale_y_log10() +
+  geom_point(alpha = 0.5, size = 1) +
+  geom_line(alpha = 0.5, size = 1) +
   theme_bw() +
@@
   ylab('Current USD (logarithmic scale)') +
-  scale_y_log10() +
   ggtitle('Foreign direct investment: net inflows')
@@
 In addition to the discovery of new resources, recent growth has also been fueled by political and economic reforms which have made it easier for foreigners to do business in Mozambique.
-Of particular note, inflation has remained relatively low, and the costs to import and export have decreased dramatically, as have the costs of starting a business and registering property.
+Of particular note, inflation remained relatively low (at least through 2014) [@deutsche].
+
+```{r}
+# Load up forex data
+if('forex.csv' %in% dir('data')){
+  forex <- read_csv('data/forex.csv')
+} else {
+  library(Quandl)
+  forex <- Quandl("BUNDESBANK/BBEX3_M_MZN_USD_CA_AC_A01")
+  write_csv(forex, 'data/forex.csv')
+}
+
+# Include only through Jan 2015
+forex <- forex[forex$Date <= '2015-01-01',]
+
+ggplot(data = forex,
+       aes(x = Date, y = Value)) +
+  geom_line(alpha = 0.5, size = 1) +
+  theme_bw() +
+  ggtitle('Meticals per US dollar') +
+  ylab('Meticals') +
+  geom_point(alpha = 0.5, size = 1) +
+  ylim(0, 40)
+
+```
+
+The massive increases in FDI have also been facilitated by dramatic decreases in the costs to import and export, as well as the costs of starting a business and registering property [@doingbusiness].
 
@@
 
-## Breakdown by industry
+### Breakdown by industry
+
+Most of recent growth has come in the "extractive" industries, a term encompassing a range of industry, but in the Mozambican context largely applying to hydrocarbons and mining [@estatistica].
 
@@
 ggplot(data = temp, aes(x = year, y = value, color = Industry)) +
-  geom_line(size = 2, alpha = 0.5) +
+  geom_line(size = 1, alpha = 0.5) +
   geom_point(aes(group = Industry)) +
@@
 
-## Breakdown by region
+The late 2015 economic slow-down in the developing world, particularly the low price of oil, could have serious repercussions for FDI in Mozambique.
+That said, the mining of metals and the service industries both make up a larger share of Mozambique's economic output than the extraction of hydrocarbons, which should somewhat buffer the Mozambican economy from the negative effects of low oil prices.
+
+### Breakdown by region
+
+Despite the concentration of private investment in regional projects, growth has been similarly large in all districts.
+From 2000 to 2009, GDP approximately doubled, with the greatest growth occurring in Inhambane (119% growth from 2000 to 2009), and the least robust grwoth in Nampula (95%) [@knoema].
 
@@
 
-#####
-# PLOT
-#####
-cols <- colorRampPalette(brewer.pal(9, 'Blues'))(ceiling(max(temp$y2009)))
-colorfy <- function(number){
-  cols[round(number)]
-}
+##### PLOT TS
+ts <- temp@data
+ts <- gather(ts, year, gdp, y2009:y2000)
+ts$year <- as.numeric(gsub('y', '', ts$year))
+ts$year <- as.Date(paste0(ts$year, '-01-01'))
 
-par(mfrow = c(1,3))
-par(mar = c(4,0,4,0))
-plot(temp, col = colorfy(temp$y2000), border = NA); title(main = '2000')
-plot(temp, col = colorfy(temp$y2005), border = NA); title(main = '2005')
-plot(temp, col = colorfy(temp$y2009), border = NA); title(main = '2009')
-par(mar = c(5,4,2,2))
+# Remove Maputo city
+ts <- ts[ts$NAME_1 != 'Maputo City',]
 
-legend_vec <- seq(6000, 24000, 6000)
-legend('right',
-       fill = cols[legend_vec],
-       bty = 'n',
-       border = NA,
-       legend = legend_vec)
+# Arrange
+ts <- ts %>% arrange(year)
+
+# Rebase to 100 %
+ts <- ts %>%
+  left_join(ts %>% 
+              group_by(NAME_1) %>%
+              summarise(start = first(gdp))) %>%
+  mutate(gdp = gdp / start * 100)
+
+ggplot(data = ts, 
+       aes(x = year, y = gdp, col = NAME_1)) +
+  geom_line(alpha = 0.6) +
+  geom_point() +
+  xlab('Year') +
+  ylab('GDP (% of 2000 value)') +
+  theme_bw() +
+  scale_colour_manual(guide = guide_legend(reverse=TRUE), 
+                      name = 'District',
+                      values = brewer.pal(length(unique(ts$NAME_1)), 'Spectral')) +
+  ggtitle('District-specific growth in GDP')
 ```
 
-## Current private investments in health
+The homogeneity in growth comes somewhat as a surprise, as it defies the general developing world pattern in that growth in areas that already had high GDP (Maputo and the coastal provinces) was as robust as growth in areas with previously low GDP.
 
-## Potential for scaling-up collaboration
+```{r}
+#####
+# PLOT MAPS
+#####
 
-Bla bla bla[@Unctad2012].  Bla bla bla.
+# Get columns for coordinates
+temp$x <- coordinates(temp)[,1]
+temp$y <- coordinates(temp)[,2]
 
----
-nocite: | 
-  @Unctad2012
-...
+# Fortify
+temp_fort <- fortify(temp, id = 'ID_0')
+
+# Bring in gdp data to fortified map
+temp_fort <- left_join(x = temp_fort,
+                       y = temp@data %>%
+                         mutate(id = as.character(ID_1)))
+
+# Gather by year
+temp_fort <- gather(temp_fort, year, gdp, y2009:y2000)
+temp_fort$year <- as.numeric(gsub('y', '', temp_fort$year))
+
+ggplot(data = temp_fort,
+       aes(x = long, 
+           y = lat, 
+           group = group,
+           fill = gdp)) +
+  geom_polygon() +
+  coord_map() +
+  facet_wrap(~year) +
+  xlab('Longitude') +
+  ylab('Latitude') +
+  theme_bw() +
+  scale_fill_gradient(low = 'yellow',
+                      high = 'darkblue',
+                      # guide = guide_legend(reverse=TRUE), 
+                      name = 'GDP')
+
+# cols <- colorRampPalette(brewer.pal(9, 'Blues'))(ceiling(max(temp$y2009, na.rm = TRUE)))
+# colorfy <- function(number){
+#   cols[round(number)]
+# }
+# 
+# par(mfrow = c(1,3))
+# par(mar = c(4,0,2,1))
+# plot(temp, col = colorfy(temp$y2000), border = NA); title(main = '2000')
+# plot(temp, col = colorfy(temp$y2005), border = NA); title(main = '2005')
+# plot(temp, col = colorfy(temp$y2009), border = NA); title(main = '2009')
+# par(mar = c(5,4,2,2))
+# 
+# legend_vec <- seq(8000, 24000, 8000)
+# legend('right',
+#        fill = cols[legend_vec],
+#        bty = 'n',
+#        border = NA,
+#        legend = paste0(legend_vec/1000,'K'),
+#        ncol = 1,
+#        title = 'GDP')
+```
+
+### CSR in Mozambique
+
+Corporate social responsibility in Mozambique is new in nomenclature, but activities which could be classified as CSR have existed for decades.
+The number of firms actively engaging in CSR cannot be ascertained (given the large and ever-changing number of small businesses), but virtually all of the largest firms, both foreign and domestic, have a CSR component.
+
+Firms with CSR activities are often large and foreign.
+Among the largest "key players" in Mozambican CSR are Coca-Cola, British Petroleum and Colgate-Palmolive.
+State conglomerates, such as Águas de Moçambique and Electricidade de Moçambique, also participate in CSR activities [@Compact2007].
+
+### Malaria-related CSR (or lack thereof)
+
+Most CSR-funded activities are focused on education, community development, women's rights and entrepreneurship.
+According to a UN poll, none of the country's largest firms invest directly in malaria-related CSR activities [@Compact2007].
+This may be due to the perceived costs of malaria control, lack of perceived PR benefits, as well as the "opt-in" nature and privacy/legal issues generally related to engaging in health-related campaigns.
+
+That said, a majority of the firms interviewed by the UN indicated that one of the principal reasons for investment in CSR is to complement government efforts.
+To the extent that malaria accounts for more of the loss in disability-adjusted life years in Mozambique than comparable countries [@ihme], malaria control's lack of representation among core CSR activities is a notable absence.
+
+# Scaling up malaria control through CSR: opportunity and risk
+
+### Opportunity
+
+The historical absence of malaria-related CSR represents an opportunity for complementarity.
+In addition to improvements in public health, both the private and public sectors stand to benefit economically from a scaling up of malaria control driven by the private sector.
+By increasing malaria control activities, as a share of total CSR expenditures, public funds could be redirected towards other areas of health.
+Likewise, if private CSR activities pivoted towards malaria control rather more general philanthropical gestures, CSR would have a more direct impact on wellbeing (with less temporal lag), thereby fulfilling the public relations goals of the firms that invest in CSR.
+Finally, for many industries, the firm itself is a potential direct beneficiary, given that improvements in employee health and a reduction in employee absenteeism can be directly correlated to productivity.
+
+### Risk
+
+Scaling up CSR while also encouraging its redirection towards malaria control is not without risks.
+The most notable downfall of this approach is the potential for the inadvertent dependence on the private sector for what is essentially a public good.
+Were CSR targeting malaria control to reach significant levels (and the government were to enact a correspoding redirection of funds towards the financing of other health areas), then a situation would be created in which the public sector had essentially divested from a public good.
+This would be unwise and dangerous.
+
+A secondary risk is that increased private sector involvement in malaria control could cause a decrease in public sector _competence_ in the prevention and treatment of malaria.
+This could have negative consequences in the case of either a financial or economic crisis (in which CSR activity would be curtailed) or an increase in malarial activity.
+By the same token, CSR involvement in malaria control could potentially portend, at least initially, less effective interventions.
+Private firms' incentives, though aligned with the public's in terms of malaria, are not identical, and pressures from shareholders and for positive public perception might motivate malaria control strategies which do not necessarily carry with them the most recent scientific knowledge.
+
+A third risk is a lack of coordination.
+Both in terms of logistical activities as well as biological realities (drug and insecticide resistance, etc.), coordination of malaria control activities is absolutely essentially if Mozambique is going to make the transition to eradication.
+The issue of coordination could be solved through an activities and outcomes reporting/surveillance structure (necessarily managed by the state), but compliance could be problematic.
+
+A final risk is that of volatility.
+By centralizing malaria control under the auspices of the government, public health authorities can effectively distribute malaria control expenditures to where and when they are most needed.
+If this control were only in the hands of private firms, expenditure would likely transform into a function of firm-specific profitability and shareholder incentives, as well as market cycles.
+This could lead to a situation in which malaria control activities are most prevalent in areas where the economy is strongest, rather than areas where the need is greatest.
+
+# Conclusion
+
+High FDI in Mozambique and growing interest in CSR both call for increased reflection on the private sector's role in the delivery of public goods.
+A refocusing of CSR expenditures into areas where the need is greatest (specifically, malaria control) could lead to better health and greater profits.
+This "win-win" for the public and private sectors represents a rare opportunity, which deserves more discourse, research and experimentation.
+
+That said, the growing interest in CSR suggests both that (a) the need for services exist and (b) that corporations (especially foreign firms) have enough excess capital to finance these services.
+Both of these factors suggest the need for a better-fitting taxation rate, and a more efficient delivery of public services.
+That said, in the short-term, increasing the efficiency and effectiveness of CSR activity through a re-pivoting towards malaria control should remain a goal.
 

### 2016-01-27 google_trends.R

@@ -1,0 +1,33 @@
+library(gtrendsR)
+
+##### Set up Google Trends
+usr <- "<Google account email>"
+psw <- "<Google account password>"
+ch <- gconnect(usr, psw) 
+# x <- gtrends(c("data is", "data are"), res="day")
+x <- gtrends(query = 'corporate social responsibility',  
+                res="week")
+
+##### Get the trend into a dataframeag
+trend <- x$trend
+
+##### Make a date object from start
+trend$date <- as.Date(trend$start)
+
+##### Get a year column
+trend$year <- format(trend$date, '%Y')
+
+##### Rename interest
+names(trend)[3] <- 'interest'
+
+##### Group by year and replot
+temp <- 
+  trend %>%
+  group_by(year) %>%
+  summarise(avg = mean(interest))
+
+## Make year numeric
+temp$year <- as.numeric(temp$year)
+
+# Plot
+plot(temp$year, temp$avg)

### 2016-02-25 abstract.Rmd
fdi abstract

---
title: Abstract - Foreign direct investment, corporate social responsibility, and malaria control in Mozambique
author: 
  - Joe Brew*
  - Elisa Sicuri*  
csl: health-economics-policy-and-law.csl
output:
  pdf_document:
    includes:
      in_header: style.sty
  html_document: default
highlight: zenburn
latex_engine: xelatex
bibliography: library.bib
---

```{r, echo=FALSE, comment=NA, message=FALSE, warning=FALSE} 
#libraries
library(ggmap)
library(ggplot2)
library(grid)
library(dplyr)
library(knitr)
library(ggthemes)
library (rgdal)
library(RColorBrewer)
library(maptools)
opts_chunk$set(comment = NA, message = FALSE, echo = FALSE, results = 'hide', fig.height = 6, warning = FALSE, fig.width = 6)
```

```{r}
suppressWarnings(library(xtable))
options(xtable.comment = FALSE)
```

```{r}
source('helpers.R')
```

\begin{changemargin}{1cm}{1cm}

\begin{center}
\Large{\textbf{Abstract}}
\end{center}
\begin{large}
\textbf{Context:} Foreign direct investment in Mozambique has increased rapidly in the last two decades. The growing interest in corporate social responsibility - combined with a recent push for malaria eradication - suggest the need for a critical examination of how firms spend money on malaria control. The interests of both the private and public sectors align in the need for more private investment in malaria control.  However, given the lack of coordination between public and private sectors, an over-reliance on foreign and private initiative for funding eradication is not without risks. \\

\textbf{Content:} We provide an overview and reflect on the experiences of private firms engaging in privately-funded malaria control activities in Mozambique. Specifically, we explore the benefits of private financing of malaria control interventions, and caution against an overreliance on such an approach from a public health perspective. Additionally, we map out the landscape of private industries, and attempt to identify areas that are ripe for a "scaling-up" of private investment. \\

\textbf{Form:} We will present an overview paper, as well as a combination of case studies. Though the latter are somewhat anecdotal, the former will rely primarily on publicly available economic and industry data, complemented by interviews and questionnaires from key industry members.\\

\end{large}

\end{changemargin}

* All authors affiliated with the Institue of Global Health of Barcelona (www.isglobal.org).

### 2016-04-14 paper.Rmd
adding dhs irs data

@@
 #####
+library(maptools)
+library(maps)
 library(dplyr)
@@
 library(RColorBrewer)
+library(ggmap)
+library(rasterVis)
+library(colorspace)
+
 loadfonts(device = 'postscript', quiet = TRUE)
@@
   ylim(0, 40)
-
 ```
@@
 gdp <- read_csv('data/gdp_by_region.csv')
+names(gdp)[1] <- 'NAME'
 
@@
                       name = 'GDP')
-
 # cols <- colorRampPalette(brewer.pal(9, 'Blues'))(ceiling(max(temp$y2009, na.rm = TRUE)))
@@
 
-### Malaria-related CSR (or lack thereof)
+### Malaria-related CSR and other corporate activity
 
@@
 According to a UN poll, none of the country's largest firms invest directly in malaria-related CSR activities [@Compact2007].
-This may be due to the perceived costs of malaria control, lack of perceived PR benefits, as well as the "opt-in" nature and privacy/legal issues generally related to engaging in health-related campaigns.
+This may be due to the perceived costs of malaria control, lack of perceived PR benefits, the government and NGO's predominant role in the area, as well as the "opt-in" nature and privacy/legal issues generally related to engaging in health-related campaigns.
 
@@
 To the extent that malaria accounts for more of the loss in disability-adjusted life years in Mozambique than comparable countries [@ihme], malaria control's lack of representation among core CSR activities is a notable absence.
+
+Whether under the guise of CSR or not, it is noteworthy that the private sector currently does play a role in malaria control activities.
+According to 2011 DHS data, greater than 7 in every 1,000 households had a private company carry out indoor residual spraying [@dhs].
+And, in some clusters, the percent of houses covered by private IRS was greater than 25%.
+
+```{r}
+# READ IN DHS DATA
+# (generated in dhs/exploration.Rmd -  https://github.com/joebrew/dhs/)
+dhs <- read_csv('data/dhs_private_irs.csv')
+
+# Get mean by cluster
+by_cluster <-
+  dhs %>%
+  group_by(cluster) %>%
+  summarise(lat = first(lat),
+            lon = first(lon),
+            urban_rural = first(urban_rural),
+            n = sum(n),
+            private = mean(private, na.rm = TRUE),
+            weight = first(weight)) %>%
+  filter(lat < -10, lon >20)
+
+# Get weighted mean
+x = weighted.mean(by_cluster$private, w = by_cluster$weight)
+
+# Make a spatial version
+houses <- 
+  by_cluster %>%
+  mutate(latitude = lat,
+         longitude = lon)
+coordinates(houses) <- ~longitude + latitude
+proj4string(houses) <- proj4string(moz0)
+
+# Get a smoothed surface for private IRS coverage
+if('irs_grid.RData' %in% dir()){
+  load('irs_grid.RData')
+} else {
+
+  # Create dataset of gridded values
+  moz <- map('world', 'moz', plot = FALSE)
+  lons <- seq(min(moz$x, na.rm = TRUE), 
+              max(moz$x, na.rm = TRUE), 
+              length = 100)
+  lats <- seq(min(moz$y, na.rm = TRUE), 
+              max(moz$y, na.rm = TRUE), 
+              length = 100)
+  df_grid <- expand.grid(lon = lons,
+                         lat = lats)
+
+  # Create latitude and longitude columns and convert to spatial
+  df_grid$latitude <- df_grid$lat
+  df_grid$longitude <- df_grid$lon
+  coordinates(df_grid) <- ~longitude + latitude
+  proj4string(df_grid) <- proj4string(moz0)
+
+  # Keep only those points within mozambique
+  x <- over(df_grid, polygons(moz0))
+  df_grid <- df_grid[!is.na(x),]
+
+  # Create a value in df_grid
+  df_grid$val <- NA
+
+  # Loop through the grid, getting a weighte dvalue of mosquito net or not
+  nn <- nrow(df_grid)
+  for (i in 1:nn){
+      message(paste0('row ', i,
+                     ' of ', nn))
+
+    # Get distance from specific point in df_grid 
+    # to every point in houses
+    distances <- spDistsN1(pts = houses,
+                          pt = df_grid[i,],
+                          longlat = TRUE)
+    # Define which are acceptably close
+    close_enough <- which(distances <= 50000)
+    # Get an val score (weighted likelihood of having net)
+    val <- stats::weighted.mean(x = houses$private[close_enough],
+                         w = ((1 / distances[close_enough]) ^2) *
+                           houses$weight[close_enough],
+                         na.rm = TRUE)
+    # Assign val to the dataframe
+    df_grid$val[i] <- val
+  }
+  save('df_grid', file = 'irs_grid.RData')
+}
+
+# Convert to raster
+temp <- df_grid@data %>% arrange(lon, lat)
+r <- rasterFromXYZ(temp[, c('lon', 'lat', 'val')])
+levelplot(r, contour = TRUE, par.settings=RdBuTheme)
+```
+
+```{r}
+# Fortify moz1
+moz2_fortified <- fortify(moz2, region = 'NAME_2')
+
+# Plot clusters
+ggplot(data = moz2_fortified,
+       aes(x = long, y = lat)) +
+  geom_polygon(aes(group = id)) +
+  coord_map() +
+  geom_point(data = by_cluster,
+             aes(x = lon, 
+                 y = lat,
+                 size = weight,
+                 color = private))
+```
 

### 2016-04-14 paper.Rmd
more mapping

@@
 library(colorspace)
+library(dismo)
 
@@
 
+Interestingly, though the UN data indicated that corporations don't actively engage in malaria control as part of CSR, the geographic distribution of households which had their homes sprayed by a private entity are clustered in areas where foreign firms operate, particularly in the south (around Maputo) in the East, where extractive industry activity is highest.
+
 ```{r}
@@
 r <- rasterFromXYZ(temp[, c('lon', 'lat', 'val')])
-levelplot(r, contour = TRUE, par.settings=RdBuTheme)
+proj4string(r) <- proj4string(moz1)
+# levelplot(r, contour = TRUE, par.settings=RdBuTheme)
+
+# # Convert raster to polygons
+# rtp <- rasterToPolygons(r)
+# rtp@data$id <- 1:nrow(rtp@data) 
 ```
@@
 # Plot clusters
+cols <- brewer.pal(9, 'Spectral')
 ggplot(data = moz2_fortified,
        aes(x = long, y = lat)) +
-  geom_polygon(aes(group = id)) +
-  coord_map() +
-  geom_point(data = by_cluster,
-             aes(x = lon, 
-                 y = lat,
-                 size = weight,
-                 color = private))
+  # coord_cartesian() + 
+  geom_raster(data = df_grid@data,
+              aes(x = lon, y = lat, fill = val * 100)) +
+    scale_fill_gradientn(name="Percent",colours = cols) +
+    coord_equal() + 
+    # geom_polygon(aes(group = id), 
+    #              color = 'black',
+    #              fill = NA,
+    #              size = 0.1,
+    #              alpha = 0.3) +
+  theme_bw() +
+  xlab('Longitude') +
+  ylab('Latitude') +
+  # theme_void() +
+  ggtitle('(Smoothed) percent of households with private IRS') +
+  theme(legend.position="bottom")
+
 ```

### 2016-11-09 paper.Rmd
more updates

@@
 Mozabique's recent economic growth has been facilitated by the government's open policy towards foreign investment, a plentiful supply of natural resources (particularly minerals and hydrocarbons) [@Rogers] and relatively inexpensive labor.
-Improvements in health have accompanied economic expansion, but Mozambique still lags behind in basic health outcomes - particularly those related to malaria - even by regional standards [@whostat].
+Improvements in health have accompanied economic expansion, but Mozambique still lags behind in basic health outcomes - particularly those related to malaria - even by regional standards.
 Despite  strong evidence showing that improved health yields substantial economic benefits [@World1999, @Bloom2008], direct foreign investment (FDI) under the auspices of "corporate social responsibility" (CSR) has largely ignored malaria prevention in favor of other social good activities.

### 2016-11-20 paper.Rmd
minor changes

@@
 title: Foreign direct investment, corporate social responsibility, and malaria control in Mozambique
-author: "Joe Brew and Elisa Sicuri"
+author: "Trends, risks, and opportunities"
 csl: health-economics-policy-and-law.csl
@@
 \begin{center}
+
+\begin{large}
+Joe Brew\footnote{Barcelona Institute for Global Health; Vrije Universiteit Amsterdam} and Elisa Sicuri\footnote{Barcelona Institute for Global Health; Imperial College London}
+\end{large}
+
 \textbf{Abstract}
@@
 
-\emph{Foreign direct investment in Mozambique has increased rapidly in the last two decades.
-The growing interest in corporate social responsibility - combined with a recent push for malaria eradication - suggest the need for a critical examination of how firms spend money on malaria control.
-The interests of both the private and public sectors align in the need for more private investment in malaria control.
-However, given the lack of coordination between public and private sectors, an over-reliance on foreign and private initiative for funding eradication is not without risks.}
+\emph{Foreign direct investment (FDI) in Mozambique has increased rapidly in the last two decades.
+The growing interest in corporate social responsibility (CSR) - combined with a recent push for malaria eradication - suggest the need for a critical examination of the research attention garnered by FDI and CSR, as well as how private firms spend money on  control.
+Through a systematic review of publications, we find that there has been a notable increase in research pertaining to FDI and CSR in recent years.
+However, given the lack of coordination between public and private sectors, even when interests coincide, an over-reliance on foreign and private initiative for funding eradication is not without risks.}
 
@@
 This paper gives an overview of trends in FDI and CSR in Mozambique, with a focus on its impact on malaria control.
-It is by no means comprehensive, but offers a consolidated starting point for understanding where the interests and incentives of the public and private sectors converge and differ in regards to malaria control.
+It also offers a quasi-systematic review of the research attention garnered by FDI and CSR in recent years.
+It is by no means comprehensive, but offers a consolidated starting point for understanding both the place of FDI and CSR in the literature, as well as where the interests and incentives of the public and private sectors converge and differ in regards to malaria control.
 
@@
 
+# Systematic review of FDI and CSR in health research
+
+```{r}
+
+# PERFORM PUBMED SEARCH
+source('search_pubmed.R')
+if('search_results.RData' %in% dir()){
+  load('search_results.RData')
+} else {
+  csr_results <- pubmed(start_year = 2000,
+                         end_year = 2016,
+                         search_topic = '(malaria) and (corporate social responsibility)',
+                         counts_only = FALSE,
+                         n_records = 5000)  
+  fdi_results <- pubmed(start_year = 2000,
+                       end_year = 2016,
+                       search_topic = '(malaria) and (foreign direct investment)',
+                       counts_only = FALSE,
+                       n_records = 5000)  
+    moz_csr_results <- pubmed(start_year = 2000,
+                     end_year = 2016,
+                     search_topic = '(mozambique) and (corporate social responsibility)',
+                     counts_only = FALSE,
+                     n_records = 5000)  
+  moz_fdi_results <- pubmed(start_year = 2000,
+                     end_year = 2016,
+                     search_topic = '(mozambique) and (foreign direct investment)',
+                     counts_only = FALSE,
+                     n_records = 5000)  
+  save(csr_results,
+       fdi_results,
+       moz_csr_results,
+       moz_fdi_results,
+       file = 'search_results.RData')
+}
+
+# Charts etc. will eventually go here
+```
+
+In order to gauge research attention and focus on FDI and CSR insofar as they affect Mozambique and malaria, four systematic searches were carried out using the NCBI/pubmed databases:
+
+1. `"(malaria) and (corporate social responsibility)"`
+2. `"(malaria) and (foreign direct investment)"`
+3. `"(mozambique) and (corporate social responsibility)"`
+4. `"(mozambique) and (foreign direct investment)"`
+
+The number of records retrieved for the 4 searches can be visualized over time below:
+
+```{r}
+barplot(1:10)
+```
+
+There are two important interpretations to the above plot. XXX YYY
+
 # Scaling up malaria control through CSR: opportunity and risk

### 2016-11-20 search_pubmed.R

@@ -1,0 +1,312 @@
+library(RISmed)
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+# http://amunategui.github.io/pubmed-query/
+
+# Define function for getting results from pubmed
+pubmed <- function(start_year = 2014,
+                   end_year = 2016,
+                   search_topic = 'malaria',
+                   counts_only = TRUE,
+                   n_records = 5000){
+  
+  # Define a year range
+  years <- start_year:end_year
+  
+  
+  # If counts only, simply get the counts by year
+  if(counts_only){
+    # JUST COUNTING RESULTS
+    
+    # Create a placeholder for results
+    return_object <- data.frame(year = years,
+                                n = NA)
+    
+    for (year in 1:length(years)){
+      message(paste0('Fetching records for year ', years[year]))
+      
+      # Perform search
+      search_query <- try({ 
+        EUtilsSummary(search_topic, 
+                      db="pubmed", 
+                      mindate=years[year], 
+                      maxdate=years[year],
+                      retmax = n_records)})
+      if(inherits(search_query, 'try-error')){
+        n <- 0
+      } else {
+        n <- QueryCount(search_query)
+      }
+      
+      # Populate results dataframe
+      return_object$n[year] <- n
+    }
+    
+  } else {
+    # NOT JUST COUNTS. ACTUALLY RETRIEVE RESULTS
+    
+    # Create an empty list for sticking results
+    results_list <- list()
+    # Create another empty list for sticking abstracts
+    abstracts_list <- list()
+    
+    # Loop through each year to get results
+    for (year in 1:length(years)){
+      message(paste0('Fetching records for year ', years[year]))
+      
+      try({
+        # Perform search
+        search_query <- 
+          EUtilsSummary(query = search_topic, 
+                        db="pubmed", 
+                        mindate=years[year], 
+                        maxdate=years[year],
+                        retmax = n_records)
+        
+        # See results summary
+        # summary(search_query)
+        
+        # See number of results
+        # QueryCount(search_query)
+        
+        # Get IDs of the articles returned in our query
+        # Qids <- queryId(search_query)
+        
+        # Actually fetch data
+        records <- EUtilsGet(search_query)
+        
+        # Turn into a data.frame
+        pubmed_data <- 
+          data.frame('title' = ArticleTitle(records),
[…]

### 2017-01-30 paper.Rmd
updated

--- 2016-11-09 paper.Rmd
@@
 ---
-title: Foreign direct investment, corporate social responsibility, and malaria control in Mozambique
-author: "Joe Brew and Elisa Sicuri"
-csl: health-economics-policy-and-law.csl
-output:
-  pdf_document:
-    includes:
-      in_header: style.sty
-  html_document: default
-highlight: zenburn
-latex_engine: xelatex
+title: Foreign direct investment, corporate social responsibility, and malaria control in Mozambique - trends, risks, and opportunities
+author:
+  - name: Joe Brew
+    email: joebrew@economicsofmalaria.com
+    affiliation: isglobal,cism,vu
+    footnote: Corresponding Author
+  - name: Elisa Sicuri
+    email: elisa.sicuri@isglobal.org 
+    affiliation: isglobal,icl,cism
+address:
+  - code: isglobal
+    address: Barcelona Institute for Global Health, c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain
+  - code: icl
+    address: Imperial College London, South Kensington Campus, London SW7 2AZ, U.K.
+  - code: cism
+    address: Centro de Investigação em Saúde de Manhiça, Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique
+  - code: vu
+    address: VU University Amsterdam, De Boelelaan 1105, 1081 HV Amsterdam, Netherlands
+abstract: |
+  Foreign direct investment (FDI) in Mozambique has increased rapidly in the last two decades.
+The growing interest in corporate social responsibility (CSR) - combined with a recent push for malaria elimination - suggest the need for a critical examination of the role of FDI in malaria control and elimination.
+Through a systematic review of the literature, we find that there has been a notable increase in research pertaining to FDI, CSR and malaria in recent years.
+Consensus among researchers is that this increase has not been accompanied by a substantial private sector investment in malaria control (with a few notable exceptions).
+This suggests a potential opportunity for scaling up privately driven malaria control as a means to both improve public health and increase return on investment.
+However, given the lack of coordination between public and private sectors, even when interests coincide, an over-reliance on foreign and private initiative for funding eradication is not without risks.
+
 bibliography: library.bib
+output: rticles::elsevier_article
 ---
@@
 # BASIC KNITR OPTIONS
-opts_chunk$set(comment = NA, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
+opts_chunk$set(comment = NA, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, fig.align = 'center', fig.height = 3, fig.width = 4)
 options(xtable.comment = FALSE)
@@
 
-\begin{changemargin}{3cm}{3cm}
-
-\begin{center}
-\textbf{Abstract}
-\end{center}
-
-\emph{Foreign direct investment in Mozambique has increased rapidly in the last two decades.
-The growing interest in corporate social responsibility - combined with a recent push for malaria eradication - suggest the need for a critical examination of how firms spend money on malaria control.
-The interests of both the private and public sectors align in the need for more private investment in malaria control.
-However, given the lack of coordination between public and private sectors, an over-reliance on foreign and private initiative for funding eradication is not without risks.}
-
-\end{changemargin}
-
 ## Introduction
-
-\columnsbegin
 
@@
 
-In 2015, the flood of FDI into Mozambique (and many other subsaharan African nations) slowed to a trickle.
+Beginning in late 2015, the flood of FDI into Mozambique (and many other subsaharan African nations) slowed to a trickle.
 The impact of this slowdown on CSR is not yet known, but it can reasonably be assumed that it will mean a reduction in CSR activities (albeit with lag).
@@
 
+We carried out a systematic review of economic and public health research pertaining to FDI, CSR, malaria and Mozambique.
 This paper gives an overview of trends in FDI and CSR in Mozambique, with a focus on its impact on malaria control.
-It is by no means comprehensive, but offers a consolidated starting point for understanding where the interests and incentives of the public and private sectors converge and differ in regards to malaria control.
+It is by no means comprehensive, but offers a consolidated starting point for understanding both the place of FDI and CSR in the literature, as well as where the interests and incentives of the public and private sectors converge and differ in regards to malaria control.
 
@@
 
-Despite the concentration of private investment in regional projects, growth has been similarly large in all districts.
-From 2000 to 2009, GDP approximately doubled, with the greatest growth occurring in Inhambane (119% growth from 2000 to 2009), and the least robust grwoth in Nampula (95%) [@knoema].
+Despite the concentration of private investment in regional projects, growth has been similarly large in all provinces  From 2000 to 2009, GDP approximately doubled, with the greatest growth occurring in Inhambane (119% growth from 2000 to 2009), and the least robust grwoth in Nampula (95%) [@knoema].
 
@@
   scale_colour_manual(guide = guide_legend(reverse=TRUE), 
-                      name = 'District',
+                      name = '',
                       values = brewer.pal(length(unique(ts$NAME_1)), 'Spectral')) +
-  ggtitle('District-specific growth in GDP')
+  ggtitle('Province-specific growth in GDP')
 ```
@@
 
-```{r}
+```{r, fig.width = 7, fig.height = 5}
 # Fortify moz1
@@
   # theme_void() +
-  ggtitle('(Smoothed) percent of households with private IRS') +
+  ggtitle('Privately supplied IRS') +
   theme(legend.position="bottom")
@@
 ```
+
+# Systematic review of FDI and CSR in health research
+
+## Search
+
+```{r}
+# PERFORM PUBMED SEARCH
+source('search_pubmed.R')
+
+if('search_results.RData' %in% dir()){
+  load('search_results.RData')
+} else {
+  csr_results <- pubmed(start_year = 2000,
+                         end_year = 2016,
+                         search_topic = '(malaria) and (corporate social responsibility)',
+                         counts_only = FALSE,
+                         n_records = 5000)  
+  fdi_results <- pubmed(start_year = 2000,
+                       end_year = 2016,
+                       search_topic = '(malaria) and (foreign direct investment)',
+                       counts_only = FALSE,
+                       n_records = 5000)  
+    moz_csr_results <- pubmed(start_year = 2000,
+                     end_year = 2016,
+                     search_topic = '(mozambique) and (corporate social responsibility)',
+                     counts_only = FALSE,
+                     n_records = 5000)  
+  moz_fdi_results <- pubmed(start_year = 2000,
+                     end_year = 2016,
+                     search_topic = '(mozambique) and (foreign direct investment)',
+                     counts_only = FALSE,
+                     n_records = 5000)  
+  save(csr_results,
+       fdi_results,
+       moz_csr_results,
+       moz_fdi_results,
+       file = 'search_results.RData')
+}
+
+```
+
+In order to gauge research attention and focus on FDI and CSR insofar as they affect Mozambique and malaria, four systematic searches were carried out using the EBSCOhost and NCBI/pubmed databases.
+Our search queries are detailed below:
+
+1. `"(malaria) and (corporate social responsibility)"`
+2. `"(malaria) and (foreign direct investment)"`
+3. `"(mozambique) and (corporate social responsibility)"`
+4. `"(mozambique) and (foreign direct investment)"`
+
+Our search is outlined in the below PRISMA-based diagram.
+
+\begin{center}
+\includegraphics[width=400pt]{img/prisma.png}
+\end{center}
+
+## Qualitative synthesis
+
+The literature is largely clear on two points: (1) that economic growth in Mozambique has been fueled in large part by foreign direct investment, and (2) that economic growth has been accompanied by a rapid decrease in malaria morbidity and mortality.
+There is no clear consensus regarding the extent to which the former is a result of the latter, and researchers disagree on how much of a role FDI has played in improving the social and economic conditions of Mozambicans.
+
+Mozambique has a unique social and economic system, which is particularly attractive to foreign investment.
+The Mozambican economy is "oriented to incentivize large-scale FDI projects" [@Robbins2012].
+Land distribution, for example, despite being igualitarian _de juro_ has been _de facto_ aimed at satisfying the interests of large foreign firms [@German2013].
+In extractive industries, and particular the sugar industry, the state has gone so far as to encourage unprofitable entreprise by propping up an internal market so as to keep foreign inflows of cash from drying out, a process described as a "mediating bureaucracy" [@Buur2012].
+Another part of the attraction of FDI to Mozambique is the extent to which the state and society, both formally and informally, subsidize the cost of labor by allowing for below subsistence wages.
+
+In one view, the last two decades of economic growth are largely irrelevant to the well-being of Mozambicans due to the "porous and extractive" nature of FDI [@CastelBranco2014].
+In other words, since both investment and profit are largely external to the Mozambican economy and society, "positive spillovers" into local firms are few and far between [@Winkler].
+
+CSR is one way of offsetting the trend, and its potential for mutually beneficial effects is high, particularly when its resources target malaria control.
+Per one analysis, malaria elimination in an endemic country like Mozambique would lead to a 16% increase in FDI [@Azemar2009].
+The possibility of re-directing resource rents to socially beneficial ends is acknolwedged by multiple authors [@Robbins2012, @CastelBranco2014].
+
+However, the re-direction or resource rents to CSR is complicated.
+If coerced (through higher taxes or fewer incentives to investment), the recent decrease in FDI may be exacerbated.
+If voluntary, particularly in the case of expenditures in medicine and health, the public's health would be largely dependent on private whim, a recipe for unsustainability at best and potential epidemiologic catastrophe at worst.
+
+In summary, Mozambique's high ratio of FDI to GDP and general allowance of "porous" foreign investment represents an opportunity for scaling up the private sector's engagement with public health, as a means to offset the lack of social benefits that traditional FDI has entailed.
+A coercive (tax-based) approach to scaling up this engagement may lead to a withdrawl of FDI from the country.
+The alternative, an increase in CSR, is a possible route for creating synergies, particular in the case of malaria elimination.
 
@@
 
-\columnsend
-
 ******

### 2017-05-22 paper.Rmd
big update

@@
 output: rticles::elsevier_article
+header-includes:
+     - \usepackage{longtable}
 ---
@@
 Improvements in health have accompanied economic expansion, but Mozambique still lags behind in basic health outcomes - particularly those related to malaria - even by regional standards.
-Despite  strong evidence showing that improved health yields substantial economic benefits [@World1999, @Bloom2008], direct foreign investment (FDI) under the auspices of "corporate social responsibility" (CSR) has largely ignored malaria prevention in favor of other social good activities.
+Most economic activities have to deal
+with malaria: particularly agricultural-based and extraction activities cannot be carried out without consistent investment towards malaria management, both
+prevention and prompt treatment. This can be done as part of the “normal activities” of private sector or can go slightly beyond what is strictly needed for the
+activities and be seen as part of the corporate social responsibilityDespite  strong evidence showing that improved health yields substantial economic benefits [@World1999, @Bloom2008], there is little evidence regarding the extent to which "corporate social responsibility" (CSR) activities of foreign firms have targeted or impacted malaria.
+Understanding the relationship between FDI, CSR and malaria is essential to identifying which incentives and barriers to private sector involvement in malaria elimination exist, as well as identifying potential opportunities and synergies for scaling up private sector involvment in malaria elimination.
 
+That said, 
 The few malaria-related CSR projects in recent years has come in large, "mega-projects" whose well-publicized malaria abatement activities are profit-driven [@Mouzin2011] or whose primary aim is not profit-related [@Han].
@@
 
+Private foreign investors showed a great optimism during the FMI meeting held in Maputo in 2014 (reference needed).
 Beginning in late 2015, the flood of FDI into Mozambique (and many other subsaharan African nations) slowed to a trickle.
 The impact of this slowdown on CSR is not yet known, but it can reasonably be assumed that it will mean a reduction in CSR activities (albeit with lag).
+Almost in parallel to this, Mozambique has had serious financial problems which have
+resulted in default (reference needed http://www.economist.com/news/middleeast-and-africa/21715030-mozambique-fails-pay-its-debtsmozambiques-default).
 Given the rapidly changing economic and epidemiologic context in Mozambique, a comprehensive and current understanding of both (a) the landscape of FDI and CSR in Mozambique and (b) public health issues (namely, malaria control) which are directly affected those investments is sorely needed.
@@
 We carried out a systematic review of economic and public health research pertaining to FDI, CSR, malaria and Mozambique.
-This paper gives an overview of trends in FDI and CSR in Mozambique, with a focus on its impact on malaria control.
+This paper gives an overview of trends in FDI and CSR in Mozambique as well as a synthesis of academic literature on the topic, with a focus on its impact on malaria control.
 It is by no means comprehensive, but offers a consolidated starting point for understanding both the place of FDI and CSR in the literature, as well as where the interests and incentives of the public and private sectors converge and differ in regards to malaria control.
 
-## FDI and CSR in Mozambique
+# Methods
+
+```{r}
+# PERFORM PUBMED SEARCH
+source('search_pubmed.R')
+
+if('search_results.RData' %in% dir()){
+  load('search_results.RData')
+} else {
+  csr_results <- pubmed(start_year = 2000,
+                         end_year = 2016,
+                         search_topic = '(malaria) and (corporate social responsibility)',
+                         counts_only = FALSE,
+                         n_records = 5000)  
+  fdi_results <- pubmed(start_year = 2000,
+                       end_year = 2016,
+                       search_topic = '(malaria) and (foreign direct investment)',
+                       counts_only = FALSE,
+                       n_records = 5000)  
+    moz_csr_results <- pubmed(start_year = 2000,
+                     end_year = 2016,
+                     search_topic = '(mozambique) and (corporate social responsibility)',
+                     counts_only = FALSE,
+                     n_records = 5000)  
+  moz_fdi_results <- pubmed(start_year = 2000,
+                     end_year = 2016,
+                     search_topic = '(mozambique) and (foreign direct investment)',
+                     counts_only = FALSE,
+                     n_records = 5000)  
+  save(csr_results,
+       fdi_results,
+       moz_csr_results,
+       moz_fdi_results,
+       file = 'search_results.RData')
+}
+pubmed_results <- rbind(csr_results$results %>% mutate(type = 'CSR'),
+                        fdi_results$results %>% mutate(type = 'FDI'),
+                        moz_csr_results$results %>% mutate(type = 'CSR MOZ'))
+
+```
+
+We sought to identify and describe sources of information regarding FDI and CSR in Mozambique, particularly in regards to malaria control and elimination.
+We carried out this identification process via 2 distinct approaches to understanding FDI, CSR and malaria in Mozambique:
+
+1. Identification and analysis of quantitative datasets.
+2. Systematic review of both grey literature (news, company websites, etc.) and academic literature.
+
+## Quantitative datasets
+
+We examined and analyzed trends from multiple datasets related to FDI, CSR and malaria in Mozambique.
+These included the "Doing Business" data pertaining to the measurement of regulatory efficiency, the GADM, the Deutsche Bundesbank Data Repository for foreign exchange rate history, the Knoema World Data Atlas for macroeconomic trends, the World Bank data portal for information containing to net foreign inflows and FDI, USAID Demographic and Health Survey data on sociodemographics and health-related practices, the Instituto Nacional de Estatistica for granular data pertaining to Mozambique's population and economic activities, and Institute for Health Metrics and Evaluation data pertaining to disease trends over time.
+
+## Systematic review
+
+### Grey literature
+
+Our grey literature review followed known practices [@Godin2015, @Adams2016], relying on internet searches and linked references.
+Sources include, but are not limited to UN reports on economic trends, WHO reports and brochures pertaining to health trends, company websites and brochures, newspaper articles, and think-tank white papers.
+
+### Academic literature
+
+In order to gauge research attention and focus on FDI and CSR insofar as they affect Mozambique and malaria, four systematic searches were carried out using the EBSCOhost and NCBI/pubmed databases.
+Our search queries are detailed below:
+
+1. `"(malaria) and (corporate social responsibility)"` (no results found in EBSCO)
+2. `"(malaria) and (foreign direct investment)"`
+3. `"(mozambique) and (corporate social responsibility)"` (no resuts found in EBSCO)
+4. `"(mozambique) and (foreign direct investment)"`
+
+# Results
+
+## Quantitiative datasets pertaining to FDI and CSR in Mozambique
 
@@
     facet_grid(.~key) +
-    theme_bw()
+    theme_bw() +
+  theme(strip.text = element_text(size=5))
 g2 <- 
@@
     facet_grid(.~key) +
-    theme_bw()
+    theme_bw() +
+   theme(strip.text = element_text(size=5))
+
 multiplot(g1, g2)  
@@
 
-# Systematic review of FDI and CSR in health research
+## Systematic review
 
-## Search
+### Grey literature
 
-```{r}
-# PERFORM PUBMED SEARCH
-source('search_pubmed.R')
+#### Grey literature search strategy
 
-if('search_results.RData' %in% dir()){
-  load('search_results.RData')
-} else {
-  csr_results <- pubmed(start_year = 2000,
-                         end_year = 2016,
-                         search_topic = '(malaria) and (corporate social responsibility)',
-                         counts_only = FALSE,
-                         n_records = 5000)  
-  fdi_results <- pubmed(start_year = 2000,
-                       end_year = 2016,
-                       search_topic = '(malaria) and (foreign direct investment)',
-                       counts_only = FALSE,
-                       n_records = 5000)  
-    moz_csr_results <- pubmed(start_year = 2000,
-                     end_year = 2016,
-                     search_topic = '(mozambique) and (corporate social responsibility)',
-                     counts_only = FALSE,
-                     n_records = 5000)  
-  moz_fdi_results <- pubmed(start_year = 2000,
-                     end_year = 2016,
-                     search_topic = '(mozambique) and (foreign direct investment)',
-                     counts_only = FALSE,
-                     n_records = 5000)  
-  save(csr_results,
-       fdi_results,
-       moz_csr_results,
-       moz_fdi_results,
-       file = 'search_results.RData')
-}
+We devised 2 simple search queries, and use www.google.com and www.bing.com to retrieve results. The queries were:
 
-```
+1. `Mozambique foreign direct investment malaria`
+2. `Mozambique corporate social responsibility malaria`
 
-In order to gauge research attention and focus on FDI and CSR insofar as they affect Mozambique and malaria, four systematic searches were carried out using the EBSCOhost and NCBI/pubmed databases.
-Our search queries are detailed below:
+#### Grey literature search results
 
-1. `"(malaria) and (corporate social responsibility)"`
-2. `"(malaria) and (foreign direct investment)"`
-3. `"(mozambique) and (corporate social responsibility)"`
-4. `"(mozambique) and (foreign direct investment)"`
+Google returned 266,000 results for the former query, and 240,000 for the latter (506,000 in total); Bing returned 4,620 and 20,900, respectively.
+We screened the top 50 results from both services (a total of 100 items) for both queries For the former, of the 100 items identified, 33 were identical in both search engines (yielding a total of 67 unique items); in the latter, 39 were identical (yielding a total of 61 unique items).
+
+\begin{center}
+\includegraphics[width=400pt]{img/prisma_grey.png}
+\end{center}
+
+#### Grey literature synthesis
+
+Our grey literature review returned relatively little information which was not readily available in those sources uncovered in the data review and systematic review.
+Very few foreign companies operating in Mozambique have any publicly available information regarding corporate social responsibility activities.
+
+One notable exception is the Nando's-lead Goodbye Malaria Trust, an umbrella organization which inclues a development impact bond, the Goodbye Malaria initiative, and partnerships with several other foreign firms operating in Mozambique (more details here).
+
+### Academic literature
 
@@
 
-## Qualitative synthesis
+The below table shows the results retrieved for the systematic review. Full article information is in the bibliography
+
+```{r}
+search_results <-
+  pubmed_results %>%
+  mutate(source = 'Pubmed') %>%
+  dplyr::select(source, year, title, type, `last _author_last_name`) %>%
+  rename(author = `last _author_last_name`)
+
+# Manually add in Elisa's search results
+econ_results <-
+  data_frame(source = rep('EBSCOhost', 20),
+             year = NA,
+             title = NA,
+             type = NA,
+             author = NA)
+econ_results[1,2:5] <- c(2014, 
+                         'A Critical Review of FDI Inflows and Economic Growth in Low-Income SADC Countries: Prospects and Challenges',
+                         'FDI MOZ',
+                         'Mahembe')
+econ_results[2,2:5] <- c(2000, 'Administrative barriers to foreign investment: Reducing red tape in Africa', 'FDI MOZ', 'Emery')
+econ_results[3,2:5] <- c(2012,
+                         'Attracting Foreign Direct Investment in Developing Countries: Determinants and Policies-A Comparative Study between Mozambique and China',
+                         'FDI MOZ', 'Tembe')
+econ_results[4,2:5] <- c(2013,
+                         'Contemporary Processes of Large-Scale Land Acquisition in Sub-Saharan Africa: Legal Deficiency or Elite Capture of the Rule of Law?',
+                         'FDI MOZ', 'German')
+econ_results[5,2:5] <- c(2012,
+                         'Corruption and Multinational Companies Entry Modes--Do Linguistic and Historical Ties Matter?',
+                         'FDI MOZ', 'Grande')
+econ_results[6,2:5] <- c(2014,
+                         'Determining the Nature and Extent of Spillovers: Empirical Assessment',
+                         'FDI MOZ', 'Winkler')
+econ_results[7,2:5] <- c(1999, 'Foreign Direct Investment in Africa: Policies Also Matter',
+                         'FDI MOZ', 'Morisset')
+econ_results[8,2:5] <- c(2002,
+                         'Foreign Direct Investment in Africa--Some Case Studies',
+                         'FDI MOZ', 'Bassu')
+econ_results[9,2:5] <- c(2014,
+                         'Growth, Capital Accumulation and Economic Porosity in Mozambique: Social Losses, Private Gains',
+                         'FDI MOZ', 'Castel-Branco')
+econ_results[10,2:5] <- c(2007,
+                          'Linkage Between foreign Direct Investment, Trade and Trade Policy: An Economic Analysis with Application to the Food Sector in OECD Countries and Case Studies in Ghana, Mozambique, Tunisia and Uganda',
+                          'FDI MOZ', 
+                          'Wilson')
+econ_results[11,2:5] <- c(2012,
+                          'Mining FDI and Infrastructure Development on Africas East Coast: Examining the Recent Experience of Tanzania and Mozambique',
+                          'FDI MOZ',
+                          'Robbins')
+econ_results[12,2:5] <- c(2013,
+                          'Potential and actual FDI spillovers in global value chains: the role of foreign investor characteristics, absorptive capacity and transmission channels',
+                          'FDI MOZ',
+                          'Winkler')
+econ_results[13,2:5] <- c(2004,
+                          'Regional integration, FDI and competitiveness in southern Africa',
+                          'FDI MOZ',
+                          'Goldstein')
+econ_results[14,2:5] <- c(2014,
+                          'Sector Case Study: Mining',
+                          'FDI MOZ',
+                          'Barnard')
+econ_results[15,2:5] <- c(2011,
+                          'Strategic Privatisation: Rehabilitating the Mozambican Sugar Industry',
+                          'FDI MOZ',
+                          'Buur')
+econ_results[16,2:5] <- c(2016,
+                          'The Economics and Politics of Local Content in African Extractives: Lessons from Tanzania, Uganda and Mozambique',
+                          'FDI MOZ',
+                          'Hansen')
+econ_results[17,2:5] <- c(2002,
+                          'The Role of FDI in Economic Development',
+                          'FDI MOZ',
+                          'Bjorvatn')
+econ_results[18,2:5] <- c(2010,
+                          'Uncovering Trends in the Accumulation of Technological Capabilities and Skills in the Mozambican Manufacturing Sector',
+                          'FDI MOZ',
+                          'Warren-Rodriguez')
+econ_results[19,2:5] <- c(2009,
+                          'Public Governance, Health and Foreign Direct Investment in Sub-Saharan Africa',
+                          'FDI Malaria',
+                          'Azemar')
+econ_results[20,2:5] <- c(2014,
+                          'Sector Case Study: Agribusiness',
+                          'FDI MOZ',
+                          'Barnard')
+
+# Combine all results
+econ_results$year <- as.numeric(as.character(econ_results$year))
+search_results <-
+  bind_rows(search_results,
+            econ_results)
+
+# Clean up
+names(search_results) <- Hmisc::capitalize(names(search_results))
+search_results$Author <- paste0(search_results$Author, ', et al')
+
+# Make title shorter
+add_n <- function(x, n = 20){
+  nc <- nchar(x)
+  if(nc < (n+15)){
+    return(x)
+  } else {
+    starts <- seq(1, nc, n)
+  stops <- c(seq(n, nc, n), nc)
+  out <- list()
+  for (i in 1:length(starts)){
+    while(stops[i] <= nc &
+          substr(x, stops[i], stops[i]) != ' '){
+      stops <- stops + 1
+    }
+    out[[i]] <- substr(x = x, start = starts[i], stop = stops[i])
+        while(!is.na(starts[i+1]) &
+              starts[i + 1] <= stops[i]){
+      starts[i + 1] <- starts[i + 1] +1
+    }
+  }
+  out <- out[!out == '']
+  final <- ''
+  for (i in 1:(length(out) -1)){
+    final <- paste0(final, trimws(out[[i]], 'both'), '\n', collapse = '')
+  }
+  final <- paste0(final, out[[length(out)]])
+  return(final)
+
+  }
+}
+
+# Make short funtion
+make_short <- function(x, n = 20){
+  if(nchar(x) < n){
+    return(x)
+  } else {
+  return(paste0(substr(x, 1, n), '...', collapse = ''))
+  }
+}
+for (s in 1:nrow(search_results)){
+  search_results$Title[s] <- 
+    make_short(search_results$Title[s])
+    # add_n(search_results$Title[s], n = 60)
+}
+search_results$Year <- as.character(search_results$Year)
+search_results <- search_results %>% arrange(Source, Type)
+```
+
+```{r, results = 'asis'}
+# kable(search_results)
+options(xtable.comment = FALSE)
+print(xtable(search_results),
+      size="\\fontsize{6pt}{7pt}\\selectfont",
+      include.rownames = FALSE,
+      include.colnames = TRUE,
+      caption.placement = "top",
+      comment=FALSE,
+      floating=FALSE,
+      tabular.environment='longtable')
+```
+
+The below chart shows our systematically retried publications by date.
+What is most striking is how little academic attention has been given to CSR, with none given to the intersection of CSR and malaria.
+
+```{r}
+x <- search_results %>%
+  group_by(Year = as.numeric(Year), Type) %>%
+  tally
+
+cols <- colorRampPalette(brewer.pal(n = 9, 'Spectral'))(length(unique(x$Type)))
+ggplot(data = x,
+       aes(x = Year,
+           y = n)) +
+  geom_bar(color = 'darkgrey',
+           stat = 'identity',
+           position = 'stack',
+           aes(#group = Type,
+                 fill = Type)) +
+  scale_fill_manual(name = 'Type', values = cols) +
+  theme_bw() +
+  labs(y = 'Articles',
+       title = 'Systematic review articles by publication date')
+```
+
+### Qualitative synthesis of systematic review of academic literature
 
@@
 
-# Scaling up malaria control through CSR: opportunity and risk
+# Discussion
+
+## Scaling up malaria control through CSR: opportunity and risk
 

### 2017-05-22 paper.tex
big update

--- 2017-01-30 paper.Rmd
@@
----
-title: Foreign direct investment, corporate social responsibility, and malaria control in Mozambique - trends, risks, and opportunities
-author:
-  - name: Joe Brew
-    email: joebrew@economicsofmalaria.com
-    affiliation: isglobal,cism,vu
-    footnote: Corresponding Author
-  - name: Elisa Sicuri
-    email: elisa.sicuri@isglobal.org 
-    affiliation: isglobal,icl,cism
-address:
-  - code: isglobal
-    address: Barcelona Institute for Global Health, c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain
-  - code: icl
-    address: Imperial College London, South Kensington Campus, London SW7 2AZ, U.K.
-  - code: cism
-    address: Centro de Investigação em Saúde de Manhiça, Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique
-  - code: vu
-    address: VU University Amsterdam, De Boelelaan 1105, 1081 HV Amsterdam, Netherlands
-abstract: |
-  Foreign direct investment (FDI) in Mozambique has increased rapidly in the last two decades.
-The growing interest in corporate social responsibility (CSR) - combined with a recent push for malaria elimination - suggest the need for a critical examination of the role of FDI in malaria control and elimination.
-Through a systematic review of the literature, we find that there has been a notable increase in research pertaining to FDI, CSR and malaria in recent years.
-Consensus among researchers is that this increase has not been accompanied by a substantial private sector investment in malaria control (with a few notable exceptions).
-This suggests a potential opportunity for scaling up privately driven malaria control as a means to both improve public health and increase return on investment.
-However, given the lack of coordination between public and private sectors, even when interests coincide, an over-reliance on foreign and private initiative for funding eradication is not without risks.
+\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
+%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+\usepackage[hyphens]{url}
+\usepackage{lineno} % add
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
-bibliography: library.bib
-output: rticles::elsevier_article
----
+\bibliographystyle{elsarticle-harv}
+\biboptions{sort&compress} % For natbib
+\usepackage{graphicx}
+\usepackage{booktabs} % book-quality tables
+%% Redefines the elsarticle footer
+%\makeatletter
+%\def\ps@pprintTitle{%
+% \let\@oddhead\@empty
+% \let\@evenhead\@empty
+% \def\@oddfoot{\it \hfill\today}%
+% \let\@evenfoot\@oddfoot}
+%\makeatother
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE }
-# No scientific notation
-options(scipen=999)
+% A modified page layout
+\textwidth 6.75in
+\oddsidemargin -0.15in
+\evensidemargin -0.15in
+\textheight 9in
+\topmargin -0.5in
+%%%%%%%%%%%%%%%% end my additions to header
 
-#####
-# PACKAGES
-#####
-library(maptools)
-library(maps)
-library(dplyr)
-library(xtable)
-library(ggplot2)
-library(knitr)
-library(png)
-library(grid)
-library(extrafont)
-library(raster)
-library(rgdal)
-library(readxl)
-library(tidyr)
-library(readr)
-library(RColorBrewer)
-library(ggmap)
-library(rasterVis)
-library(colorspace)
-library(dismo)
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={Foreign direct investment, corporate social responsibility, and malaria control in Mozambique - trends, risks, and opportunities},
+            colorlinks=true,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{0}
+\usepackage{longtable}
+% Pandoc toggle for numbering sections (defaults to be off)
+\setcounter{secnumdepth}{0}
+% Pandoc header
+\usepackage{longtable}
 
-loadfonts(device = 'postscript', quiet = TRUE)
+\usepackage[nomarkers]{endfloat}
 
-# BASIC KNITR OPTIONS
-opts_chunk$set(comment = NA, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, fig.align = 'center', fig.height = 3, fig.width = 4)
-options(xtable.comment = FALSE)
+\begin{document}
+\begin{frontmatter}
 
-# HELPER FUNCTIONS
-source('helpers.R')
-```
+  \title{Foreign direct investment, corporate social responsibility, and malaria
+control in Mozambique - trends, risks, and opportunities}
+    \author[isglobal,cism,vu]{Joe Brew\corref{c1}}
+   \ead{joebrew@economicsofmalaria.com} 
+   \cortext[c1]{Corresponding Author}
+    \author[isglobal,icl,cism]{Elisa Sicuri}
+   \ead{elisa.sicuri@isglobal.org}
 
-```{r}
-#####
-# GET MAPS OF MOZAMBIQUE
-#####
-setwd('data')
-# This will get three different administrative
-# levels of Mozambican shapefiles from the GADM database
-# using the raster library
-for (i in 0:3){
-  assign(paste0('moz', i),
-         getData('GADM', country = 'MOZ', level = i))
-}
-setwd('..')
-```
+      \address[isglobal]{Barcelona Institute for Global Health, c/ Rosselló, 132, 5è 2a. 08036,
+Barcelona, Spain}
+    \address[icl]{Imperial College London, South Kensington Campus, London SW7 2AZ, U.K.}
+    \address[cism]{Centro de Investigação em Saúde de Manhiça, Vila da Manhiça, Bairro
+Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique}
+    \address[vu]{VU University Amsterdam, De Boelelaan 1105, 1081 HV Amsterdam,
+Netherlands}
 
-```{r, fig.height = 3, fig.align='center', eval = FALSE}
-plot(moz0)
-```
+  \begin{abstract}
+  Foreign direct investment (FDI) in Mozambique has increased rapidly in
+  the last two decades. The growing interest in corporate social
+  responsibility (CSR) - combined with a recent push for malaria
+  elimination - suggest the need for a critical examination of the role of
+  FDI in malaria control and elimination. Through a systematic review of
+  the literature, we find that there has been a notable increase in
+  research pertaining to FDI, CSR and malaria in recent years. Consensus
+  among researchers is that this increase has not been accompanied by a
+  substantial private sector investment in malaria control (with a few
+  notable exceptions). This suggests a potential opportunity for scaling
+  up privately driven malaria control as a means to both improve public
+  health and increase return on investment. However, given the lack of
+  coordination between public and private sectors, even when interests
+  coincide, an over-reliance on foreign and private initiative for funding
+  eradication is not without risks.
+  \end{abstract}
 
-## Introduction
+ \end{frontmatter}
 
-Mozabique's recent economic growth has been facilitated by the government's open policy towards foreign investment, a plentiful supply of natural resources (particularly minerals and hydrocarbons) [@Rogers] and relatively inexpensive labor.
-Improvements in health have accompanied economic expansion, but Mozambique still lags behind in basic health outcomes - particularly those related to malaria - even by regional standards.
-Despite  strong evidence showing that improved health yields substantial economic benefits [@World1999, @Bloom2008], direct foreign investment (FDI) under the auspices of "corporate social responsibility" (CSR) has largely ignored malaria prevention in favor of other social good activities.
+\subsection{Introduction}\label{introduction}
 
-The few malaria-related CSR projects in recent years has come in large, "mega-projects" whose well-publicized malaria abatement activities are profit-driven [@Mouzin2011] or whose primary aim is not profit-related [@Han].
-Though the former is often portrayed as a "win-win" for business and public health, the latter also offers tangible benefits for private industry, and should be understood as operating under the same conditions and with the same motivations.
+Mozabique's recent economic growth has been facilitated by the
+government's open policy towards foreign investment, a plentiful supply
+of natural resources (particularly minerals and hydrocarbons) (Rogers
+2014) and relatively inexpensive labor. Improvements in health have
+accompanied economic expansion, but Mozambique still lags behind in
+basic health outcomes - particularly those related to malaria - even by
+regional standards. Most economic activities have to deal with malaria:
+particularly agricultural-based and extraction activities cannot be
+carried out without consistent investment towards malaria management,
+both prevention and prompt treatment. This can be done as part of the
+``normal activities'' of private sector or can go slightly beyond what
+is strictly needed for the activities and be seen as part of the
+corporate social responsibilityDespite strong evidence showing that
+improved health yields substantial economic benefits (Brundtland 1999,
+Bloom and Canning (2008)), there is little evidence regarding the extent
+to which ``corporate social responsibility'' (CSR) activities of foreign
+firms have targeted or impacted malaria. Understanding the relationship
+between FDI, CSR and malaria is essential to identifying which
+incentives and barriers to private sector involvement in malaria
+elimination exist, as well as identifying potential opportunities and
+synergies for scaling up private sector involvment in malaria
+elimination.
 
-Beginning in late 2015, the flood of FDI into Mozambique (and many other subsaharan African nations) slowed to a trickle.
-The impact of this slowdown on CSR is not yet known, but it can reasonably be assumed that it will mean a reduction in CSR activities (albeit with lag).
-Given the rapidly changing economic and epidemiologic context in Mozambique, a comprehensive and current understanding of both (a) the landscape of FDI and CSR in Mozambique and (b) public health issues (namely, malaria control) which are directly affected those investments is sorely needed.
-Such an understanding may foster greater private interest in for-profit investment in public health measures.
-Likewise, it can facilitate better public sector understanding of potential industry partners and stakeholders (a prerequisite to greater public-private collaborations), and guide the public sector away from an inflexible dependence on FDI for the provision of public health necessities.
+That said, The few malaria-related CSR projects in recent years has come
+in large, ``mega-projects'' whose well-publicized malaria abatement
+activities are profit-driven (Mouzin and al. 2011) or whose primary aim
+is not profit-related (Han 2015). Though the former is often portrayed
+as a ``win-win'' for business and public health, the latter also offers
+tangible benefits for private industry, and should be understood as
+operating under the same conditions and with the same motivations.
 
-We carried out a systematic review of economic and public health research pertaining to FDI, CSR, malaria and Mozambique.
-This paper gives an overview of trends in FDI and CSR in Mozambique, with a focus on its impact on malaria control.
-It is by no means comprehensive, but offers a consolidated starting point for understanding both the place of FDI and CSR in the literature, as well as where the interests and incentives of the public and private sectors converge and differ in regards to malaria control.
+Private foreign investors showed a great optimism during the FMI meeting
+held in Maputo in 2014 (reference needed). Beginning in late 2015, the
+flood of FDI into Mozambique (and many other subsaharan African nations)
+slowed to a trickle. The impact of this slowdown on CSR is not yet
+known, but it can reasonably be assumed that it will mean a reduction in
+CSR activities (albeit with lag). Almost in parallel to this, Mozambique
+has had serious financial problems which have resulted in default
+(reference needed
+http://www.economist.com/news/middleeast-and-africa/21715030-mozambique-fails-pay-its-debtsmozambiques-default).
+Given the rapidly changing economic and epidemiologic context in
+Mozambique, a comprehensive and current understanding of both (a) the
+landscape of FDI and CSR in Mozambique and (b) public health issues
+(namely, malaria control) which are directly affected those investments
+is sorely needed. Such an understanding may foster greater private
+interest in for-profit investment in public health measures. Likewise,
+it can facilitate better public sector understanding of potential
+industry partners and stakeholders (a prerequisite to greater
+public-private collaborations), and guide the public sector away from an
+inflexible dependence on FDI for the provision of public health
+necessities.
 
-## FDI and CSR in Mozambique
+We carried out a systematic review of economic and public health
+research pertaining to FDI, CSR, malaria and Mozambique. This paper
+gives an overview of trends in FDI and CSR in Mozambique as well as a
+synthesis of academic literature on the topic, with a focus on its
+impact on malaria control. It is by no means comprehensive, but offers a
+consolidated starting point for understanding both the place of FDI and
+CSR in the literature, as well as where the interests and incentives of
+the public and private sectors converge and differ in regards to malaria
+control.
 
-### Massive increase in FDI
+\section{Methods}\label{methods}
 
-Following independence (1974), Mozambique saw two decades of low and unsteady foreign investment, largely due to the civil war (which did not end until 1992).
-Thereafter, foreign investment began a steady increase but leveled off by the late 1990s.
-However, the discovery of novel sources of oil and gas set off a new spur of investments beginning in 2007, and continuing through last year.
-From 2010 to 2013, foreign direct investment grew from 1.26 to 6.70 billion USD, a more than five-fold increase [@wbdata].
+We sought to identify and describe sources of information regarding FDI
+and CSR in Mozambique, particularly in regards to malaria control and
+elimination. We carried out this identification process via 2 distinct
+approaches to understanding FDI, CSR and malaria in Mozambique:
 
-```{r}
-#####
-# READ IN WORLD BANK FOREIGN INVESTMENT DATA
-#####
-# http://data.worldbank.org/indicator/BX.KLT.DINV.CD.WD
-wb <- read_excel('data/bx.klt.dinv.cd.wd_Indicator_en_excel_v2.xls',
-                 skip = 3)
-wb <- data.frame(wb)
-names(wb[5:ncol(wb)]) <- paste0('y', names(wb[5:ncol(wb)]))
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Identification and analysis of quantitative datasets.
+\item
+  Systematic review of both grey literature (news, company websites,
+  etc.) and academic literature.
+\end{enumerate}
 
-# Clean up to just get mozambique
-temp <- gather(wb, year, value, X1961:X2015)
-# Clean up year column
-temp$year <- as.numeric(gsub('X', '', temp$year))
-# Subset to only include Mozambique
-temp <- temp[temp$Country.Name == 'Mozambique',]
-# Remove years without data
-temp <- temp[temp$year >= 1970 & temp$year <= 2014,]
-temp <- temp[!is.na(temp$value),]
-temp <- temp[temp$value > 0,]
+\subsection{Quantitative datasets}\label{quantitative-datasets}
 
-#####
-# PLOT FOREIGN DIRECT INVESTMENT NET INFLOWS
-#####
-ggplot(data = temp, aes(x = year, y = value)) +
-    scale_y_log10() +
-  geom_point(alpha = 0.5, size = 1) +
-  geom_line(alpha = 0.5, size = 1) +
-  theme_bw() +
-  xlab('Year') +
-  ylab('Current USD (logarithmic scale)') +
-  ggtitle('Foreign direct investment: net inflows')
-```
+We examined and analyzed trends from multiple datasets related to FDI,
+CSR and malaria in Mozambique. These included the ``Doing Business''
+data pertaining to the measurement of regulatory efficiency, the GADM,
+the Deutsche Bundesbank Data Repository for foreign exchange rate
+history, the Knoema World Data Atlas for macroeconomic trends, the World
+Bank data portal for information containing to net foreign inflows and
+FDI, USAID Demographic and Health Survey data on sociodemographics and
+health-related practices, the Instituto Nacional de Estatistica for
+granular data pertaining to Mozambique's population and economic
+activities, and Institute for Health Metrics and Evaluation data
+pertaining to disease trends over time.
 
-In addition to the discovery of new resources, recent growth has also been fueled by political and economic reforms which have made it easier for foreigners to do business in Mozambique.
-Of particular note, inflation remained relatively low (at least through 2014) [@deutsche].
+\subsection{Systematic review}\label{systematic-review}
 
-```{r}
-# Load up forex data
-if('forex.csv' %in% dir('data')){
-  forex <- read_csv('data/forex.csv')
-} else {
-  library(Quandl)
-  forex <- Quandl("BUNDESBANK/BBEX3_M_MZN_USD_CA_AC_A01")
-  write_csv(forex, 'data/forex.csv')
-}
+\subsubsection{Grey literature}\label{grey-literature}
 
-# Include only through Jan 2015
-forex <- forex[forex$Date <= '2015-01-01',]
+Our grey literature review followed known practices (Godin et al. 2015,
+Adams et al. (2016)), relying on internet searches and linked
+references. Sources include, but are not limited to UN reports on
+economic trends, WHO reports and brochures pertaining to health trends,
+company websites and brochures, newspaper articles, and think-tank white
+papers.
 
-ggplot(data = forex,
-       aes(x = Date, y = Value)) +
-  geom_line(alpha = 0.5, size = 1) +
-  theme_bw() +
-  ggtitle('Meticals per US dollar') +
-  ylab('Meticals') +
-  geom_point(alpha = 0.5, size = 1) +
-  ylim(0, 40)
-```
+\subsubsection{Academic literature}\label{academic-literature}
 
-The massive increases in FDI have also been facilitated by dramatic decreases in the costs to import and export, as well as the costs of starting a business and registering property [@doingbusiness].
+In order to gauge research attention and focus on FDI and CSR insofar as
+they affect Mozambique and malaria, four systematic searches were
+carried out using the EBSCOhost and NCBI/pubmed databases. Our search
+queries are detailed below:
 
-```{r, comment = NA, message = FALSE, results = 'hide'}
-######
-# READ IN DOING BUSINESS DATA
-#####
-db <- read_excel('data/doing_business.xls')
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \texttt{"(malaria)\ and\ (corporate\ social\ responsibility)"} (no
+  results found in EBSCO)
+\item
+  \texttt{"(malaria)\ and\ (foreign\ direct\ investment)"}
+\item
+  \texttt{"(mozambique)\ and\ (corporate\ social\ responsibility)"} (no
+  resuts found in EBSCO)
+\item
+  \texttt{"(mozambique)\ and\ (foreign\ direct\ investment)"}
+\end{enumerate}
 
-# Clean up columns
-db <- data.frame(db)
-macro_names <- names(db)
-for (j in 1:length(macro_names)){
-  number <- j
-  while(grepl('Var', macro_names[number])){
-    number <- number -1
-  }
-  macro_names[j] <- macro_names[number]
-}
-micro_names <- as.character(db[1,])
-combined_names <- gsub(' ', '', paste0(macro_names, '..', ifelse(is.na(micro_names), '', micro_names)))
-# Remove extra columns and replace column names
-db <- db[-1,]
-names(db) <- combined_names
+\section{Results}\label{results}
 
-# Clean up year
-db$year <- as.numeric(gsub('DB', '', db$Year.., fixed = TRUE))
+\subsection{Quantitiative datasets pertaining to FDI and CSR in
+Mozambique}\label{quantitiative-datasets-pertaining-to-fdi-and-csr-in-mozambique}
 
-# Get my columns of interest
-db$cost_export_container <- db[,"Trading.Across.Borders..Costtoimport(deflatedUS$percontainer)"]
-db$start_business <- db[,"Starting.a.Business..Time(days)"]
-db$cost_import_container <- db[,"Trading.Across.Borders..Costtoimport(deflatedUS$percontainer)" ]
-db$register_property <- 
-  db[,"Registering.Property..Cost(%ofpropertyvalue)"]
+\subsubsection{Massive increase in FDI}\label{massive-increase-in-fdi}
 
-db <- db[,c('year', 'cost_export_container', 'cost_import_container', 'start_business', 'register_property')]
+Following independence (1974), Mozambique saw two decades of low and
+unsteady foreign investment, largely due to the civil war (which did not
+end until 1992). Thereafter, foreign investment began a steady increase
+but leveled off by the late 1990s. However, the discovery of novel
+sources of oil and gas set off a new spur of investments beginning in
+2007, and continuing through last year. From 2010 to 2013, foreign
+direct investment grew from 1.26 to 6.70 billion USD, a more than
+five-fold increase (WB 2015).
 
-# Clean up
-temp <- gather(db, key, value, cost_export_container:register_property)
-temp$value <- as.numeric(temp$value)
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1} \end{center}
 
-# Clean up names/labels
-temp$key <- factor(temp$key,
-                   labels = c('Cost to export container (deflated USD)',
-                              'Cost to import container (deflated USD)',
-                              'Cost (% of GDP per capita) to start business',
-                              'Cost (% of GDP per capita) to register property'))
+In addition to the discovery of new resources, recent growth has also
+been fueled by political and economic reforms which have made it easier
+for foreigners to do business in Mozambique. Of particular note,
+inflation remained relatively low (at least through 2014) (Bundesbank
+2015).
 
-# PLOT
-g1 <- 
-  ggplot(data = temp[temp$key %in% unique(sort(temp$key))[1:2],], 
-         aes(x = year, y = value)) +
-    geom_line() +
-    facet_grid(.~key) +
-    theme_bw()
-g2 <- 
-  ggplot(data = temp[temp$key %in% unique(sort(temp$key))[3:4],], 
-         aes(x = year, y = value)) +
-    geom_line() +
-    facet_grid(.~key) +
-    theme_bw()
-multiplot(g1, g2)  
-```
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1} \end{center}
 
-### Breakdown by industry
+The massive increases in FDI have also been facilitated by dramatic
+decreases in the costs to import and export, as well as the costs of
+starting a business and registering property (Bank 2016).
 
-Most of recent growth has come in the "extractive" industries, a term encompassing a range of industry, but in the Mozambican context largely applying to hydrocarbons and mining [@estatistica].
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1} \end{center}
 
-```{r}
-#####
-# MANUALLY ENTER DATA FROM INSTiTUtE OF STATISTICS
-#####
-industry <- data.frame(var = c('extractivas', 
-                               'alimentares',
-                               'madeira',
-                               'coque',
-                               'naometalicos',
-                               'metalicos',
-                               'equipamentos'),
-                       description = c('Extractive',
-                                       'Food, drink, tobacco',
-                                       'Wood, paper, printing',
-                                       'Chemical, petroluem products',
-                                       'Non-metalic minerals',
-                                       'Metalic minerals',
-                                       'Equipment (including vehicles)'))
-industry[,paste0('y', 2011:2014)] <- NA
+\subsubsection{Breakdown by industry}\label{breakdown-by-industry}
 
-industry[industry$var == 'extractivas' ,paste0('y', 2011:2014)] <-
-  c(100, 194.4, 203.3, 230.1)
-industry[industry$var == 'alimentares' ,paste0('y', 2011:2014)] <-
-  c(100, 100.3, 147.2, 214.1)
-industry[industry$var == 'madeira' ,paste0('y', 2011:2014)] <-
-  c(100, 93.5, 103.3, 118.5)
-industry[industry$var == 'coque' ,paste0('y', 2011:2014)] <-
-  c(100, 105.6, 100.2, 79.5)
-industry[industry$var == 'naometalicos' ,paste0('y', 2011:2014)] <-
-  c(100, 102.9, 110, 131.5)
-industry[industry$var == 'metalicos' ,paste0('y', 2011:2014)] <-
-  c(100, 78.9, 83.2, 120.7)
-industry[industry$var == 'equipamentos' ,paste0('y', 2011:2014)] <-
-  c(100, 73.8, 167.5, 25.7)
+Most of recent growth has come in the ``extractive'' industries, a term
+encompassing a range of industry, but in the Mozambican context largely
+applying to hydrocarbons and mining (INE 2015).
 
-# Clean up a bit
-temp <- gather(industry, year, value, y2011:y2014)
-temp$year <- as.numeric(gsub('y', '', temp$year))
-temp$Industry <- temp$description
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1} \end{center}
 
-ggplot(data = temp, aes(x = year, y = value, color = Industry)) +
-  geom_line(size = 1, alpha = 0.5) +
-  geom_point(aes(group = Industry)) +
-  theme_bw() +
-  geom_hline(yintercept = 100) +
-  scale_color_brewer(palette = 'Set1') +
-  xlab('Year') +
-  ylab('Percent of 2011 Volume') +
-  ggtitle('Industry volume index') +
-  theme(legend.position="bottom") +
-  guides(col = guide_legend(nrow = 3))
+The late 2015 economic slow-down in the developing world, particularly
+the low price of oil, could have serious repercussions for FDI in
+Mozambique. That said, the mining of metals and the service industries
+both make up a larger share of Mozambique's economic output than the
+extraction of hydrocarbons, which should somewhat buffer the Mozambican
+economy from the negative effects of low oil prices.
 
-```
+\subsubsection{Breakdown by region}\label{breakdown-by-region}
 
-The late 2015 economic slow-down in the developing world, particularly the low price of oil, could have serious repercussions for FDI in Mozambique.
-That said, the mining of metals and the service industries both make up a larger share of Mozambique's economic output than the extraction of hydrocarbons, which should somewhat buffer the Mozambican economy from the negative effects of low oil prices.
+Despite the concentration of private investment in regional projects,
+growth has been similarly large in all provinces From 2000 to 2009, GDP
+approximately doubled, with the greatest growth occurring in Inhambane
+(119\% growth from 2000 to 2009), and the least robust grwoth in Nampula
+(95\%) (Knoema 2015).
 
-### Breakdown by region
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1} \end{center}
 
-Despite the concentration of private investment in regional projects, growth has been similarly large in all provinces  From 2000 to 2009, GDP approximately doubled, with the greatest growth occurring in Inhambane (119% growth from 2000 to 2009), and the least robust grwoth in Nampula (95%) [@knoema].
+The homogeneity in growth comes somewhat as a surprise, as it defies the
+general developing world pattern in that growth in areas that already
+had high GDP (Maputo and the coastal provinces) was as robust as growth
+in areas with previously low GDP.
 
-```{r}
-#####
-# READ IN GDP BY REGION/TIME DATA FROM
-# http://knoema.com/atlas/Mozambique/ranks/GDP-at-Constant-Prices
-#####
-gdp <- read_csv('data/gdp_by_region.csv')
-names(gdp)[1] <- 'NAME'
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1} \end{center}
 
-# Clean names to match map
-gdp$NAME[which(gdp$NAME == 'Niassa')] <- 'Nassa'
+\subsubsection{CSR in Mozambique}\label{csr-in-mozambique}
 
-# Remove city
-gdp <- gdp[gdp$NAME != 'Maputo cidade',]
+Corporate social responsibility in Mozambique is new in nomenclature,
+but activities which could be classified as CSR have existed for
+decades. The number of firms actively engaging in CSR cannot be
+ascertained (given the large and ever-changing number of small
+businesses), but virtually all of the largest firms, both foreign and
+domestic, have a CSR component.
 
-# Join map and data
-temp <- moz1
-temp@data <- left_join(x = temp@data, y = gdp, by = c('NAME_1' = 'NAME'))
+Firms with CSR activities are often large and foreign. Among the largest
+``key players'' in Mozambican CSR are Coca-Cola, British Petroleum and
+Colgate-Palmolive. State conglomerates, such as Águas de Moçambique and
+Electricidade de Moçambique, also participate in CSR activities (Compact
+2007).
 
-##### PLOT TS
-ts <- temp@data
-ts <- gather(ts, year, gdp, y2009:y2000)
-ts$year <- as.numeric(gsub('y', '', ts$year))
-ts$year <- as.Date(paste0(ts$year, '-01-01'))
+\subsubsection{Malaria-related CSR and other corporate
+activity}\label{malaria-related-csr-and-other-corporate-activity}
 
-# Remove Maputo city
-ts <- ts[ts$NAME_1 != 'Maputo City',]
+Most CSR-funded activities are focused on education, community
+development, women's rights and entrepreneurship. According to a UN
+poll, none of the country's largest firms invest directly in
+malaria-related CSR activities (Compact 2007). This may be due to the
+perceived costs of malaria control, lack of perceived PR benefits, the
+government and NGO's predominant role in the area, as well as the
+``opt-in'' nature and privacy/legal issues generally related to engaging
+in health-related campaigns.
 
-# Arrange
-ts <- ts %>% arrange(year)
+That said, a majority of the firms interviewed by the UN indicated that
+one of the principal reasons for investment in CSR is to complement
+government efforts. To the extent that malaria accounts for more of the
+loss in disability-adjusted life years in Mozambique than comparable
+countries (IHME 2015), malaria control's lack of representation among
+core CSR activities is a notable absence.
 
-# Rebase to 100 %
-ts <- ts %>%
-  left_join(ts %>% 
-              group_by(NAME_1) %>%
-              summarise(start = first(gdp))) %>%
-  mutate(gdp = gdp / start * 100)
+Whether under the guise of CSR or not, it is noteworthy that the private
+sector currently does play a role in malaria control activities.
+According to 2011 DHS data, greater than 7 in every 1,000 households had
+a private company carry out indoor residual spraying (DHS 2011). And, in
+some clusters, the percent of houses covered by private IRS was greater
+than 25\%.
 
-ggplot(data = ts, 
-       aes(x = year, y = gdp, col = NAME_1)) +
-  geom_line(alpha = 0.6) +
-  geom_point() +
-  xlab('Year') +
-  ylab('GDP (% of 2000 value)') +
-  theme_bw() +
-  scale_colour_manual(guide = guide_legend(reverse=TRUE), 
-                      name = '',
-                      values = brewer.pal(length(unique(ts$NAME_1)), 'Spectral')) +
-  ggtitle('Province-specific growth in GDP')
-```
+Interestingly, though the UN data indicated that corporations don't
+actively engage in malaria control as part of CSR, the geographic
+distribution of households which had their homes sprayed by a private
+entity are clustered in areas where foreign firms operate, particularly
+in the south (around Maputo) in the East, where extractive industry
+activity is highest.
 
-The homogeneity in growth comes somewhat as a surprise, as it defies the general developing world pattern in that growth in areas that already had high GDP (Maputo and the coastal provinces) was as robust as growth in areas with previously low GDP.
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-12-1} \end{center}
 
-```{r}
-#####
-# PLOT MAPS
-#####
+\subsection{Systematic review}\label{systematic-review-1}
 
-# Get columns for coordinates
-temp$x <- coordinates(temp)[,1]
-temp$y <- coordinates(temp)[,2]
+\subsubsection{Grey literature}\label{grey-literature-1}
 
-# Fortify
-temp_fort <- fortify(temp, id = 'ID_0')
+\paragraph{Grey literature search
+strategy}\label{grey-literature-search-strategy}
 
-# Bring in gdp data to fortified map
-temp_fort <- left_join(x = temp_fort,
-                       y = temp@data %>%
-                         mutate(id = as.character(ID_1)))
+We devised 2 simple search queries, and use www.google.com and
+www.bing.com to retrieve results. The queries were:
 
-# Gather by year
-temp_fort <- gather(temp_fort, year, gdp, y2009:y2000)
-temp_fort$year <- as.numeric(gsub('y', '', temp_fort$year))
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \texttt{Mozambique\ foreign\ direct\ investment\ malaria}
+\item
+  \texttt{Mozambique\ corporate\ social\ responsibility\ malaria}
+\end{enumerate}
 
-ggplot(data = temp_fort,
-       aes(x = long, 
-           y = lat, 
-           group = group,
-           fill = gdp)) +
-  geom_polygon() +
-  coord_map() +
-  facet_wrap(~year) +
-  xlab('Longitude') +
-  ylab('Latitude') +
-  theme_bw() +
-  scale_fill_gradient(low = 'yellow',
-                      high = 'darkblue',
-                      # guide = guide_legend(reverse=TRUE), 
-                      name = 'GDP')
-# cols <- colorRampPalette(brewer.pal(9, 'Blues'))(ceiling(max(temp$y2009, na.rm = TRUE)))
-# colorfy <- function(number){
-#   cols[round(number)]
-# }
-# 
-# par(mfrow = c(1,3))
-# par(mar = c(4,0,2,1))
-# plot(temp, col = colorfy(temp$y2000), border = NA); title(main = '2000')
-# plot(temp, col = colorfy(temp$y2005), border = NA); title(main = '2005')
-# plot(temp, col = colorfy(temp$y2009), border = NA); title(main = '2009')
-# par(mar = c(5,4,2,2))
-# 
-# legend_vec <- seq(8000, 24000, 8000)
-# legend('right',
-#        fill = cols[legend_vec],
-#        bty = 'n',
-#        border = NA,
-#        legend = paste0(legend_vec/1000,'K'),
-#        ncol = 1,
-#        title = 'GDP')
-```
+\paragraph{Grey literature search
+results}\label{grey-literature-search-results}
 
-### CSR in Mozambique
+Google returned 266,000 results for the former query, and 240,000 for
+the latter (506,000 in total); Bing returned 4,620 and 20,900,
+respectively. We screened the top 50 results from both services (a total
+of 100 items) for both queries For the former, of the 100 items
+identified, 33 were identical in both search engines (yielding a total
+of 67 unique items); in the latter, 39 were identical (yielding a total
+of 61 unique items).
 
-Corporate social responsibility in Mozambique is new in nomenclature, but activities which could be classified as CSR have existed for decades.
-The number of firms actively engaging in CSR cannot be ascertained (given the large and ever-changing number of small businesses), but virtually all of the largest firms, both foreign and domestic, have a CSR component.
+\begin{center}
+\includegraphics[width=400pt]{img/prisma_grey.png}
+\end{center}
 
-Firms with CSR activities are often large and foreign.
-Among the largest "key players" in Mozambican CSR are Coca-Cola, British Petroleum and Colgate-Palmolive.
-State conglomerates, such as Águas de Moçambique and Electricidade de Moçambique, also participate in CSR activities [@Compact2007].
+\paragraph{Grey literature synthesis}\label{grey-literature-synthesis}
 
-### Malaria-related CSR and other corporate activity
+Our grey literature review returned relatively little information which
+was not readily available in those sources uncovered in the data review
+and systematic review. Very few foreign companies operating in
+Mozambique have any publicly available information regarding corporate
+social responsibility activities.
 
-Most CSR-funded activities are focused on education, community development, women's rights and entrepreneurship.
-According to a UN poll, none of the country's largest firms invest directly in malaria-related CSR activities [@Compact2007].
-This may be due to the perceived costs of malaria control, lack of perceived PR benefits, the government and NGO's predominant role in the area, as well as the "opt-in" nature and privacy/legal issues generally related to engaging in health-related campaigns.
+One notable exception is the Nando's-lead Goodbye Malaria Trust, an
+umbrella organization which inclues a development impact bond, the
+Goodbye Malaria initiative, and partnerships with several other foreign
+firms operating in Mozambique (more details here).
 
-That said, a majority of the firms interviewed by the UN indicated that one of the principal reasons for investment in CSR is to complement government efforts.
-To the extent that malaria accounts for more of the loss in disability-adjusted life years in Mozambique than comparable countries [@ihme], malaria control's lack of representation among core CSR activities is a notable absence.
-
-Whether under the guise of CSR or not, it is noteworthy that the private sector currently does play a role in malaria control activities.
-According to 2011 DHS data, greater than 7 in every 1,000 households had a private company carry out indoor residual spraying [@dhs].
-And, in some clusters, the percent of houses covered by private IRS was greater than 25%.
-
-Interestingly, though the UN data indicated that corporations don't actively engage in malaria control as part of CSR, the geographic distribution of households which had their homes sprayed by a private entity are clustered in areas where foreign firms operate, particularly in the south (around Maputo) in the East, where extractive industry activity is highest.
-
-```{r}
-# READ IN DHS DATA
-# (generated in dhs/exploration.Rmd -  https://github.com/joebrew/dhs/)
-dhs <- read_csv('data/dhs_private_irs.csv')
-
-# Get mean by cluster
-by_cluster <-
-  dhs %>%
-  group_by(cluster) %>%
-  summarise(lat = first(lat),
-            lon = first(lon),
-            urban_rural = first(urban_rural),
-            n = sum(n),
-            private = mean(private, na.rm = TRUE),
-            weight = first(weight)) %>%
-  filter(lat < -10, lon >20)
-
-# Get weighted mean
-x = weighted.mean(by_cluster$private, w = by_cluster$weight)
-
-# Make a spatial version
-houses <- 
-  by_cluster %>%
-  mutate(latitude = lat,
-         longitude = lon)
-coordinates(houses) <- ~longitude + latitude
-proj4string(houses) <- proj4string(moz0)
-
-# Get a smoothed surface for private IRS coverage
-if('irs_grid.RData' %in% dir()){
-  load('irs_grid.RData')
-} else {
-
-  # Create dataset of gridded values
-  moz <- map('world', 'moz', plot = FALSE)
-  lons <- seq(min(moz$x, na.rm = TRUE), 
-              max(moz$x, na.rm = TRUE), 
-              length = 100)
-  lats <- seq(min(moz$y, na.rm = TRUE), 
-              max(moz$y, na.rm = TRUE), 
-              length = 100)
-  df_grid <- expand.grid(lon = lons,
-                         lat = lats)
-
-  # Create latitude and longitude columns and convert to spatial
-  df_grid$latitude <- df_grid$lat
-  df_grid$longitude <- df_grid$lon
-  coordinates(df_grid) <- ~longitude + latitude
-  proj4string(df_grid) <- proj4string(moz0)
-
-  # Keep only those points within mozambique
-  x <- over(df_grid, polygons(moz0))
-  df_grid <- df_grid[!is.na(x),]
-
-  # Create a value in df_grid
-  df_grid$val <- NA
-
-  # Loop through the grid, getting a weighte dvalue of mosquito net or not
-  nn <- nrow(df_grid)
-  for (i in 1:nn){
-      message(paste0('row ', i,
-                     ' of ', nn))
-
-    # Get distance from specific point in df_grid 
-    # to every point in houses
-    distances <- spDistsN1(pts = houses,
-                          pt = df_grid[i,],
-                          longlat = TRUE)
-    # Define which are acceptably close
-    close_enough <- which(distances <= 50000)
-    # Get an val score (weighted likelihood of having net)
-    val <- stats::weighted.mean(x = houses$private[close_enough],
-                         w = ((1 / distances[close_enough]) ^2) *
-                           houses$weight[close_enough],
-                         na.rm = TRUE)
-    # Assign val to the dataframe
-    df_grid$val[i] <- val
-  }
-  save('df_grid', file = 'irs_grid.RData')
-}
-
-# Convert to raster
-temp <- df_grid@data %>% arrange(lon, lat)
-r <- rasterFromXYZ(temp[, c('lon', 'lat', 'val')])
-proj4string(r) <- proj4string(moz1)
-# levelplot(r, contour = TRUE, par.settings=RdBuTheme)
-
-# # Convert raster to polygons
-# rtp <- rasterToPolygons(r)
-# rtp@data$id <- 1:nrow(rtp@data) 
-```
-
-```{r, fig.width = 7, fig.height = 5}
-# Fortify moz1
-moz2_fortified <- fortify(moz2, region = 'NAME_2')
-
-# Plot clusters
-cols <- brewer.pal(9, 'Spectral')
-ggplot(data = moz2_fortified,
-       aes(x = long, y = lat)) +
-  # coord_cartesian() + 
-  geom_raster(data = df_grid@data,
-              aes(x = lon, y = lat, fill = val * 100)) +
-    scale_fill_gradientn(name="Percent",colours = cols) +
-    coord_equal() + 
-    # geom_polygon(aes(group = id), 
-    #              color = 'black',
-    #              fill = NA,
-    #              size = 0.1,
-    #              alpha = 0.3) +
-  theme_bw() +
-  xlab('Longitude') +
-  ylab('Latitude') +
-  # theme_void() +
-  ggtitle('Privately supplied IRS') +
-  theme(legend.position="bottom")
-
-```
-
-# Systematic review of FDI and CSR in health research
-
-## Search
-
-```{r}
-# PERFORM PUBMED SEARCH
-source('search_pubmed.R')
-
-if('search_results.RData' %in% dir()){
-  load('search_results.RData')
-} else {
-  csr_results <- pubmed(start_year = 2000,
-                         end_year = 2016,
-                         search_topic = '(malaria) and (corporate social responsibility)',
-                         counts_only = FALSE,
-                         n_records = 5000)  
-  fdi_results <- pubmed(start_year = 2000,
-                       end_year = 2016,
-                       search_topic = '(malaria) and (foreign direct investment)',
-                       counts_only = FALSE,
-                       n_records = 5000)  
-    moz_csr_results <- pubmed(start_year = 2000,
-                     end_year = 2016,
-                     search_topic = '(mozambique) and (corporate social responsibility)',
-                     counts_only = FALSE,
-                     n_records = 5000)  
-  moz_fdi_results <- pubmed(start_year = 2000,
-                     end_year = 2016,
-                     search_topic = '(mozambique) and (foreign direct investment)',
-                     counts_only = FALSE,
-                     n_records = 5000)  
-  save(csr_results,
-       fdi_results,
-       moz_csr_results,
-       moz_fdi_results,
-       file = 'search_results.RData')
-}
-
-```
-
-In order to gauge research attention and focus on FDI and CSR insofar as they affect Mozambique and malaria, four systematic searches were carried out using the EBSCOhost and NCBI/pubmed databases.
-Our search queries are detailed below:
-
-1. `"(malaria) and (corporate social responsibility)"`
-2. `"(malaria) and (foreign direct investment)"`
-3. `"(mozambique) and (corporate social responsibility)"`
-4. `"(mozambique) and (foreign direct investment)"`
+\subsubsection{Academic literature}\label{academic-literature-1}
 
@@
 
-## Qualitative synthesis
+The below table shows the results retrieved for the systematic review.
+Full article information is in the bibliography
 
-The literature is largely clear on two points: (1) that economic growth in Mozambique has been fueled in large part by foreign direct investment, and (2) that economic growth has been accompanied by a rapid decrease in malaria morbidity and mortality.
-There is no clear consensus regarding the extent to which the former is a result of the latter, and researchers disagree on how much of a role FDI has played in improving the social and economic conditions of Mozambicans.
+\begingroup\fontsize{6pt}{7pt}\selectfont
 
-Mozambique has a unique social and economic system, which is particularly attractive to foreign investment.
-The Mozambican economy is "oriented to incentivize large-scale FDI projects" [@Robbins2012].
-Land distribution, for example, despite being igualitarian _de juro_ has been _de facto_ aimed at satisfying the interests of large foreign firms [@German2013].
-In extractive industries, and particular the sugar industry, the state has gone so far as to encourage unprofitable entreprise by propping up an internal market so as to keep foreign inflows of cash from drying out, a process described as a "mediating bureaucracy" [@Buur2012].
-Another part of the attraction of FDI to Mozambique is the extent to which the state and society, both formally and informally, subsidize the cost of labor by allowing for below subsistence wages.
+\begin{longtable}{lllll}
+  \hline
+Source & Year & Title & Type & Author \\ 
+  \hline
+EBSCOhost & 2009 & Public Governance, H... & FDI Malaria & Azemar, et al \\ 
+  EBSCOhost & 2014 & A Critical Review of... & FDI MOZ & Mahembe, et al \\ 
+  EBSCOhost & 2000 & Administrative barri... & FDI MOZ & Emery, et al \\ 
+  EBSCOhost & 2012 & Attracting Foreign D... & FDI MOZ & Tembe, et al \\ 
+  EBSCOhost & 2013 & Contemporary Process... & FDI MOZ & German, et al \\ 
+  EBSCOhost & 2012 & Corruption and Multi... & FDI MOZ & Grande, et al \\ 
+  EBSCOhost & 2014 & Determining the Natu... & FDI MOZ & Winkler, et al \\ 
+  EBSCOhost & 1999 & Foreign Direct Inves... & FDI MOZ & Morisset, et al \\ 
+  EBSCOhost & 2002 & Foreign Direct Inves... & FDI MOZ & Bassu, et al \\ 
+  EBSCOhost & 2014 & Growth, Capital Accu... & FDI MOZ & Castel-Branco, et al \\ 
+  EBSCOhost & 2007 & Linkage Between fore... & FDI MOZ & Wilson, et al \\ 
+  EBSCOhost & 2012 & Mining FDI and Infra... & FDI MOZ & Robbins, et al \\ 
+  EBSCOhost & 2013 & Potential and actual... & FDI MOZ & Winkler, et al \\ 
+  EBSCOhost & 2004 & Regional integration... & FDI MOZ & Goldstein, et al \\ 
+  EBSCOhost & 2014 & Sector Case Study: M... & FDI MOZ & Barnard, et al \\ 
+  EBSCOhost & 2011 & Strategic Privatisat... & FDI MOZ & Buur, et al \\ 
+  EBSCOhost & 2016 & The Economics and Po... & FDI MOZ & Hansen, et al \\ 
+  EBSCOhost & 2002 & The Role of FDI in E... & FDI MOZ & Bjorvatn, et al \\ 
+  EBSCOhost & 2010 & Uncovering Trends in... & FDI MOZ & Warren-Rodriguez, et al \\ 
+  EBSCOhost & 2014 & Sector Case Study: A... & FDI MOZ & Barnard, et al \\ 
+  Pubmed & 2004 & Community health out... & CSR & Singer, et al \\ 
+  Pubmed & 2014 & Public-private partn... & CSR & Hutton, et al \\ 
+  Pubmed & 2007 & Feasibility of water... & CSR MOZ & Tang, et al \\ 
+  Pubmed & 2002 & The economic impact ... & FDI & Mills, et al \\ 
+  Pubmed & 2004 & The economic burden ... & FDI & Russell, et al \\ 
+  Pubmed & 2012 & The economic benefit... & FDI & Feachem, et al \\ 
+  Pubmed & 2012 & Global health fundin... & FDI & D'Agostino, et al \\ 
+  Pubmed & 2015 & Tracking Global Fund... & FDI & Huntington, et al \\ 
+   \hline
+\hline
+\end{longtable}
 
-In one view, the last two decades of economic growth are largely irrelevant to the well-being of Mozambicans due to the "porous and extractive" nature of FDI [@CastelBranco2014].
-In other words, since both investment and profit are largely external to the Mozambican economy and society, "positive spillovers" into local firms are few and far between [@Winkler].
+\endgroup
 
-CSR is one way of offsetting the trend, and its potential for mutually beneficial effects is high, particularly when its resources target malaria control.
-Per one analysis, malaria elimination in an endemic country like Mozambique would lead to a 16% increase in FDI [@Azemar2009].
-The possibility of re-directing resource rents to socially beneficial ends is acknolwedged by multiple authors [@Robbins2012, @CastelBranco2014].
+The below chart shows our systematically retried publications by date.
+What is most striking is how little academic attention has been given to
+CSR, with none given to the intersection of CSR and malaria.
 
-However, the re-direction or resource rents to CSR is complicated.
-If coerced (through higher taxes or fewer incentives to investment), the recent decrease in FDI may be exacerbated.
-If voluntary, particularly in the case of expenditures in medicine and health, the public's health would be largely dependent on private whim, a recipe for unsustainability at best and potential epidemiologic catastrophe at worst.
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-15-1} \end{center}
 
-In summary, Mozambique's high ratio of FDI to GDP and general allowance of "porous" foreign investment represents an opportunity for scaling up the private sector's engagement with public health, as a means to offset the lack of social benefits that traditional FDI has entailed.
-A coercive (tax-based) approach to scaling up this engagement may lead to a withdrawl of FDI from the country.
-The alternative, an increase in CSR, is a possible route for creating synergies, particular in the case of malaria elimination.
+\subsubsection{Qualitative synthesis of systematic review of academic
+literature}\label{qualitative-synthesis-of-systematic-review-of-academic-literature}
 
-# Scaling up malaria control through CSR: opportunity and risk
+The literature is largely clear on two points: (1) that economic growth
+in Mozambique has been fueled in large part by foreign direct
+investment, and (2) that economic growth has been accompanied by a rapid
+decrease in malaria morbidity and mortality. There is no clear consensus
+regarding the extent to which the former is a result of the latter, and
+researchers disagree on how much of a role FDI has played in improving
+the social and economic conditions of Mozambicans.
 
-### Opportunity
+Mozambique has a unique social and economic system, which is
+particularly attractive to foreign investment. The Mozambican economy is
+``oriented to incentivize large-scale FDI projects'' (Robbins and
+Perkins 2012). Land distribution, for example, despite being
+igualitarian \emph{de juro} has been \emph{de facto} aimed at satisfying
+the interests of large foreign firms (German, Schoneveld, and Mwangi
+2013). In extractive industries, and particular the sugar industry, the
+state has gone so far as to encourage unprofitable entreprise by
+propping up an internal market so as to keep foreign inflows of cash
+from drying out, a process described as a ``mediating bureaucracy''
+(Buur, Tembe, and Baloi 2012). Another part of the attraction of FDI to
+Mozambique is the extent to which the state and society, both formally
+and informally, subsidize the cost of labor by allowing for below
+subsistence wages.
 
-The historical absence of malaria-related CSR represents an opportunity for complementarity.
-In addition to improvements in public health, both the private and public sectors stand to benefit economically from a scaling up of malaria control driven by the private sector.
-By increasing malaria control activities, as a share of total CSR expenditures, public funds could be redirected towards other areas of health.
-Likewise, if private CSR activities pivoted towards malaria control rather more general philanthropical gestures, CSR would have a more direct impact on wellbeing (with less temporal lag), thereby fulfilling the public relations goals of the firms that invest in CSR.
-Finally, for many industries, the firm itself is a potential direct beneficiary, given that improvements in employee health and a reduction in employee absenteeism can be directly correlated to productivity.
+In one view, the last two decades of economic growth are largely
+irrelevant to the well-being of Mozambicans due to the ``porous and
+extractive'' nature of FDI (Castel-Branco 2014). In other words, since
+both investment and profit are largely external to the Mozambican
+economy and society, ``positive spillovers'' into local firms are few
+and far between (Winkler 2013).
 
-### Risk
+CSR is one way of offsetting the trend, and its potential for mutually
+beneficial effects is high, particularly when its resources target
+malaria control. Per one analysis, malaria elimination in an endemic
+country like Mozambique would lead to a 16\% increase in FDI (Azemar and
+Desbordes 2009). The possibility of re-directing resource rents to
+socially beneficial ends is acknolwedged by multiple authors (Robbins
+and Perkins 2012, Castel-Branco (2014)).
 
-Scaling up CSR while also encouraging its redirection towards malaria control is not without risks.
-The most notable downfall of this approach is the potential for the inadvertent dependence on the private sector for what is essentially a public good.
-Were CSR targeting malaria control to reach significant levels (and the government were to enact a correspoding redirection of funds towards the financing of other health areas), then a situation would be created in which the public sector had essentially divested from a public good.
-This would be unwise and dangerous.
+However, the re-direction or resource rents to CSR is complicated. If
+coerced (through higher taxes or fewer incentives to investment), the
+recent decrease in FDI may be exacerbated. If voluntary, particularly in
+the case of expenditures in medicine and health, the public's health
+would be largely dependent on private whim, a recipe for
+unsustainability at best and potential epidemiologic catastrophe at
+worst.
 
-A secondary risk is that increased private sector involvement in malaria control could cause a decrease in public sector _competence_ in the prevention and treatment of malaria.
-This could have negative consequences in the case of either a financial or economic crisis (in which CSR activity would be curtailed) or an increase in malarial activity.
-By the same token, CSR involvement in malaria control could potentially portend, at least initially, less effective interventions.
-Private firms' incentives, though aligned with the public's in terms of malaria, are not identical, and pressures from shareholders and for positive public perception might motivate malaria control strategies which do not necessarily carry with them the most recent scientific knowledge.
+In summary, Mozambique's high ratio of FDI to GDP and general allowance
+of ``porous'' foreign investment represents an opportunity for scaling
+up the private sector's engagement with public health, as a means to
+offset the lack of social benefits that traditional FDI has entailed. A
+coercive (tax-based) approach to scaling up this engagement may lead to
+a withdrawl of FDI from the country. The alternative, an increase in
+CSR, is a possible route for creating synergies, particular in the case
+of malaria elimination.
 
-A third risk is a lack of coordination.
-Both in terms of logistical activities as well as biological realities (drug and insecticide resistance, etc.), coordination of malaria control activities is absolutely essentially if Mozambique is going to make the transition to eradication.
-The issue of coordination could be solved through an activities and outcomes reporting/surveillance structure (necessarily managed by the state), but compliance could be problematic.
+\section{Discussion}\label{discussion}
 
-A final risk is that of volatility.
-By centralizing malaria control under the auspices of the government, public health authorities can effectively distribute malaria control expenditures to where and when they are most needed.
-If this control were only in the hands of private firms, expenditure would likely transform into a function of firm-specific profitability and shareholder incentives, as well as market cycles.
-This could lead to a situation in which malaria control activities are most prevalent in areas where the economy is strongest, rather than areas where the need is greatest.
+\subsection{Scaling up malaria control through CSR: opportunity and
+risk}\label{scaling-up-malaria-control-through-csr-opportunity-and-risk}
 
-# Conclusion
+\subsubsection{Opportunity}\label{opportunity}
 
-High FDI in Mozambique and growing interest in CSR both call for increased reflection on the private sector's role in the delivery of public goods.
-A refocusing of CSR expenditures into areas where the need is greatest (specifically, malaria control) could lead to better health and greater profits.
-This "win-win" for the public and private sectors represents a rare opportunity, which deserves more discourse, research and experimentation.
+The historical absence of malaria-related CSR represents an opportunity
+for complementarity. In addition to improvements in public health, both
+the private and public sectors stand to benefit economically from a
+scaling up of malaria control driven by the private sector. By
+increasing malaria control activities, as a share of total CSR
+expenditures, public funds could be redirected towards other areas of
+health. Likewise, if private CSR activities pivoted towards malaria
+control rather more general philanthropical gestures, CSR would have a
+more direct impact on wellbeing (with less temporal lag), thereby
+fulfilling the public relations goals of the firms that invest in CSR.
+Finally, for many industries, the firm itself is a potential direct
+beneficiary, given that improvements in employee health and a reduction
+in employee absenteeism can be directly correlated to productivity.
 
-That said, the growing interest in CSR suggests both that (a) the need for services exist and (b) that corporations (especially foreign firms) have enough excess capital to finance these services.
-Both of these factors suggest the need for a better-fitting taxation rate, and a more efficient delivery of public services.
-That said, in the short-term, increasing the efficiency and effectiveness of CSR activity through a re-pivoting towards malaria control should remain a goal.
+\subsubsection{Risk}\label{risk}
 
-******
+Scaling up CSR while also encouraging its redirection towards malaria
+control is not without risks. The most notable downfall of this approach
+is the potential for the inadvertent dependence on the private sector
+for what is essentially a public good. Were CSR targeting malaria
+control to reach significant levels (and the government were to enact a
+correspoding redirection of funds towards the financing of other health
+areas), then a situation would be created in which the public sector had
+essentially divested from a public good. This would be unwise and
+dangerous.
 
-## References
+A secondary risk is that increased private sector involvement in malaria
+control could cause a decrease in public sector \emph{competence} in the
+prevention and treatment of malaria. This could have negative
+consequences in the case of either a financial or economic crisis (in
+which CSR activity would be curtailed) or an increase in malarial
+activity. By the same token, CSR involvement in malaria control could
+potentially portend, at least initially, less effective interventions.
+Private firms' incentives, though aligned with the public's in terms of
+malaria, are not identical, and pressures from shareholders and for
+positive public perception might motivate malaria control strategies
+which do not necessarily carry with them the most recent scientific
+knowledge.
+
+A third risk is a lack of coordination. Both in terms of logistical
+activities as well as biological realities (drug and insecticide
+resistance, etc.), coordination of malaria control activities is
+absolutely essentially if Mozambique is going to make the transition to
+eradication. The issue of coordination could be solved through an
+activities and outcomes reporting/surveillance structure (necessarily
+managed by the state), but compliance could be problematic.
+
+A final risk is that of volatility. By centralizing malaria control
+under the auspices of the government, public health authorities can
+effectively distribute malaria control expenditures to where and when
+they are most needed. If this control were only in the hands of private
+firms, expenditure would likely transform into a function of
+firm-specific profitability and shareholder incentives, as well as
+market cycles. This could lead to a situation in which malaria control
+activities are most prevalent in areas where the economy is strongest,
+rather than areas where the need is greatest.
+
+\section{Conclusion}\label{conclusion}
+
+High FDI in Mozambique and growing interest in CSR both call for
+increased reflection on the private sector's role in the delivery of
+public goods. A refocusing of CSR expenditures into areas where the need
+is greatest (specifically, malaria control) could lead to better health
+and greater profits. This ``win-win'' for the public and private sectors
+represents a rare opportunity, which deserves more discourse, research
+and experimentation.
+
+That said, the growing interest in CSR suggests both that (a) the need
+for services exist and (b) that corporations (especially foreign firms)
+have enough excess capital to finance these services. Both of these
+factors suggest the need for a better-fitting taxation rate, and a more
+efficient delivery of public services. That said, in the short-term,
+increasing the efficiency and effectiveness of CSR activity through a
+re-pivoting towards malaria control should remain a goal.
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\subsection*{References}\label{references}
+\addcontentsline{toc}{subsection}{References}
+
+\hypertarget{refs}{}
+\hypertarget{ref-Adams2016}{}
+Adams, Jean, Frances C. Hillier-Brown, Helen J. Moore, Amelia A. Lake,
+Vera Araujo-Soares, Martin White, and Carolyn Summerbell. 2016.
+``Searching and Synthesising `Grey Literature' and `Grey Information' in
+Public Health: Critical Reflections on Three Case Studies.''
+\emph{Systematic Reviews} 5 (1). Springer Nature.
+doi:\href{https://doi.org/10.1186/s13643-016-0337-y}{10.1186/s13643-016-0337-y}.
+
+\hypertarget{ref-Azemar2009}{}
+Azemar, C., and R. Desbordes. 2009. ``Public Governance, Health and
+Foreign Direct Investment in Sub-Saharan Africa.'' \emph{Journal of
+African Economies} 18 (4). Oxford University Press (OUP): 667--709.
+doi:\href{https://doi.org/10.1093/jae/ejn028}{10.1093/jae/ejn028}.
+
+\hypertarget{ref-doingbusiness}{}
+Bank, World. 2016. \emph{Doing Business 2016: Measuring Regulatory
+Quality and Efficiency}. World Bank.
+\url{http://www.doingbusiness.org/~/media/GIAWB/Doing\%20Business/Documents/Annual-Reports/English/DB16-Chapters/DB16-Mini-Book.pdf}.
+
+\hypertarget{ref-Bloom2008}{}
+Bloom, David, and David Canning. 2008. ``Population Health and Economic
+Growth,'' 1--25.
+
+\hypertarget{ref-World1999}{}
+Brundtland, Gro Harlem. 1999. ``WHO on Health and Economic
+Productivity'' 25 (2): 396--402.
+
+\hypertarget{ref-deutsche}{}
+Bundesbank. 2015. ``Exchange Rates for the Us Dollar in Mozambique / Usd
+1 = Mzn.''
+\url{https://www.quandl.com/data/BUNDESBANK/BBEX3_M_MZN_USD_CA_AC_A01}.
+
+\hypertarget{ref-Buur2012}{}
+Buur, Lars, Carlota Mondlane Tembe, and Obede Baloi. 2012. ``The White
+Gold: The Role of Government and State in Rehabilitating the Sugar
+Industry in Mozambique.'' \emph{Journal of Development Studies} 48 (3).
+Informa UK Limited: 349--62.
+doi:\href{https://doi.org/10.1080/00220388.2011.635200}{10.1080/00220388.2011.635200}.
+
+\hypertarget{ref-CastelBranco2014}{}
+Castel-Branco, Carlos Nuno. 2014. ``Growth, Capital Accumulation and
+Economic Porosity in Mozambique: Social Losses, Private Gains.''
+\emph{Review of African Political Economy} 41 (sup1). Informa UK
+Limited: S26--S48.
+doi:\href{https://doi.org/10.1080/03056244.2014.976363}{10.1080/03056244.2014.976363}.
+
+\hypertarget{ref-Compact2007}{}
+Compact, U N Global. 2007. ``Corporate Social Responsibility: Country
+Report Mozambique.''
+\url{http://www.undp.org/content/dam/mozambique/docs/Poverty/UNDP}.
+
+\hypertarget{ref-dhs}{}
+DHS. 2011. ``USAID.''
+\url{http://dhsprogram.com/what-we-do/survey/survey-display-362.cfm}.
+
+\hypertarget{ref-German2013}{}
+German, Laura, George Schoneveld, and Esther Mwangi. 2013.
+``Contemporary Processes of Large-Scale Land Acquisition in Sub-Saharan
+Africa: Legal Deficiency or Elite Capture of the Rule of Law?''
+\emph{World Development} 48 (August). Elsevier BV: 1--18.
+doi:\href{https://doi.org/10.1016/j.worlddev.2013.03.006}{10.1016/j.worlddev.2013.03.006}.
+
+\hypertarget{ref-Godin2015}{}
+Godin, Katelyn, Jackie Stapleton, Sharon I. Kirkpatrick, Rhona M.
+Hanning, and Scott T. Leatherdale. 2015. ``Applying Systematic Review
+Search Methods to the Grey Literature: A Case Study Examining Guidelines
+for School-Based Breakfast Programs in Canada.'' \emph{Systematic
+Reviews} 4 (1). Springer Nature.
+doi:\href{https://doi.org/10.1186/s13643-015-0125-0}{10.1186/s13643-015-0125-0}.
+
+\hypertarget{ref-Han}{}
+Han, Lily. 2015. ``Malaria in Mozambique: trialling payment by
+results.''
+\url{http://www.theguardian.com/global-development-professionals-network/2014/mar/31/malaria-control-payment-by-results}.
+
+\hypertarget{ref-ihme}{}
+IHME. 2015. ``GBD Prifle: Mozambique.''
+\url{www.medbox.org/gbd-profile-mozambique/download.pdf}.
+
+\hypertarget{ref-estatistica}{}
+INE. 2015. ``Economic Statistics.'' \url{http://www.ine.gov.mz/}.
+
+\hypertarget{ref-knoema}{}
+Knoema. 2015. ``GDP of Mozambique by Region, Province and Country.''
+\url{http://knoema.com/atlas/Mozambique/ranks/GDP-at-Constant-Prices}.
+
+\hypertarget{ref-Mouzin2011}{}
+Mouzin, Eric, and Et al. 2011. ``Business Investing in Malaria Control:
+Economic Returns and a Healthy Workforce for Africa.'' \emph{Progress \&
+Impact Series}, no. 6.
+
+\hypertarget{ref-Robbins2012}{}
+Robbins, Glen, and David Perkins. 2012. ``MINING FDI AND INFRASTRUCTURE
+DEVELOPMENT ON AFRICAS EAST COAST: EXAMINING THE RECENT EXPERIENCE OF
+TANZANIA AND MOZAMBIQUE.'' \emph{Journal of International Development}
+24 (2). Wiley-Blackwell: 220--36.
+doi:\href{https://doi.org/10.1002/jid.2817}{10.1002/jid.2817}.
+
+\hypertarget{ref-Rogers}{}
+Rogers, Lina. 2014. ``Natural resources boom sustaining growth in
+Mozambique.''
+\url{http://www.abo.net/oilportal/topic/view.do?contentId=2195109}.
+
+\hypertarget{ref-wbdata}{}
+WB. 2015. ``Foreign Direct Investment, Net Inflows (Bop, Current
+Us\$).'' \url{http://data.worldbank.org/indicator/BX.KLT.DINV.CD.WD}.
+
+\hypertarget{ref-Winkler}{}
+Winkler, Deborah. 2013. ``Potential and Actual FDI Spillovers in Global
+Value Chains.''
+
+\end{document}

### 2017-05-22 paper_word.Rmd
big update

--- 2017-05-22 paper.Rmd
@@
 title: Foreign direct investment, corporate social responsibility, and malaria control in Mozambique - trends, risks, and opportunities
-author:
-  - name: Joe Brew
-    email: joebrew@economicsofmalaria.com
-    affiliation: isglobal,cism,vu
-    footnote: Corresponding Author
-  - name: Elisa Sicuri
-    email: elisa.sicuri@isglobal.org 
-    affiliation: isglobal,icl,cism
-address:
-  - code: isglobal
-    address: Barcelona Institute for Global Health, c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain
-  - code: icl
-    address: Imperial College London, South Kensington Campus, London SW7 2AZ, U.K.
-  - code: cism
-    address: Centro de Investigação em Saúde de Manhiça, Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique
-  - code: vu
-    address: VU University Amsterdam, De Boelelaan 1105, 1081 HV Amsterdam, Netherlands
-abstract: |
-  Foreign direct investment (FDI) in Mozambique has increased rapidly in the last two decades.
-The growing interest in corporate social responsibility (CSR) - combined with a recent push for malaria elimination - suggest the need for a critical examination of the role of FDI in malaria control and elimination.
-Through a systematic review of the literature, we find that there has been a notable increase in research pertaining to FDI, CSR and malaria in recent years.
-Consensus among researchers is that this increase has not been accompanied by a substantial private sector investment in malaria control (with a few notable exceptions).
-This suggests a potential opportunity for scaling up privately driven malaria control as a means to both improve public health and increase return on investment.
-However, given the lack of coordination between public and private sectors, even when interests coincide, an over-reliance on foreign and private initiative for funding eradication is not without risks.
+date: "April 5, 2017"
+output: word_document
+bibliography: library.bib
 
-bibliography: library.bib
-output: rticles::elsevier_article
-header-includes:
-     - \usepackage{longtable}
 ---
@@
 ```
+
+## Abstract
+
+Foreign direct investment (FDI) in Mozambique has increased rapidly in the last two decades.
+The growing interest in corporate social responsibility (CSR) - combined with a recent push for malaria elimination - suggest the need for a critical examination of the role of FDI in malaria control and elimination.
+Through a systematic review of the literature, we find that there has been a notable increase in research pertaining to FDI, CSR and malaria in recent years.
+Consensus among researchers is that this increase has not been accompanied by a substantial private sector investment in malaria control (with a few notable exceptions).
+This suggests a potential opportunity for scaling up privately driven malaria control as a means to both improve public health and increase return on investment.
+However, given the lack of coordination between public and private sectors, even when interests coincide, an over-reliance on foreign and private initiative for funding eradication is not without risks.
 

### 2017-07-11 paper.Rmd
big update

@@
 
+### Sources of FDI in Mozambique
+
+```{r}
+srcs <- readxl::read_excel('data/data.imf.orgSLASHregular.aspx?key=60564262.xls', skip = 4)
+# Rename columns
+srcs <- srcs[,1:3]
+srcs <- srcs[2:nrow(srcs),]
+names(srcs) <- c('from_country', 'inward', 'outward')
+# Make numberic
+srcs <- data.frame(srcs)
+for (j in 2:3){
+  this_column <- as.numeric(as.character(gsub('.', '', srcs[,j], fixed = TRUE)))
+  srcs[,j] <- this_column
+}
+srcs <- srcs %>%
+  filter(from_country != 'World')
+
+# Get amount by inward
+x <- srcs %>%
+  arrange(desc(inward)) %>%
+  filter(!is.na(inward) & inward > 0) %>%
+  mutate(from_country = ifelse(from_country == 'China, P.R.: Mainland',
+                               'China',
+                               ifelse(from_country == 'United Arab Emirates',
+                                      'U.A.E.',
+                                      ifelse(from_country == 'United Kingdom',
+                                             'U.K.',
+                                             from_country)))) %>%
+  mutate(from_country = factor(from_country, levels = from_country))
+```
+
+According to the IMF, `r nrow(x)` countries had foreign direct investment in Mozambique as of the end of 2015. The below chart shows the top 10 countries ranked by amount invested in Mozambique.
+
+```{r}
+ggplot(data = x[1:10,],
+       aes(x = from_country,
+           y = inward)) +
+  geom_bar(stat = 'identity',
+           alpha = 0.6,
+           fill = 'darkblue') +
+  labs(x = 'Source of investment',
+       y = '2015 USD, Millions',
+       title = 'Sources of FDI in Mozambique',
+       subtitle = 'By investor country') +
+  theme_bw() +
+    theme(axis.text.x = element_text(angle = 90)) 
+```
+
+What stands out is the presence of many "non-Western" countries in the top 10.
+Though data are not available over time (ie, we cannot make temporal comparisons), the landscape of investment 20 years prior would have been both (a) far smaller and (b) primarily composed of western countries.
+
 ### Massive increase in FDI
@@
 macro_names <- names(db)
+macro_names <- ifelse(grepl('X__', macro_names), 'Var', macro_names)
+
 for (j in 1:length(macro_names)){

### 2017-07-11 paper.tex
big update

@@
  \end{frontmatter}
+
+\subsection{Abstract}\label{abstract}
+
+Foreign direct investment (FDI) in Mozambique has increased rapidly in
+the last two decades. The growing interest in corporate social
+responsibility (CSR) - combined with a recent push for malaria
+elimination - suggest the need for a critical examination of the role of
+FDI in malaria control and elimination. Through a systematic review of
+the literature, we find that there has been a notable increase in
+research pertaining to FDI, CSR and malaria in recent years. Consensus
+among researchers is that this increase has not been accompanied by a
+substantial private sector investment in malaria control (with a few
+notable exceptions). This suggests a potential opportunity for scaling
+up privately driven malaria control as a means to both improve public
+health and increase return on investment. However, given the lack of
+coordination between public and private sectors, even when interests
+coincide, an over-reliance on foreign and private initiative for funding
+eradication is not without risks.
 
@@
 
+\subsubsection{Sources of FDI in
+Mozambique}\label{sources-of-fdi-in-mozambique}
+
+According to the IMF, 69 countries had foreign direct investment in
+Mozambique as of the end of 2015. The below chart shows the top 10
+countries ranked by amount invested in Mozambique.
+
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1} \end{center}
+
+What stands out is the presence of many ``non-Western'' countries in the
+top 10. Though data are not available over time (ie, we cannot make
+temporal comparisons), the landscape of investment 20 years prior would
+have been both (a) far smaller and (b) primarily composed of western
+countries.
+
 \subsubsection{Massive increase in FDI}\label{massive-increase-in-fdi}
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-11-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-12-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-12-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-14-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-15-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-17-1} \end{center}
 

### 2017-07-11 paper.Rmd
incorporating elisas fmi data

--- 2017-05-22 paper.Rmd
@@
 
+### Sources of FDI in Mozambique
+
+```{r}
+# Define function for reading fdi data from imf
+read_imf <- function(file){
+  srcs <- readxl::read_excel(file, skip = 4)
+  # Rename columns
+  srcs <- srcs[,1:3]
+  srcs <- srcs[2:nrow(srcs),]
+  names(srcs) <- c('from_country', 'inward', 'outward')
+  # Make numberic
+  srcs <- data.frame(srcs)
+  for (j in 2:3){
+    this_column <- as.numeric(as.character(gsub('.', '', srcs[,j], fixed = TRUE)))
+    srcs[,j] <- this_column
+  }
+  srcs <- srcs %>%
+    filter(from_country != 'World')
+
+  # Get amount by inward
+  x <- srcs %>%
+    arrange(desc(inward)) %>%
+    filter(!is.na(inward) & inward > 0) %>%
+    mutate(from_country = ifelse(from_country == 'China, P.R.: Mainland',
+                                 'China',
+                                 ifelse(from_country == 'United Arab Emirates',
+                                        'U.A.E.',
+                                        ifelse(from_country == 'United Kingdom',
+                                               'U.K.',
+                                               from_country)))) #%>%
+    # mutate(from_country = factor(from_country, levels = from_country))
+  return(x)
+}
+
+years <- 2009:2015
+results_list <- list()
+for (i in 1:length(years)){
+  this_year <- years[i]
+  this_data <- read_imf(paste0('data/imf_inward_direct_investment_positions/', this_year, '.xls')) %>%
+    mutate(year = this_year)
+  results_list[[i]] <- this_data
+}
+x <- bind_rows(results_list)
+```
+
+According to the IMF, `r x$from_country[x$year == 2015 & !is.na(x$inward)]` countries had foreign direct investment in Mozambique as of the end of 2015.
+The below chart shows the top 10 countries ranked by amount invested in Mozambique.
+
+```{r}
+last_year <- x %>%
+  filter(year == 2015)
+last_year <- last_year %>% 
+  arrange(desc(inward)) %>%
+  mutate(from_country = factor(from_country, levels = as.character(from_country)))
+last_year <- last_year[1:10,]
+ggplot(data = last_year,
+       aes(x = from_country,
+           y = inward)) +
+  geom_bar(stat = 'identity',
+           alpha = 0.6,
+           fill = 'darkblue') +
+  labs(x = 'Source of investment',
+       y = '2015 USD, Millions',
+       title = 'Sources of FDI in Mozambique',
+       subtitle = 'By investor country') +
+  theme_bw() +
+    theme(axis.text.x = element_text(angle = 90)) 
+```
+
+What stands out is the presence of many "non-Western" countries in the top 10.
+Though data are not available over time (ie, we cannot make temporal comparisons), the landscape of investment 20 years prior would have been both (a) far smaller and (b) primarily composed of western countries, especially given that Mozambique emerged so late from civil war (1994).
+
+If we examine the same IMF data over time, we can see that even in the short period for which data is available (2009 to 2015) there is a general upward trend, but that the trend is steepest in the last two years from China:
+
+```{r}
+cols <- colorRampPalette(brewer.pal(n = 9, name = 'Spectral'))(nrow(last_year))
+cols[3] <- 'black'
+ggplot(data = x %>%
+         filter(from_country %in% last_year$from_country),
+       aes(x = year,
+           y = inward,
+           group = from_country,
+           color = from_country)) +
+  geom_line() +
+  scale_color_manual(name = 'Country',
+                     values = cols) +
+labs(x = 'Source of investment',
+       y = '2015 USD, Millions',
+       title = 'Sources of FDI in Mozambique over time',
+       subtitle = 'By investor country') +
+  theme_bw() 
+```
+
 ### Massive increase in FDI

### 2017-07-11 paper.tex
incorporating elisas fmi data

--- 2017-05-22 paper.tex
@@
 
+\subsubsection{Sources of FDI in
+Mozambique}\label{sources-of-fdi-in-mozambique}
+
+According to the IMF, U.K., Zambia, China, Canada, Turkey, U.A.E.,
+Romania, India, Bulgaria, Lebanon, Argentina, Singapore, Netherlands,
+Spain, Norway, Angola, United States, China, P.R.: Macao, Denmark,
+Botswana, Serbia, Republic of, Guadeloupe, Swaziland, Bermuda, Panama,
+Japan, Switzerland, Ireland, Malaysia, Russian Federation, Nigeria,
+Uruguay, Mauritius, Italy, Burundi, Kuwait, Guinea-Bissau, Australia,
+Zimbabwe, Togo, China, P.R.: Hong Kong, Virgin Islands, British,
+Portugal, Iceland, Austria, Sweden, Malta, Isle of Man, Suriname,
+Tanzania, Brazil, Korea, Republic of, Bahamas, The, Saudi Arabia, Ghana,
+Seychelles, Reunion, Kenya, South Africa, France, Congo, Democratic
+Republic of, Germany, Uganda, Thailand, Slovak Republic, Estonia, Puerto
+Rico, Cyprus, Indonesia countries had foreign direct investment in
+Mozambique as of the end of 2015. The below chart shows the top 10
+countries ranked by amount invested in Mozambique.
+
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1} \end{center}
+
+What stands out is the presence of many ``non-Western'' countries in the
+top 10. Though data are not available over time (ie, we cannot make
+temporal comparisons), the landscape of investment 20 years prior would
+have been both (a) far smaller and (b) primarily composed of western
+countries, especially given that Mozambique emerged so late from civil
+war (1994).
+
+If we examine the same IMF data over time, we can see that even in the
+short period for which data is available (2009 to 2015) there is a
+general upward trend, but that the trend is steepest in the last two
+years from China:
+
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1} \end{center}
+
 \subsubsection{Massive increase in FDI}\label{massive-increase-in-fdi}
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-11-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-12-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-13-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-12-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-15-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-15-1} \end{center}
+\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-18-1} \end{center}
 

### 2017-12-13 paper.Rmd
major update, addressing elisas comments

@@
 title: Foreign direct investment, corporate social responsibility, and malaria control in Mozambique - trends, risks, and opportunities
-author:
-  - name: Joe Brew
-    email: joebrew@economicsofmalaria.com
-    affiliation: isglobal,cism,vu
-    footnote: Corresponding Author
-  - name: Elisa Sicuri
-    email: elisa.sicuri@isglobal.org 
-    affiliation: isglobal,icl,cism
-address:
-  - code: isglobal
-    address: Barcelona Institute for Global Health, c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain
-  - code: icl
-    address: Imperial College London, South Kensington Campus, London SW7 2AZ, U.K.
-  - code: cism
-    address: Centro de Investigação em Saúde de Manhiça, Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique
-  - code: vu
-    address: VU University Amsterdam, De Boelelaan 1105, 1081 HV Amsterdam, Netherlands
-abstract: |
-  Foreign direct investment (FDI) in Mozambique has increased rapidly in the last two decades.
-The growing interest in corporate social responsibility (CSR) - combined with a recent push for malaria elimination - suggest the need for a critical examination of the role of FDI in malaria control and elimination.
-Through a systematic review of the literature, we find that there has been a notable increase in research pertaining to FDI, CSR and malaria in recent years.
-Consensus among researchers is that this increase has not been accompanied by a substantial private sector investment in malaria control (with a few notable exceptions).
-This suggests a potential opportunity for scaling up privately driven malaria control as a means to both improve public health and increase return on investment.
-However, given the lack of coordination between public and private sectors, even when interests coincide, an over-reliance on foreign and private initiative for funding eradication is not without risks.
-
+date: ''
+output: 
+  pdf_document:
+    includes:
+      in_header: preamble.sty
+fig_width: 4
+fig_height: 2.6
 bibliography: library.bib
-output: rticles::elsevier_article
-header-includes:
-     - \usepackage{longtable}
 ---
@@
 
+# Authorship
+source('make_authors.R')
+
 # BASIC KNITR OPTIONS
@@
 
+\begin{center}
+\begin{large}
+
+```{r, results = 'asis', eval = TRUE}
+make_authors()
+```
+
+\end{large}
+\end{center}
+
+\vspace{5mm}
+
 ```{r, fig.height = 3, fig.align='center', eval = FALSE}
@@
 
-## Abstract
+\begin{center}
+\textbf{Abstract}  
+\end{center}
+
+\vspace{5mm}
+\begin{center}
+\begin{changemargin}{3cm}{3cm}
 
@@
 
+\end{changemargin}
+\end{center}
+
 ## Introduction
 
-Mozabique's recent economic growth has been facilitated by the government's open policy towards foreign investment, a plentiful supply of natural resources (particularly minerals and hydrocarbons) [@Rogers] and relatively inexpensive labor.
-Improvements in health have accompanied economic expansion, but Mozambique still lags behind in basic health outcomes - particularly those related to malaria - even by regional standards.
-Most economic activities have to deal
-with malaria: particularly agricultural-based and extraction activities cannot be carried out without consistent investment towards malaria management, both
-prevention and prompt treatment. This can be done as part of the “normal activities” of private sector or can go slightly beyond what is strictly needed for the
-activities and be seen as part of the corporate social responsibilityDespite  strong evidence showing that improved health yields substantial economic benefits [@World1999, @Bloom2008], there is little evidence regarding the extent to which "corporate social responsibility" (CSR) activities of foreign firms have targeted or impacted malaria.
-Understanding the relationship between FDI, CSR and malaria is essential to identifying which incentives and barriers to private sector involvement in malaria elimination exist, as well as identifying potential opportunities and synergies for scaling up private sector involvment in malaria elimination.
-
-That said, 
-The few malaria-related CSR projects in recent years has come in large, "mega-projects" whose well-publicized malaria abatement activities are profit-driven [@Mouzin2011] or whose primary aim is not profit-related [@Han].
-Though the former is often portrayed as a "win-win" for business and public health, the latter also offers tangible benefits for private industry, and should be understood as operating under the same conditions and with the same motivations.
-
-Private foreign investors showed a great optimism during the FMI meeting held in Maputo in 2014 (reference needed).
+Mozambique is a Southern-East African country with a surface area of nearly 800,000 m2 (more than 3 times that of the United Kingdom) and a population of nearly 30 million inhabitants.
+Though low-income, Mozambique experienced an upward trend in gross domestic product (GDP) per capita from approximately 200 USD in the late 90s to 623 in 2014, followed by a steep decrease in 2015 down to 382 in 2016 [@doingbusiness].
+
+Mozambique's economic growth up to 2014 was facilitated by the government's open policy towards foreign direct investment (FDI), a plentiful supply of natural resources [@Rogers], and relatively inexpensive labor.
+Foreign investment in Mozambique is largely geared towards the export market, and heavily concentrated.
+63% of exports come from aluminum, electricity, minerals, and gas, with each sector displaying high degrees of concentration (greater than 50% of the export share of the previous four industries are attributable to one company) [@Sutton2014].
+The total number of foreign enterprises operating in Mozambique is not ascertainable, but in both quantity and diversity of sources, FDI has increased dramatically in recent years [@Sutton2014].
+Mozambique was the Sub-Saharan African country with the greatest increase in FDI (defined by the World Bank as cross-border investment to establish a lasting interest) inflows from 2006 to 2014, registering a boom between 2010 and 2014 of about 5000 million USD [@Unctad2012].
+
+The impact of FDI on economic growth and other economic indicators has been largely assessed [@Alfaro_2006] [@Almfraji_2014] [@World1999].
+As FDI has been found to improve working conditions in low-income countries and, consequently, population ability to pay [@Bloom2008], they are likely to improve access to healthcare and, therefore, improve health [@Feenstra].
+However, there is some evidence that FDI may have adverse effects on health, through increasing the consumption of “bads” such as tobacco and alcohol and the level of harmful pollution [@Pazienza_2015] [@Shahbaz_2015].
+It has been recently assessed that FDIs are weakly associated with a marginal benefit in adults’ life expectancy in low and middle income countries (no impact has been found on children’s health), yet investments into the secondary sector (for being polluting) have been found potentially harmful [@Burns_2017].
+
+In Mozambique, improvements in health have accompanied economic expansion, but the country still lags behind in many basic health outcomes [@Williams_2015].
+For example, clusters exist in Southern Mozambique with HIV prevalence as high as 50\% [@Gonz_lez_2015].
+
+Malaria is a protozoan parasitic disease, transmitted to humans by mosquitoes.
+The _Plasmodium falciparum_ species, transmitted by the female _Anopheles_ mosquito, accounts for a large majority of deaths [@White2014].
+Estimates for the year 2015 quantify the burden of malaria in Mozambique in 8,300,000 cases and 15,000 associated deaths [@World2016].
+Despite the still high malaria burden, Mozambique is part of the "Elimination Eight (E8)" malaria initative, with the country-specific goal of malarai elimination by 2030.
+Elimination efforts have been particularly active in Southern Mozambique, in border areas contiguous with Swaziland and South Africa, where the goal is elimination by 2020 [@Moonasar_2016].
+
+Several economic activities are strictly linked to malaria.
+On the one hand, particularly agricultural-based and extraction activities are responsible for considerable increases in the malaria burden (REFERENCE NEED FROM ELISA), through land and environment modifications that favor the efficiency of mosquitos in spreading the malaria parasites [@Sheela_2017].
+On the other hand, the same activities, are barely carried out without consistent investment towards malaria management, both prevention and prompt treatment (REFERENCE NEEDED FROM ELISA): firms need to protect their workers against infections to maintain their productivity high.
+In this regards, there is little evidence on economic burden malaria places on the private sector in endemic countries: however, available reports point to profitability of founding malaria control [@Nonvignon_2016].
+Malaria control carried out by some multinational companies have been highly successful and strategies implemented were adopted as a model for the whole country [@Asante_2011].
+
+Private investments in malaria control can be done as part of the “normal actions” of private sector or can go slightly beyond what is strictly needed for having workers’ health to carry out activities and be seen as part of the corporate social responsibility (CSR).
+CSR is defined as those cases in which a company goes "beyond the legal requirements of the country in which they operate" in their work towards that country's long-term interests, in such a systematic manner that it becomes "business as usual" [@Ortega_2016].
+
+Motivations for engaging in CSR has been analyzed in the economic literature.
+Enterprises would have no incentives to engage in CSR according to Friedman [@Friedman] as firms are organized to maximize their profits only.
+The opposite view is that CSR is actually a strategy for maximizing profits.
+According to this vision, CSR helps enterprises to make their products known and, therefore, foment consumers’ demand [@Porter2002].
+CSR also helps building a positive reputation of enterprises among institutions and communities where they operate: this is particularly useful for foreign firms that have no previous contacts with the host country/communities [@Zhu_2013].
+
+A study has recently looked at the propensity and at the magnitude of CSR to the host communities in the US by comparing foreign and local enterprises [@NBERw17282].
+They found that foreign-owned enterprises are less likely to give, but that when they do give, it is substantially more in magnitude than domestic firms, everything else equal.
+Their null hypothesis was that foreign firms give as much as local ones based on the assumptions that firms decide on CSR decisions from a strict profit maximization perspective and that, therefore, foreign or local ownership does not affect such a choice.
+One alternative hypothesis was that foreign firms are less likely to give or give less than local because their products are mainly not for the local markets and because foreign managers have higher propensity to give not locally but either internationally or to their country of origin.
+A second alternative hypothesis is that foreign firms give more than the local ones.
+This may be supported by the fact that foreign companies have different culture from local, more in favor of giving or because foreign firms use CSR to overcome political and cultural barriers in the host country that would not be able to mitigate differently.
+
+The relationship between FDI and foreign aid may also be an interesting determinant of CSR: if aid from a particular country increases FDI from the same country to the same host communities, incentives to engage in CSR from foreign private companies may disappear [@Selaya_2012].
+In other words, why should foreign companies donate to the host country when the government of the country of origin is already giving?
+
+Private foreign investors showed a great optimism during the International Monetary Fund meeting held in Maputo, the Mozambican city capital, in 2014 [@Pfeiffer_2017].
 Beginning in late 2015, the flood of FDI into Mozambique (and many other subsaharan African nations) slowed to a trickle.
 The impact of this slowdown on CSR is not yet known, but it can reasonably be assumed that it will mean a reduction in CSR activities (albeit with lag).
-Almost in parallel to this, Mozambique has had serious financial problems which have
-resulted in default (reference needed http://www.economist.com/news/middleeast-and-africa/21715030-mozambique-fails-pay-its-debtsmozambiques-default).
+Almost in parallel to this, Mozambique has had serious financial problems which have resulted in default [@Economist2017].
 Given the rapidly changing economic and epidemiologic context in Mozambique, a comprehensive and current understanding of both (a) the landscape of FDI and CSR in Mozambique and (b) public health issues (namely, malaria control) which are directly affected those investments is sorely needed.
@@
 It is by no means comprehensive, but offers a consolidated starting point for understanding both the place of FDI and CSR in the literature, as well as where the interests and incentives of the public and private sectors converge and differ in regards to malaria control.
+
+I AM HERE.
 
@@
 # Define function for reading fdi data from imf
+if('imf.RData' %in% dir()){
+  load('imf.RData')
+} else {
   read_imf <- function(file){
@@
   x <- bind_rows(results_list)
-```
-
-According to the IMF, `r x$from_country[x$year == 2015 & !is.na(x$inward)]` countries had foreign direct investment in Mozambique as of the end of 2015.
+  save(x, file = 'imf.RData')
+}
+```
+
+According to the IMF, `r length(x$from_country[x$year == 2015 & !is.na(x$inward)])` countries had foreign direct investment in Mozambique as of the end of 2015\footnote{`r paste0(x$from_country[x$year == 2015 & !is.na(x$inward)], collapse = ', ')`}.
 The below chart shows the top 10 countries ranked by amount invested in Mozambique.
@@
   geom_line() +
-  scale_color_manual(name = 'Country',
+  scale_color_manual(name = '',
                      values = cols) +
@@
        subtitle = 'By investor country') +
-  theme_bw() 
+  theme_bw() +
+  theme(legend.position = 'bottom') +
+  guides(col = guide_legend(nrow = 4))
 ```
@@
 
+FIND A PLACE FOR THIS PARAGRAPH: 
+The few malaria-related CSR projects in recent years has come in large, "mega-projects" whose well-publicized malaria abatement activities are profit-driven [@Mouzin2011] or whose primary aim is not profit-related [@Han].
+Though the former is often portrayed as a "win-win" for business and public health, the latter also offers tangible benefits for private industry, and should be understood as operating under the same conditions and with the same motivations.
+
 ## Scaling up malaria control through CSR: opportunity and risk

### 2017-12-13 make_authors.R

@@ -1,0 +1,91 @@
+#' Make authors
+#'
+#' Cat the authors in latex-ready format
+#' @return Latex text
+#' @param authors A dataframe with columns: name, email, affiliation, footnote
+#' @param institutions A dataframe with columns: code, institution, address
+#' @param cat_it Whether to cat (as opposed to return)
+#' @param include_address Whether to include the address
+#' @param include_country Whether to include the country; only relevant if include_address is FALSE
+#' @param seperator What to put between author names: usually a comma or line break
+#' @export
+
+make_authors <- function(
+
+  authors = data.frame(name = c('Joe Brew',
+                                'Celine Aerts',
+                                'Elisa Sicuri'),
+                       email = c('joe@databrew.cc',
+                                 'celine.aerts@isglobal.org',
+                                 'elisa.sicuri@isglobal.org'),
+                       affilitation = c("isglobal,cism,vu",
+                                        'isglobal',
+                                        "isglobal,cism,icl"),
+                       footnote = c("Corresponding Author",
+                                    rep('', 2)),
+                       stringsAsFactors = FALSE),
+
+  institutions = data.frame(code = c('isglobal',
+                                     'icl',
+                                     'cism',
+                                     'vu',
+                                     'uva',
+                                     'ma'),
+                            institution = c('Barcelona Institute for Global Health',
+                                            'Imperial College London',
+                                            'Centro de Investigação em Saúde de Manhiça',
+                                            'VU University Amsterdam',
+                                            'University of Amsterdam',
+                                            'Maragra Açucar SA, Subsidiary of Illovo Sugar Ltd'),
+                            country = c('Spain', 'UK', 'Mozambique', 'Netherlands', 'Netherlands', 'Mozambique'),
+                            address = c('c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain',
+                                        'South Kensington Campus, London SW7 2AZ, U.K.',
+                                        'Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique',
+                                        'De Boelelaan 1105, 1081 HV Amsterdam, Netherlands',
+                                        'REC E, Roetersstraat 11, Amsterdam, Netherlands',
+                                        'CP 2789, Maputo, Mozambique'),
+                            stringsAsFactors = FALSE),
+  cat_it = TRUE,
+  include_address = TRUE,
+  include_country = FALSE,
[…]

### 2017-12-27 paper.Rmd
updated

@@
   pdf_document:
+    keep_tex: true
     includes:
@@
 # BASIC KNITR OPTIONS
-opts_chunk$set(comment = NA, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, fig.align = 'center', fig.height = 3, fig.width = 4)
+opts_chunk$set(comment = NA, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, fig.align = 'center', fig.height = 4, fig.width = 5, fig.path = "figures/")
 options(xtable.comment = FALSE)
@@
 Several economic activities are strictly linked to malaria.
-On the one hand, particularly agricultural-based and extraction activities are responsible for considerable increases in the malaria burden (REFERENCE NEED FROM ELISA), through land and environment modifications that favor the efficiency of mosquitos in spreading the malaria parasites [@Sheela_2017].
-On the other hand, the same activities, are barely carried out without consistent investment towards malaria management, both prevention and prompt treatment (REFERENCE NEEDED FROM ELISA): firms need to protect their workers against infections to maintain their productivity high.
+On the one hand, particularly agricultural-based and extraction activities are responsible for considerable increases in the malaria burden, through land and environment modifications that favor the efficiency of mosquitos in spreading the malaria parasites [@Sheela_2017].
+On the other hand, the same activities, are barely carried out without consistent investment towards malaria management, both prevention and prompt treatment: firms need to protect their workers against infections to maintain their productivity high.
 In this regards, there is little evidence on economic burden malaria places on the private sector in endemic countries: however, available reports point to profitability of founding malaria control [@Nonvignon_2016].
@@
 It is by no means comprehensive, but offers a consolidated starting point for understanding both the place of FDI and CSR in the literature, as well as where the interests and incentives of the public and private sectors converge and differ in regards to malaria control.
-
-I AM HERE.
 
@@
 
-We examined and analyzed trends from multiple datasets related to FDI, CSR and malaria in Mozambique.
-These included the "Doing Business" data pertaining to the measurement of regulatory efficiency, the GADM, the Deutsche Bundesbank Data Repository for foreign exchange rate history, the Knoema World Data Atlas for macroeconomic trends, the World Bank data portal for information containing to net foreign inflows and FDI, USAID Demographic and Health Survey data on sociodemographics and health-related practices, the Instituto Nacional de Estatistica for granular data pertaining to Mozambique's population and economic activities, and Institute for Health Metrics and Evaluation data pertaining to disease trends over time.
+We examined trends from multiple datasets related to FDI, CSR and malaria in Mozambique.
+These included the "Doing Business" data pertaining to the measurement of regulatory efficiency [@doingbusiness], the GADM [@gadm], the Deutsche Bundesbank Data Repository for foreign exchange rate history [@deutsche], the Knoema World Data Atlas for macroeconomic trends [@knoema], the World Bank data portal for information containing to net foreign inflows and FDI, USAID Demographic and Health Survey data on sociodemographics and health-related practices [@wbd], the Instituto Nacional de Estatistica for granular data pertaining to Mozambique's population and economic activities [@ine], the Institute for Health Metrics and Evaluation data pertaining to disease trends over time [@ihme], the International Monetary Fund's open datasets pertaining to FDI [@imf], and the United Nations Conference on Trade and Development's data on FDI [@Unctad2012].
 
@@
 ## Quantitiative datasets pertaining to FDI and CSR in Mozambique
-
-### Sources of FDI in Mozambique
 
@@
 What stands out is the presence of many "non-Western" countries in the top 10.
-Though data are not available over time (ie, we cannot make temporal comparisons), the landscape of investment 20 years prior would have been both (a) far smaller and (b) primarily composed of western countries, especially given that Mozambique emerged so late from civil war (1994).
-
-If we examine the same IMF data over time, we can see that even in the short period for which data is available (2009 to 2015) there is a general upward trend, but that the trend is steepest in the last two years from China:
-
-```{r}
+Though reliable data only go as far back as 2009, the landscape of investment 20 years prior would have been both (a) far smaller and (b) primarily composed of western countries, especially given that Mozambique emerged so late from civil war (1994).
+Though as of 2015, China did not figure in the top 10 FDI countries in Mozambique, its upward trend is notable, with approximately 1.5 million USD invested in 2010 to nearly 100 million by 2015 (a 60-fold increase).
+
+```{r, eval = FALSE}
+china <- x %>%
+  filter(from_country == 'China')
+
 cols <- colorRampPalette(brewer.pal(n = 9, name = 'Spectral'))(nrow(last_year))
@@
 #####
-ggplot(data = temp, aes(x = year, y = value)) +
+g1 <- ggplot(data = temp, aes(x = year, y = value)) +
     scale_y_log10() +
@@
 
-```{r}
+```{r, fig.width = 7}
 # Load up forex data
@@
 
-ggplot(data = forex,
+g2 <- ggplot(data = forex,
        aes(x = Date, y = Value)) +
@@
   ylim(0, 40)
+multiplot(g1, g2, cols = 2)
 ```
@@
 # Clean up
-temp <- gather(db, key, value, cost_export_container:register_property)
-temp$value <- as.numeric(temp$value)
+tempo <- gather(db, key, value, cost_export_container:register_property)
+tempo$value <- as.numeric(tempo$value)
 
 # Clean up names/labels
-temp$key <- factor(temp$key,
+tempo$key <- factor(tempo$key,
                    labels = c('Cost to export container (deflated USD)',
@@
 # PLOT
-g1 <- 
-  ggplot(data = temp[temp$key %in% unique(sort(temp$key))[1:2],], 
+g3 <- 
+  ggplot(data = tempo[tempo$key %in% unique(sort(tempo$key))[1:2],], 
          aes(x = year, y = value)) +
@@
     theme_bw() +
-  theme(strip.text = element_text(size=5))
-g2 <- 
-  ggplot(data = temp[temp$key %in% unique(sort(temp$key))[3:4],], 
+  theme(strip.text = element_text(size=5))+
+  labs(x = 'Year', y = 'USD')
+g4 <- 
+  ggplot(data = tempo[tempo$key %in% unique(sort(tempo$key))[3:4],], 
          aes(x = year, y = value)) +
@@
     theme_bw() +
-   theme(strip.text = element_text(size=5))
-
-multiplot(g1, g2)  
+   theme(strip.text = element_text(size=5)) +
+  labs(x = 'Year', y = '%')
+
+multiplot(g3, g4)  
 ```
@@
 
-### CSR in Mozambique
-
-Corporate social responsibility in Mozambique is new in nomenclature, but activities which could be classified as CSR have existed for decades.
-The number of firms actively engaging in CSR cannot be ascertained (given the large and ever-changing number of small businesses), but virtually all of the largest firms, both foreign and domestic, have a CSR component.
-
-Firms with CSR activities are often large and foreign.
-Among the largest "key players" in Mozambican CSR are Coca-Cola, British Petroleum and Colgate-Palmolive.
-State conglomerates, such as Águas de Moçambique and Electricidade de Moçambique, also participate in CSR activities [@Compact2007].
-
-### Malaria-related CSR and other corporate activity
-
-Most CSR-funded activities are focused on education, community development, women's rights and entrepreneurship.
-According to a UN poll, none of the country's largest firms invest directly in malaria-related CSR activities [@Compact2007].
-This may be due to the perceived costs of malaria control, lack of perceived PR benefits, the government and NGO's predominant role in the area, as well as the "opt-in" nature and privacy/legal issues generally related to engaging in health-related campaigns.
-
-That said, a majority of the firms interviewed by the UN indicated that one of the principal reasons for investment in CSR is to complement government efforts.
-To the extent that malaria accounts for more of the loss in disability-adjusted life years in Mozambique than comparable countries [@ihme], malaria control's lack of representation among core CSR activities is a notable absence.
-
+FDI is associated with CSR.
 Whether under the guise of CSR or not, it is noteworthy that the private sector currently does play a role in malaria control activities.
@@
 
-One notable exception is the Nando's-lead Goodbye Malaria Trust, an umbrella organization which inclues a development impact bond, the Goodbye Malaria initiative, and partnerships with several other foreign firms operating in Mozambique (more details here).
+Corporate social responsibility in Mozambique is new in nomenclature, but activities which could be classified as CSR have existed for decades.
+The number of firms actively engaging in CSR cannot be ascertained (given the large and ever-changing number of small businesses), but virtually all of the largest firms, both foreign and domestic, have a CSR component.
+
+Firms with CSR activities are often large and foreign.
+Among the largest "key players" in Mozambican CSR are Coca-Cola, British Petroleum and Colgate-Palmolive.
+State conglomerates, such as Águas de Moçambique and Electricidade de Moçambique, also participate in CSR activities [@Compact2007].
+Large communications firms, such as Vodacom and MCEL, and banks (BIM and BCI) have a CSR component, but generally focus on culture, environment and sport, avoiding activities which could either complement or conflict with the health sector [@Kaufmann_2015].
+
+Among major foreign companies, those operating in extraction - particularly mining - appear to be the most involved with malaria.
+BHP Billiton has been active in vector control support to the governments of Mozambique, Swaziland, and South Africa, particularly in the early 2000s [@bhp].
+Kropfmuhl AG and Vale, also mining companies, have carried out CSR related to both community development and malaria control in Northern Mozambique [@Kaufmann_2015].
+
+Most CSR-funded activities are focused on education, community development, women's rights and entrepreneurship.
+According to a UN poll, none of the country's largest firms invest directly in malaria-related CSR activities [@Compact2007].
+That said, a majority of the firms interviewed by the UN indicated that one of the principal reasons for investment in CSR is to complement government efforts.
+To the extent that malaria accounts for more of the loss in disability-adjusted life years in Mozambique than comparable countries [@ihme], malaria control's lack of representation among core CSR activities is a notable absence.
+One important exception is the Nando's-lead Goodbye Malaria Trust, an umbrella organization which inclues a development impact bond, the Goodbye Malaria initiative, and partnerships with several other foreign firms operating in Mozambique.
 
@@
 
-FIND A PLACE FOR THIS PARAGRAPH: 
+The intensity of CSR activity seems to closely track FDI, but generally avoids malaria control. 
 The few malaria-related CSR projects in recent years has come in large, "mega-projects" whose well-publicized malaria abatement activities are profit-driven [@Mouzin2011] or whose primary aim is not profit-related [@Han].
 Though the former is often portrayed as a "win-win" for business and public health, the latter also offers tangible benefits for private industry, and should be understood as operating under the same conditions and with the same motivations.
+To the extent that most CSR activity has avoided the health sector (including malaria) in favor of cultural, environmental, education and more general "development" activities, it is reasonable to assume that this may be due to the perceived costs of malaria control, lack of perceived PR benefits, the government and NGO's predominant role in the area, as well as the "opt-in" nature and privacy/legal issues generally related to engaging in health-related campaigns.
 
@@
 
-## References
+# References

### 2017-12-27 paper.tex
updated

--- 2017-05-22 paper.tex
@@
-\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
-%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
-\usepackage[hyphens]{url}
-\usepackage{lineno} % add
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-\bibliographystyle{elsarticle-harv}
-\biboptions{sort&compress} % For natbib
-\usepackage{graphicx}
-\usepackage{booktabs} % book-quality tables
-%% Redefines the elsarticle footer
-%\makeatletter
-%\def\ps@pprintTitle{%
-% \let\@oddhead\@empty
-% \let\@evenhead\@empty
-% \def\@oddfoot{\it \hfill\today}%
-% \let\@evenfoot\@oddfoot}
-%\makeatother
-
-% A modified page layout
-\textwidth 6.75in
-\oddsidemargin -0.15in
-\evensidemargin -0.15in
-\textheight 9in
-\topmargin -0.5in
-%%%%%%%%%%%%%%%% end my additions to header
-
-\usepackage[T1]{fontenc}
+\documentclass[]{article}
 \usepackage{lmodern}
@@
 \usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
 % use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \usepackage{fontspec}
-  \ifxetex
-    \usepackage{xltxtra,xunicode}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
 % use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
             pdftitle={Foreign direct investment, corporate social responsibility, and malaria control in Mozambique - trends, risks, and opportunities},
-            colorlinks=true,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
+            pdfborder={0 0 0},
+            breaklinks=true}
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 \setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+  \title{Foreign direct investment, corporate social responsibility, and malaria
+control in Mozambique - trends, risks, and opportunities}
+  \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+  \author{}
+  \preauthor{}\postauthor{}
+  \date{}
+  \predate{}\postdate{}
+
 \usepackage{longtable}
-% Pandoc toggle for numbering sections (defaults to be off)
-\setcounter{secnumdepth}{0}
-% Pandoc header
-\usepackage{longtable}
+\usepackage{multicol}
+\usepackage{hyperref}
+\usepackage{geometry}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{fancyhdr} % Headers and footers
+\pagestyle{fancy} % All pages have headers and footers
+\fancyhead{} % Blank out the default header
+\fancyfoot{} % Blank out the default footer
+\fancyhead[C]{joebrew@economicsofmalaria.com $\bullet$ ROUGH DRAFT $\bullet$ FDI in Mozambique}
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
 
-\usepackage[nomarkers]{endfloat}
+\newcommand{\footremember}[2]{%
+    \footnote{#2}
+    \newcounter{#1}
+    \setcounter{#1}{\value{footnote}}%
+}
+\newcommand{\footrecall}[1]{%
+    \footnotemark[\value{#1}]%
+}
+
+\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
+\let\endchangemargin=\endlist
+
+\widowpenalties 1 150
+
+\makeatletter
+\renewcommand\footnotesize{%
+   \@setfontsize\footnotesize\@ixpt{11}%
+   \abovedisplayskip 8\p@ \@plus2\p@ \@minus4\p@
+   \abovedisplayshortskip \z@ \@plus\p@
+   \belowdisplayshortskip 4\p@ \@plus2\p@ \@minus2\p@
+   \def\@listi{\leftmargin\leftmargini
+               \topsep 4\p@ \@plus2\p@ \@minus2\p@
+               \parsep 2\p@ \@plus\p@ \@minus\p@
+               \itemsep \parsep}%
+   \belowdisplayskip \abovedisplayskip
+}
+\makeatother
 
 \begin{document}
-\begin{frontmatter}
+\maketitle
 
-  \title{Foreign direct investment, corporate social responsibility, and malaria
-control in Mozambique - trends, risks, and opportunities}
-    \author[isglobal,cism,vu]{Joe Brew\corref{c1}}
-   \ead{joebrew@economicsofmalaria.com} 
-   \cortext[c1]{Corresponding Author}
-    \author[isglobal,icl,cism]{Elisa Sicuri}
-   \ead{elisa.sicuri@isglobal.org}
+\begin{center}
+\begin{large}
 
-      \address[isglobal]{Barcelona Institute for Global Health, c/ Rosselló, 132, 5è 2a. 08036,
-Barcelona, Spain}
-    \address[icl]{Imperial College London, South Kensington Campus, London SW7 2AZ, U.K.}
-    \address[cism]{Centro de Investigação em Saúde de Manhiça, Vila da Manhiça, Bairro
-Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique}
-    \address[vu]{VU University Amsterdam, De Boelelaan 1105, 1081 HV Amsterdam,
-Netherlands}
+Joe Brew\footremember{isglobal}{Barcelona Institute for Global Health: c/ Rosselló, 132, 5è 2a.
+08036, Barcelona, Spain}\footremember{cism}{Centro de Investigação em Saúde de Manhiça: Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique}\footremember{vu}{VU University Amsterdam: De Boelelaan 1105, 1081 HV Amsterdam, Netherlands}
+Celine Aerts\footrecall{isglobal}
+Elisa Sicuri\footrecall{isglobal}\footrecall{cism}\footremember{icl}{Imperial College London: South Kensington Campus, London SW7 2AZ, U.K.}
 
-  \begin{abstract}
-  Foreign direct investment (FDI) in Mozambique has increased rapidly in
-  the last two decades. The growing interest in corporate social
-  responsibility (CSR) - combined with a recent push for malaria
-  elimination - suggest the need for a critical examination of the role of
-  FDI in malaria control and elimination. Through a systematic review of
-  the literature, we find that there has been a notable increase in
-  research pertaining to FDI, CSR and malaria in recent years. Consensus
-  among researchers is that this increase has not been accompanied by a
-  substantial private sector investment in malaria control (with a few
-  notable exceptions). This suggests a potential opportunity for scaling
-  up privately driven malaria control as a means to both improve public
-  health and increase return on investment. However, given the lack of
-  coordination between public and private sectors, even when interests
-  coincide, an over-reliance on foreign and private initiative for funding
-  eradication is not without risks.
-  \end{abstract}
+\end{large}
+\end{center}
 
- \end{frontmatter}
+\vspace{5mm}
+
+\begin{center}
+\textbf{Abstract}  
+\end{center}
+
+\vspace{5mm}
+
+\begin{center}
+\begin{changemargin}{3cm}{3cm}
+
+Foreign direct investment (FDI) in Mozambique has increased rapidly in the last two decades.
+The growing interest in corporate social responsibility (CSR) - combined with a recent push for malaria elimination - suggest the need for a critical examination of the role of FDI in malaria control and elimination.
+Through a systematic review of the literature, we find that there has been a notable increase in research pertaining to FDI, CSR and malaria in recent years.
+Consensus among researchers is that this increase has not been accompanied by a substantial private sector investment in malaria control (with a few notable exceptions).
+This suggests a potential opportunity for scaling up privately driven malaria control as a means to both improve public health and increase return on investment.
+However, given the lack of coordination between public and private sectors, even when interests coincide, an over-reliance on foreign and private initiative for funding eradication is not without risks.
+
+\end{changemargin}
+\end{center}
 
@@
 
-Mozabique's recent economic growth has been facilitated by the
-government's open policy towards foreign investment, a plentiful supply
-of natural resources (particularly minerals and hydrocarbons) (Rogers
-2014) and relatively inexpensive labor. Improvements in health have
-accompanied economic expansion, but Mozambique still lags behind in
-basic health outcomes - particularly those related to malaria - even by
-regional standards. Most economic activities have to deal with malaria:
-particularly agricultural-based and extraction activities cannot be
-carried out without consistent investment towards malaria management,
-both prevention and prompt treatment. This can be done as part of the
-``normal activities'' of private sector or can go slightly beyond what
-is strictly needed for the activities and be seen as part of the
-corporate social responsibilityDespite strong evidence showing that
-improved health yields substantial economic benefits (Brundtland 1999,
-Bloom and Canning (2008)), there is little evidence regarding the extent
-to which ``corporate social responsibility'' (CSR) activities of foreign
-firms have targeted or impacted malaria. Understanding the relationship
-between FDI, CSR and malaria is essential to identifying which
-incentives and barriers to private sector involvement in malaria
-elimination exist, as well as identifying potential opportunities and
-synergies for scaling up private sector involvment in malaria
-elimination.
+Mozambique is a Southern-East African country with a surface area of
+nearly 800,000 m2 (more than 3 times that of the United Kingdom) and a
+population of nearly 30 million inhabitants. Though low-income,
+Mozambique experienced an upward trend in gross domestic product (GDP)
+per capita from approximately 200 USD in the late 90s to 623 in 2014,
+followed by a steep decrease in 2015 down to 382 in 2016 (WB 2016).
 
-That said, The few malaria-related CSR projects in recent years has come
-in large, ``mega-projects'' whose well-publicized malaria abatement
-activities are profit-driven (Mouzin and al. 2011) or whose primary aim
-is not profit-related (Han 2015). Though the former is often portrayed
-as a ``win-win'' for business and public health, the latter also offers
-tangible benefits for private industry, and should be understood as
-operating under the same conditions and with the same motivations.
+Mozambique's economic growth up to 2014 was facilitated by the
+government's open policy towards foreign direct investment (FDI), a
+plentiful supply of natural resources (Rogers 2014), and relatively
+inexpensive labor. Foreign investment in Mozambique is largely geared
+towards the export market, and heavily concentrated. 63\% of exports
+come from aluminum, electricity, minerals, and gas, with each sector
+displaying high degrees of concentration (greater than 50\% of the
+export share of the previous four industries are attributable to one
+company) (Sutton 2014). The total number of foreign enterprises
+operating in Mozambique is not ascertainable, but in both quantity and
+diversity of sources, FDI has increased dramatically in recent years
+(Sutton 2014). Mozambique was the Sub-Saharan African country with the
+greatest increase in FDI (defined by the World Bank as cross-border
+investment to establish a lasting interest) inflows from 2006 to 2014,
+registering a boom between 2010 and 2014 of about 5000 million USD
+(UNCTAD 2012).
 
-Private foreign investors showed a great optimism during the FMI meeting
-held in Maputo in 2014 (reference needed). Beginning in late 2015, the
+The impact of FDI on economic growth and other economic indicators has
+been largely assessed (Alfaro et al. 2006) (Almfraji and Almsafir 2014)
+(Brundtland 1999). As FDI has been found to improve working conditions
+in low-income countries and, consequently, population ability to pay
+(Bloom and Canning 2008), they are likely to improve access to
+healthcare and, therefore, improve health (Feenstra and Hanson 1997).
+However, there is some evidence that FDI may have adverse effects on
+health, through increasing the consumption of ``bads'' such as tobacco
+and alcohol and the level of harmful pollution (Pazienza 2015) (Shahbaz
+et al. 2015). It has been recently assessed that FDIs are weakly
+associated with a marginal benefit in adults' life expectancy in low and
+middle income countries (no impact has been found on children's health),
+yet investments into the secondary sector (for being polluting) have
+been found potentially harmful (Burns et al. 2017).
+
+In Mozambique, improvements in health have accompanied economic
+expansion, but the country still lags behind in many basic health
+outcomes (Williams et al. 2015). For example, clusters exist in Southern
+Mozambique with HIV prevalence as high as 50\% (González et al. 2015).
+
+Malaria is a protozoan parasitic disease, transmitted to humans by
+mosquitoes. The \emph{Plasmodium falciparum} species, transmitted by the
+female \emph{Anopheles} mosquito, accounts for a large majority of
+deaths (N. J. White et al. 2014). Estimates for the year 2015 quantify
+the burden of malaria in Mozambique in 8,300,000 cases and 15,000
+associated deaths (WHO 2016). Despite the still high malaria burden,
+Mozambique is part of the ``Elimination Eight (E8)'' malaria initative,
+with the country-specific goal of malarai elimination by 2030.
+Elimination efforts have been particularly active in Southern
+Mozambique, in border areas contiguous with Swaziland and South Africa,
+where the goal is elimination by 2020 (Moonasar et al. 2016).
+
+Several economic activities are strictly linked to malaria. On the one
+hand, particularly agricultural-based and extraction activities are
+responsible for considerable increases in the malaria burden, through
+land and environment modifications that favor the efficiency of
+mosquitos in spreading the malaria parasites (Sheela et al. 2017). On
+the other hand, the same activities, are barely carried out without
+consistent investment towards malaria management, both prevention and
+prompt treatment: firms need to protect their workers against infections
+to maintain their productivity high. In this regards, there is little
+evidence on economic burden malaria places on the private sector in
+endemic countries: however, available reports point to profitability of
+founding malaria control (Nonvignon et al. 2016). Malaria control
+carried out by some multinational companies have been highly successful
+and strategies implemented were adopted as a model for the whole country
+(Asante et al. 2011).
+
+Private investments in malaria control can be done as part of the
+``normal actions'' of private sector or can go slightly beyond what is
+strictly needed for having workers' health to carry out activities and
+be seen as part of the corporate social responsibility (CSR). CSR is
+defined as those cases in which a company goes ``beyond the legal
+requirements of the country in which they operate'' in their work
+towards that country's long-term interests, in such a systematic manner
+that it becomes ``business as usual'' (Ortega et al. 2016).
+
+Motivations for engaging in CSR has been analyzed in the economic
+literature. Enterprises would have no incentives to engage in CSR
+according to Friedman (Friedman 1970) as firms are organized to maximize
+their profits only. The opposite view is that CSR is actually a strategy
+for maximizing profits. According to this vision, CSR helps enterprises
+to make their products known and, therefore, foment consumers' demand
+(Porter and Kramer 2002). CSR also helps building a positive reputation
+of enterprises among institutions and communities where they operate:
+this is particularly useful for foreign firms that have no previous
+contacts with the host country/communities (Zhu, Sun, and Leung 2013).
+
+A study has recently looked at the propensity and at the magnitude of
+CSR to the host communities in the US by comparing foreign and local
+enterprises (Blonigen and O'Fallon 2011). They found that foreign-owned
+enterprises are less likely to give, but that when they do give, it is
+substantially more in magnitude than domestic firms, everything else
+equal. Their null hypothesis was that foreign firms give as much as
+local ones based on the assumptions that firms decide on CSR decisions
+from a strict profit maximization perspective and that, therefore,
+foreign or local ownership does not affect such a choice. One
+alternative hypothesis was that foreign firms are less likely to give or
+give less than local because their products are mainly not for the local
+markets and because foreign managers have higher propensity to give not
+locally but either internationally or to their country of origin. A
+second alternative hypothesis is that foreign firms give more than the
+local ones. This may be supported by the fact that foreign companies
+have different culture from local, more in favor of giving or because
+foreign firms use CSR to overcome political and cultural barriers in the
+host country that would not be able to mitigate differently.
+
+The relationship between FDI and foreign aid may also be an interesting
+determinant of CSR: if aid from a particular country increases FDI from
+the same country to the same host communities, incentives to engage in
+CSR from foreign private companies may disappear (Selaya and Sunesen
+2012). In other words, why should foreign companies donate to the host
+country when the government of the country of origin is already giving?
+
+Private foreign investors showed a great optimism during the
+International Monetary Fund meeting held in Maputo, the Mozambican city
+capital, in 2014 (Pfeiffer et al. 2017). Beginning in late 2015, the
 flood of FDI into Mozambique (and many other subsaharan African nations)
@@
 has had serious financial problems which have resulted in default
-(reference needed
-http://www.economist.com/news/middleeast-and-africa/21715030-mozambique-fails-pay-its-debtsmozambiques-default).
-Given the rapidly changing economic and epidemiologic context in
-Mozambique, a comprehensive and current understanding of both (a) the
-landscape of FDI and CSR in Mozambique and (b) public health issues
-(namely, malaria control) which are directly affected those investments
-is sorely needed. Such an understanding may foster greater private
-interest in for-profit investment in public health measures. Likewise,
-it can facilitate better public sector understanding of potential
-industry partners and stakeholders (a prerequisite to greater
+(Economist 2017). Given the rapidly changing economic and epidemiologic
+context in Mozambique, a comprehensive and current understanding of both
+(a) the landscape of FDI and CSR in Mozambique and (b) public health
+issues (namely, malaria control) which are directly affected those
+investments is sorely needed. Such an understanding may foster greater
+private interest in for-profit investment in public health measures.
+Likewise, it can facilitate better public sector understanding of
+potential industry partners and stakeholders (a prerequisite to greater
 public-private collaborations), and guide the public sector away from an
@@
 
-We examined and analyzed trends from multiple datasets related to FDI,
-CSR and malaria in Mozambique. These included the ``Doing Business''
-data pertaining to the measurement of regulatory efficiency, the GADM,
-the Deutsche Bundesbank Data Repository for foreign exchange rate
-history, the Knoema World Data Atlas for macroeconomic trends, the World
-Bank data portal for information containing to net foreign inflows and
-FDI, USAID Demographic and Health Survey data on sociodemographics and
-health-related practices, the Instituto Nacional de Estatistica for
-granular data pertaining to Mozambique's population and economic
-activities, and Institute for Health Metrics and Evaluation data
-pertaining to disease trends over time.
+We examined trends from multiple datasets related to FDI, CSR and
+malaria in Mozambique. These included the ``Doing Business'' data
+pertaining to the measurement of regulatory efficiency (WB 2016), the
+GADM (Gadm 2009), the Deutsche Bundesbank Data Repository for foreign
+exchange rate history (Bundesbank 2015), the Knoema World Data Atlas for
+macroeconomic trends (Knoema 2015), the World Bank data portal for
+information containing to net foreign inflows and FDI, USAID Demographic
+and Health Survey data on sociodemographics and health-related practices
+(WB 2017), the Instituto Nacional de Estatistica for granular data
+pertaining to Mozambique's population and economic activities (INE
+2017), the Institute for Health Metrics and Evaluation data pertaining
+to disease trends over time (IHME 2015), the International Monetary
+Fund's open datasets pertaining to FDI (IMF 2003), and the United
+Nations Conference on Trade and Development's data on FDI (UNCTAD 2012).
 
@@
 
+According to the IMF, 69 countries had foreign direct investment in
+Mozambique as of the end of
+2015\footnote{U.A.E., United States, South Africa, Mauritius, Italy, Australia, Portugal, U.K., India, Netherlands, Japan, Switzerland, Malaysia, France, Germany, Virgin Islands, British, Tanzania, Brazil, Bahamas, The, China, Spain, Norway, Botswana, Kuwait, Guinea-Bissau, Zimbabwe, Togo, China, P.R.: Hong Kong, Puerto Rico, Austria, Sweden, Malta, Isle of Man, Korea, Republic of, Lebanon, Singapore, Angola, China, P.R.: Macao, Denmark, Swaziland, Kenya, Estonia, Suriname, Zambia, Canada, Thailand, Ghana, Turkey, Seychelles, Argentina, Slovak Republic, Serbia, Republic of, Guadeloupe, Reunion, Panama, Ireland, Nigeria, Uruguay, Saudi Arabia, Indonesia, Romania, Bulgaria, Bermuda, Russian Federation, Burundi, Congo, Democratic Republic of, Iceland, Uganda, Cyprus}.
+The below chart shows the top 10 countries ranked by amount invested in
+Mozambique.
+
+\begin{center}\includegraphics{figures/unnamed-chunk-7-1} \end{center}
+
+What stands out is the presence of many ``non-Western'' countries in the
+top 10. Though reliable data only go as far back as 2009, the landscape
+of investment 20 years prior would have been both (a) far smaller and
+(b) primarily composed of western countries, especially given that
+Mozambique emerged so late from civil war (1994). Though as of 2015,
+China did not figure in the top 10 FDI countries in Mozambique, its
+upward trend is notable, with approximately 1.5 million USD invested in
+2010 to nearly 100 million by 2015 (a 60-fold increase).
+
 \subsubsection{Massive increase in FDI}\label{massive-increase-in-fdi}
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1} \end{center}
-
 In addition to the discovery of new resources, recent growth has also
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1} \end{center}
+\begin{center}\includegraphics{figures/unnamed-chunk-10-1} \end{center}
 
@@
 decreases in the costs to import and export, as well as the costs of
-starting a business and registering property (Bank 2016).
+starting a business and registering property (WB 2016).
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1} \end{center}
+\begin{center}\includegraphics{figures/unnamed-chunk-11-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1} \end{center}
+\begin{center}\includegraphics{figures/unnamed-chunk-12-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1} \end{center}
+\begin{center}\includegraphics{figures/unnamed-chunk-13-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1} \end{center}
+\begin{center}\includegraphics{figures/unnamed-chunk-14-1} \end{center}
 
-\subsubsection{CSR in Mozambique}\label{csr-in-mozambique}
-
-Corporate social responsibility in Mozambique is new in nomenclature,
-but activities which could be classified as CSR have existed for
-decades. The number of firms actively engaging in CSR cannot be
-ascertained (given the large and ever-changing number of small
-businesses), but virtually all of the largest firms, both foreign and
-domestic, have a CSR component.
-
-Firms with CSR activities are often large and foreign. Among the largest
-``key players'' in Mozambican CSR are Coca-Cola, British Petroleum and
-Colgate-Palmolive. State conglomerates, such as Águas de Moçambique and
-Electricidade de Moçambique, also participate in CSR activities (Compact
-2007).
-
-\subsubsection{Malaria-related CSR and other corporate
-activity}\label{malaria-related-csr-and-other-corporate-activity}
-
-Most CSR-funded activities are focused on education, community
-development, women's rights and entrepreneurship. According to a UN
-poll, none of the country's largest firms invest directly in
-malaria-related CSR activities (Compact 2007). This may be due to the
-perceived costs of malaria control, lack of perceived PR benefits, the
-government and NGO's predominant role in the area, as well as the
-``opt-in'' nature and privacy/legal issues generally related to engaging
-in health-related campaigns.
-
-That said, a majority of the firms interviewed by the UN indicated that
-one of the principal reasons for investment in CSR is to complement
-government efforts. To the extent that malaria accounts for more of the
-loss in disability-adjusted life years in Mozambique than comparable
-countries (IHME 2015), malaria control's lack of representation among
-core CSR activities is a notable absence.
-
-Whether under the guise of CSR or not, it is noteworthy that the private
-sector currently does play a role in malaria control activities.
-According to 2011 DHS data, greater than 7 in every 1,000 households had
-a private company carry out indoor residual spraying (DHS 2011). And, in
-some clusters, the percent of houses covered by private IRS was greater
-than 25\%.
+FDI is associated with CSR. Whether under the guise of CSR or not, it is
+noteworthy that the private sector currently does play a role in malaria
+control activities. According to 2011 DHS data, greater than 7 in every
+1,000 households had a private company carry out indoor residual
+spraying (DHS 2011). And, in some clusters, the percent of houses
+covered by private IRS was greater than 25\%.
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-12-1} \end{center}
+\begin{center}\includegraphics{figures/unnamed-chunk-16-1} \end{center}
 
@@
 
-One notable exception is the Nando's-lead Goodbye Malaria Trust, an
-umbrella organization which inclues a development impact bond, the
-Goodbye Malaria initiative, and partnerships with several other foreign
-firms operating in Mozambique (more details here).
+Corporate social responsibility in Mozambique is new in nomenclature,
+but activities which could be classified as CSR have existed for
+decades. The number of firms actively engaging in CSR cannot be
+ascertained (given the large and ever-changing number of small
+businesses), but virtually all of the largest firms, both foreign and
+domestic, have a CSR component.
+
+Firms with CSR activities are often large and foreign. Among the largest
+``key players'' in Mozambican CSR are Coca-Cola, British Petroleum and
+Colgate-Palmolive. State conglomerates, such as Águas de Moçambique and
+Electricidade de Moçambique, also participate in CSR activities (Compact
+2007). Large communications firms, such as Vodacom and MCEL, and banks
+(BIM and BCI) have a CSR component, but generally focus on culture,
+environment and sport, avoiding activities which could either complement
+or conflict with the health sector (Kaufmann and Simons-Kaufmann 2015).
+
+Among major foreign companies, those operating in extraction -
+particularly mining - appear to be the most involved with malaria. BHP
+Billiton has been active in vector control support to the governments of
+Mozambique, Swaziland, and South Africa, particularly in the early 2000s
+(BHP 2014). Kropfmuhl AG and Vale, also mining companies, have carried
+out CSR related to both community development and malaria control in
+Northern Mozambique (Kaufmann and Simons-Kaufmann 2015).
+
+Most CSR-funded activities are focused on education, community
+development, women's rights and entrepreneurship. According to a UN
+poll, none of the country's largest firms invest directly in
+malaria-related CSR activities (Compact 2007). That said, a majority of
+the firms interviewed by the UN indicated that one of the principal
+reasons for investment in CSR is to complement government efforts. To
+the extent that malaria accounts for more of the loss in
+disability-adjusted life years in Mozambique than comparable countries
+(IHME 2015), malaria control's lack of representation among core CSR
+activities is a notable absence. One important exception is the
+Nando's-lead Goodbye Malaria Trust, an umbrella organization which
+inclues a development impact bond, the Goodbye Malaria initiative, and
+partnerships with several other foreign firms operating in Mozambique.
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-15-1} \end{center}
+\begin{center}\includegraphics{figures/unnamed-chunk-19-1} \end{center}
 
@@
 \section{Discussion}\label{discussion}
+
+The intensity of CSR activity seems to closely track FDI, but generally
+avoids malaria control. The few malaria-related CSR projects in recent
+years has come in large, ``mega-projects'' whose well-publicized malaria
+abatement activities are profit-driven (Mouzin and al. 2011) or whose
+primary aim is not profit-related (Han 2015). Though the former is often
+portrayed as a ``win-win'' for business and public health, the latter
+also offers tangible benefits for private industry, and should be
+understood as operating under the same conditions and with the same
+motivations. To the extent that most CSR activity has avoided the health
+sector (including malaria) in favor of cultural, environmental,
+education and more general ``development'' activities, it is reasonable
+to assume that this may be due to the perceived costs of malaria
+control, lack of perceived PR benefits, the government and NGO's
+predominant role in the area, as well as the ``opt-in'' nature and
+privacy/legal issues generally related to engaging in health-related
+campaigns.
 
@@
 
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
 
@@
 
+\hypertarget{ref-Alfaro_2006}{}
+Alfaro, Laura, Areendam Chanda, Sebnem Kalemli-Ozcan, and Selin Sayek.
+2006. ``How Does Foreign Direct Investment Promote Economic Growth?
+Exploring the Effects of Financial Markets on Linkages,'' September.
+National Bureau of Economic Research.
+doi:\href{https://doi.org/10.3386/w12522}{10.3386/w12522}.
+
+\hypertarget{ref-Almfraji_2014}{}
+Almfraji, Mohammad Amin, and Mahmoud Khalid Almsafir. 2014. ``Foreign
+Direct Investment and Economic Growth Literature Review from 1994 to
+2012.'' \emph{Procedia - Social and Behavioral Sciences} 129 (May).
+Elsevier BV: 206--13.
+doi:\href{https://doi.org/10.1016/j.sbspro.2014.03.668}{10.1016/j.sbspro.2014.03.668}.
+
+\hypertarget{ref-Asante_2011}{}
+Asante, Kwaku P, Charles Zandoh, Dominic B Dery, Charles Brown, George
+Adjei, Yaw Antwi-Dadzie, Martin Adjuik, et al. 2011. ``Malaria
+Epidemiology in the Ahafo Area of Ghana.'' \emph{Malaria Journal} 10
+(1). Springer Nature: 211.
+doi:\href{https://doi.org/10.1186/1475-2875-10-211}{10.1186/1475-2875-10-211}.
+
 \hypertarget{ref-Azemar2009}{}
@@
 
-\hypertarget{ref-doingbusiness}{}
-Bank, World. 2016. \emph{Doing Business 2016: Measuring Regulatory
-Quality and Efficiency}. World Bank.
-\url{http://www.doingbusiness.org/~/media/GIAWB/Doing\%20Business/Documents/Annual-Reports/English/DB16-Chapters/DB16-Mini-Book.pdf}.
+\hypertarget{ref-bhp}{}
+BHP. 2014. ``BHP Billiton Case Study: Malaria.''
+\url{https://www.bhp.com/-/media/bhp/documents/investors/reports/2004/norvatispresentation.pdf}.
+
+\hypertarget{ref-NBERw17282}{}
+Blonigen, Bruce, and Cheyney O'Fallon. 2011. ``Foreign Firms and Local
+Communities.'' Working Paper 17282. Working Paper Series. National
+Bureau of Economic Research.
+doi:\href{https://doi.org/10.3386/w17282}{10.3386/w17282}.
 
@@
 \url{https://www.quandl.com/data/BUNDESBANK/BBEX3_M_MZN_USD_CA_AC_A01}.
+
+\hypertarget{ref-Burns_2017}{}
+Burns, Darren K., Andrew P. Jones, Yevgeniy Goryakin, and Marc Suhrcke.
+2017. ``Is Foreign Direct Investment Good for Health in Low and Middle
+Income Countries? An Instrumental Variable Approach.'' \emph{Social
+Science \& Medicine} 181 (May). Elsevier BV: 74--82.
+doi:\href{https://doi.org/10.1016/j.socscimed.2017.03.054}{10.1016/j.socscimed.2017.03.054}.
 
@@
 
+\hypertarget{ref-Economist2017}{}
+Economist. 2017. ``Mozambique's Default: Mozambique Fails to Pay Its
+Debts.'' \emph{The Economist}, January. Economist.
+\url{https://www.economist.com/news/middle-east-and-africa/21715030-mozambique-fails-pay-its-debts-mozambiques-default}.
+
+\hypertarget{ref-Feenstra}{}
+Feenstra, Robert, and Gordon Hanson. 1997. ``Foreign Direct Investment
+and Relative Wages: Evidence from Mexico's Maquiladoras.'' \emph{Journal
+of International Economics} 42 (3-4): 371--93.
+\url{https://EconPapers.repec.org/RePEc:eee:inecon:v:42:y:1997:i:3-4:p:371-393}.
+
+\hypertarget{ref-Friedman}{}
+Friedman, Milton. 1970. ``The Social Responsibility of Business Is to
+Increase Its Profits.'' \emph{Corporate Ethics and Corporate
+Governance}. Springer Berlin Heidelberg, 173--78.
+doi:\href{https://doi.org/10.1007/978-3-540-70818-6_14}{10.1007/978-3-540-70818-6\_14}.
+
+\hypertarget{ref-gadm}{}
+Gadm. 2009. ``GADM Database of Global Administrative Areas Version
+1.0.''
+
 \hypertarget{ref-German2013}{}
@@
 
+\hypertarget{ref-Gonz_lez_2015}{}
+González, Raquel, Orvalho J. Augusto, Khátia Munguambe, Charlotte
+Pierrat, Elpidia N. Pedro, Charfudin Sacoor, Elisa De Lazzari, et al.
+2015. ``HIV Incidence and Spatial Clustering in a Rural Area of Southern
+Mozambique.'' Edited by Jean KEditor Carr. \emph{PLOS ONE} 10 (7).
+Public Library of Science (PLoS): e0132053.
+doi:\href{https://doi.org/10.1371/journal.pone.0132053}{10.1371/journal.pone.0132053}.
+
 \hypertarget{ref-Han}{}
@@
 
+\hypertarget{ref-imf}{}
+IMF. 2003. ``International Monetary Fund's Foreign Direct Investment
+Trends and Statistics.''
+\url{https://www.imf.org/external/np/sta/fdi/eng/2003/102803.htm}.
+
 \hypertarget{ref-estatistica}{}
 INE. 2015. ``Economic Statistics.'' \url{http://www.ine.gov.mz/}.
+
+\hypertarget{ref-ine}{}
+---------. 2017. ``Instituto Nacional de Estatistica.''
+\url{http://www.ine.gov.mz/}.
+
+\hypertarget{ref-Kaufmann_2015}{}
+Kaufmann, Friedrich, and Claudia Simons-Kaufmann. 2015. ``Corporate
+Social Responsibility in Mozambique.'' \emph{CSR, Sustainability, Ethics
+\& Governance}, December. Springer International Publishing, 31--50.
+doi:\href{https://doi.org/10.1007/978-3-319-26668-8_2}{10.1007/978-3-319-26668-8\_2}.
 
@@
 
+\hypertarget{ref-Moonasar_2016}{}
+Moonasar, Devanand, Rajendra Maharaj, Simon Kunene, Baltazar Candrinho,
+Francisco Saute, Nyasatu Ntshalintshali, and Natashia Morris. 2016.
+``Towards Malaria Elimination in the Mosaswa (Mozambique, South Africa
+and Swaziland) Region.'' \emph{Malaria Journal} 15 (1). Springer Nature.
+doi:\href{https://doi.org/10.1186/s12936-016-1470-8}{10.1186/s12936-016-1470-8}.
+
 \hypertarget{ref-Mouzin2011}{}
@@
 Impact Series}, no. 6.
+
+\hypertarget{ref-Nonvignon_2016}{}
+Nonvignon, Justice, Genevieve Cecilia Aryeetey, Keziah L. Malm, Samuel
+Agyei Agyemang, Vivian N. A. Aubyn, Nana Yaw Peprah, Constance N.
+Bart-Plange, and Moses Aikins. 2016. ``Economic Burden of Malaria on
+Businesses in Ghana: A Case for Private Sector Investment in Malaria
+Control.'' \emph{Malaria Journal} 15 (1). Springer Nature.
+doi:\href{https://doi.org/10.1186/s12936-016-1506-0}{10.1186/s12936-016-1506-0}.
+
+\hypertarget{ref-Ortega_2016}{}
+Ortega, María Isabel, Samantha Sabo, Patricia Aranda Gallegos, Jill
+Eileen Guernsey De Zapien, Antonio Zapien, Gloria Elena Portillo Abril,
+and Cecilia Rosales. 2016. ``Agribusiness, Corporate Social
+Responsibility, and Health of Agricultural Migrant Workers.''
+\emph{Frontiers in Public Health} 4 (March). Frontiers Media SA.
+doi:\href{https://doi.org/10.3389/fpubh.2016.00054}{10.3389/fpubh.2016.00054}.
+
+\hypertarget{ref-Pazienza_2015}{}
+Pazienza, Pasquale. 2015. ``The Relationship Between Co2 and Foreign
+Direct Investment in the Agriculture and Fishing Sector of Oecd
+Countries: Evidence and Policy Considerations.'' \emph{Intellectual
+Economics} 9 (1). Elsevier BV: 55--66.
+doi:\href{https://doi.org/10.1016/j.intele.2015.08.001}{10.1016/j.intele.2015.08.001}.
+
+\hypertarget{ref-Pfeiffer_2017}{}
+Pfeiffer, James, Sarah Gimbel, Baltazar Chilundo, Stephen Gloyd, Rachel
+Chapman, and Kenneth Sherr. 2017. ``Austerity and the `Sector-Wide
+Approach' to Health: The Mozambique Experience.'' \emph{Social Science
+\& Medicine} 187 (August). Elsevier BV: 208--16.
+doi:\href{https://doi.org/10.1016/j.socscimed.2017.05.008}{10.1016/j.socscimed.2017.05.008}.
+
+\hypertarget{ref-Porter2002}{}
+Porter, KE, and MR Kramer. 2002. ``The Competitive Advantage of
+Corporate Philanthropy.'' \emph{Harvard Business Review}, no. 80.
+Harvard University Press.
+doi:\href{https://doi.org/10.1186/s12936-016-1506-0}{10.1186/s12936-016-1506-0}.
 
@@
 
+\hypertarget{ref-Selaya_2012}{}
+Selaya, Pablo, and Eva Rytter Sunesen. 2012. ``Does Foreign Aid Increase
+Foreign Direct Investment?'' \emph{World Development} 40 (11). Elsevier
+BV: 2155--76.
+doi:\href{https://doi.org/10.1016/j.worlddev.2012.06.001}{10.1016/j.worlddev.2012.06.001}.
+
+\hypertarget{ref-Shahbaz_2015}{}
+Shahbaz, Muhammad, Samia Nasreen, Faisal Abbas, and Omri Anis. 2015.
+``Does Foreign Direct Investment Impede Environmental Quality in High-,
+Middle-, and Low-Income Countries?'' \emph{Energy Economics} 51
+(September). Elsevier BV: 275--87.
+doi:\href{https://doi.org/10.1016/j.eneco.2015.06.014}{10.1016/j.eneco.2015.06.014}.
+
+\hypertarget{ref-Sheela_2017}{}
+Sheela, A.M., A. Ghermandi, P. Vineetha, R.V. Sheeja, J. Justus, and K.
+Ajayakrishna. 2017. ``Assessment of Relation of Land Use Characteristics
+with Vector-Borne Diseases in Tropical Areas.'' \emph{Land Use Policy}
+63 (April). Elsevier BV: 369--80.
+doi:\href{https://doi.org/10.1016/j.landusepol.2017.01.047}{10.1016/j.landusepol.2017.01.047}.
+
+\hypertarget{ref-Sutton2014}{}
+Sutton, John. 2014. \emph{MAPA Empresarial de Moçambique}. International
+Growth Centre.
+\url{http://personal.lse.ac.uk/sutton/mozambique_portuguese_edn_updated_version_web.pdf}.
+
+\hypertarget{ref-Unctad2012}{}
+UNCTAD. 2012. \emph{Investment Policy Review: Mozambique}. United
+Nations.
+
 \hypertarget{ref-wbdata}{}
@@
 Us\$).'' \url{http://data.worldbank.org/indicator/BX.KLT.DINV.CD.WD}.
+
+\hypertarget{ref-doingbusiness}{}
+---------. 2016. \emph{Doing Business 2016: Measuring Regulatory Quality
+and Efficiency}. World Bank.
+\url{http://www.doingbusiness.org/~/media/GIAWB/Doing\%20Business/Documents/Annual-Reports/English/DB16-Chapters/DB16-Mini-Book.pdf}.
+
+\hypertarget{ref-wbd}{}
+---------. 2017. ``World Bank Open Data.''
+\url{https://data.worldbank.org/}.
+
+\hypertarget{ref-White2014}{}
+White, Nicholas J, Sasithon Pukrittayakamee, Tran Tinh Hien, M Abul
+Faiz, Olugbenga A Mokuolu, and Arjen M Dondorp. 2014. ``Malaria.''
+\emph{The Lancet} 383 (9918). Elsevier BV: 723--35.
+doi:\href{https://doi.org/10.1016/s0140-6736(13)60024-0}{10.1016/s0140-6736(13)60024-0}.
+
+\hypertarget{ref-World2016}{}
+WHO. 2016. ``World Malaria Report.''
+\url{http://apps.who.int/iris/bitstream/10665/252038/1/9789241511711-eng.pdf?ua=1}.
+
+\hypertarget{ref-Williams_2015}{}
+Williams, Brian G., Eleanor Gouws, Pierre Somse, Mpho Mmelesi, Chibwe
+Lwamba, Trouble Chikoko, Erika Fazito, et al. 2015. ``Epidemiological
+Trends for Hiv in Southern Africa: Implications for Reaching the
+Elimination Targets.'' \emph{Current HIV/AIDS Reports} 12 (2). Springer
+Nature: 196--206.
+doi:\href{https://doi.org/10.1007/s11904-015-0264-x}{10.1007/s11904-015-0264-x}.
 
@@
 
+\hypertarget{ref-Zhu_2013}{}
+Zhu, Yan, Li-Yun Sun, and Alicia S. M. Leung. 2013. ``Corporate Social
+Responsibility, Firm Reputation, and Firm Performance: The Role of
+Ethical Leadership.'' \emph{Asia Pacific Journal of Management} 31 (4).
+Springer Nature: 925--47.
+doi:\href{https://doi.org/10.1007/s10490-013-9369-1}{10.1007/s10490-013-9369-1}.
+
 \end{document}

### 2019-06-19 paper.Rmd
major update

@@
 fig_height: 2.6
-bibliography: library.bib
+bibliography: fdi_moz.bib
 ---
@@
 Mozambique's economic growth up to 2014 was facilitated by the government's open policy towards foreign direct investment (FDI), a plentiful supply of natural resources [@Rogers], and relatively inexpensive labor.
-Foreign investment in Mozambique is largely geared towards the export market, and heavily concentrated.
+These factors combined to push Maputo (Mozambique's capital) into the top-20 African countries in terms of gross FDI [@africa2018].
+Its location (a large, eastward-facing coastline, proximity to rapidly developing South Africa, and integration into trade corridor and zones such as those with Zimbabwe [@africa2018] and the Southern African community as a whole [@Asiedu_2005]) are favorable to foreign investment, which is largely geared towards the export market, and heavily concentrated.
 63% of exports come from aluminum, electricity, minerals, and gas, with each sector displaying high degrees of concentration (greater than 50% of the export share of the previous four industries are attributable to one company) [@Sutton2014].
@@
 Though reliable data only go as far back as 2009, the landscape of investment 20 years prior would have been both (a) far smaller and (b) primarily composed of western countries, especially given that Mozambique emerged so late from civil war (1994).
+As of 2014, Chinese capital had entered into 52 countries in Africa [@africa2018], and though more recent data are not available, it's likely that this figure has grown since then.
 Though as of 2015, China did not figure in the top 10 FDI countries in Mozambique, its upward trend is notable, with approximately 1.5 million USD invested in 2010 to nearly 100 million by 2015 (a 60-fold increase).
+As of 2017, 3.2% of Chinese FDI flows went to Mozambique [@He_2015].
+Though much of the growth is fueled by centralized State actors, an increasing number of provincial and decentralized investors are playing a role in African FDI [@Gu_2016].
 
@@
 macro_names <- names(db)
+macro_names <- gsub('...', 'X__...', macro_names, fixed = TRUE)
 macro_names <- ifelse(grepl('X__', macro_names), 'Var', macro_names)
@@
 # Get columns for coordinates
-temp$x <- coordinates(temp)[,1]
-temp$y <- coordinates(temp)[,2]
+temp@data$x <- coordinates(temp)[,1]
+temp@data$y <- coordinates(temp)[,2]
+
+# Get a better id
+temp@data$id <-
+  unlist(lapply(strsplit(gsub('MOZ.', '', temp@data$GID_1), '_'), function(x){x[1]}))
 
@@
 temp_fort <- left_join(x = temp_fort,
-                       y = temp@data %>%
-                         mutate(id = as.character(ID_1)))
+                       y = temp@data)
 
@@
 # rtp@data$id <- 1:nrow(rtp@data)
+
 ```
@@
 State conglomerates, such as Águas de Moçambique and Electricidade de Moçambique, also participate in CSR activities [@Compact2007].
-Large communications firms, such as Vodacom and MCEL, and banks (BIM and BCI) have a CSR component, but generally focus on culture, environment and sport, avoiding activities which could either complement or conflict with the health sector [@Kaufmann_2015].
+Large communications firms, such as Vodacom and MCEL, and banks (BIM and BCI) have a CSR component, but generally focus on culture, environment and sport, avoiding activities which could either complement or conflict with the health sector.
+Much of CSR in Mozambique is basic servcies only, a phenomenon Kaufmann described as "CSR towards compliance" as opposed to "beyond compliance" [@Kaufmann_2015].
 
@@
 The few malaria-related CSR projects in recent years has come in large, "mega-projects" whose well-publicized malaria abatement activities are profit-driven [@Mouzin2011] or whose primary aim is not profit-related [@Han].
+This should come as no surprise given that "mega-projects" make up a majority of FDI to Mozambique (albeit declining slightly in recent years) [@Unctad2012].
 Though the former is often portrayed as a "win-win" for business and public health, the latter also offers tangible benefits for private industry, and should be understood as operating under the same conditions and with the same motivations.
@@
 Private firms' incentives, though aligned with the public's in terms of malaria, are not identical, and pressures from shareholders and for positive public perception might motivate malaria control strategies which do not necessarily carry with them the most recent scientific knowledge.
+The methods by which private initiatives could contribute to malaria elimination and control while at the same time only "complementing" (rather than "substituting") State activities is not entirely clear, and would require significant organization [@Kaufmann_2015].
 
@@
 High FDI in Mozambique and growing interest in CSR both call for increased reflection on the private sector's role in the delivery of public goods.
-A refocusing of CSR expenditures into areas where the need is greatest (specifically, malaria control) could lead to better health and greater profits.
+A refocusing of CSR expenditures into areas where the need is greatest (specifically, malaria control) could lead to better health and greater profits, and go beyond the perception of superficial CSR which doesn't address core social problems (ie, "green washing") [@Kaufmann_2015].
 This "win-win" for the public and private sectors represents a rare opportunity, which deserves more discourse, research and experimentation.

### 2019-06-19 paper.tex
major update

@@
 % Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
+\providecommand{\subtitle}[1]{
   \posttitle{
@@
 \setlength{\droptitle}{-2em}
+
   \title{Foreign direct investment, corporate social responsibility, and malaria
@@
 plentiful supply of natural resources (Rogers 2014), and relatively
-inexpensive labor. Foreign investment in Mozambique is largely geared
-towards the export market, and heavily concentrated. 63\% of exports
-come from aluminum, electricity, minerals, and gas, with each sector
-displaying high degrees of concentration (greater than 50\% of the
-export share of the previous four industries are attributable to one
-company) (Sutton 2014). The total number of foreign enterprises
-operating in Mozambique is not ascertainable, but in both quantity and
-diversity of sources, FDI has increased dramatically in recent years
-(Sutton 2014). Mozambique was the Sub-Saharan African country with the
-greatest increase in FDI (defined by the World Bank as cross-border
-investment to establish a lasting interest) inflows from 2006 to 2014,
-registering a boom between 2010 and 2014 of about 5000 million USD
-(UNCTAD 2012).
+inexpensive labor. These factors combined to push Maputo (Mozambique's
+capital) into the top-20 African countries in terms of gross FDI
+(UNHabitat 2018). Its location (a large, eastward-facing coastline,
+proximity to rapidly developing South Africa, and integration into trade
+corridor and zones such as those with Zimbabwe (UNHabitat 2018) and the
+Southern African community as a whole (Asiedu 2005)) are favorable to
+foreign investment, which is largely geared towards the export market,
+and heavily concentrated. 63\% of exports come from aluminum,
+electricity, minerals, and gas, with each sector displaying high degrees
+of concentration (greater than 50\% of the export share of the previous
+four industries are attributable to one company) (Sutton 2014). The
+total number of foreign enterprises operating in Mozambique is not
+ascertainable, but in both quantity and diversity of sources, FDI has
+increased dramatically in recent years (Sutton 2014). Mozambique was the
+Sub-Saharan African country with the greatest increase in FDI (defined
+by the World Bank as cross-border investment to establish a lasting
+interest) inflows from 2006 to 2014, registering a boom between 2010 and
+2014 of about 5000 million USD (UNCTAD 2012).
 
@@
 this is particularly useful for foreign firms that have no previous
-contacts with the host country/communities (Zhu, Sun, and Leung 2013).
+contacts with the host country/communities (Y. Zhu, Sun, and Leung
+2013).
 
@@
 (b) primarily composed of western countries, especially given that
-Mozambique emerged so late from civil war (1994). Though as of 2015,
-China did not figure in the top 10 FDI countries in Mozambique, its
-upward trend is notable, with approximately 1.5 million USD invested in
-2010 to nearly 100 million by 2015 (a 60-fold increase).
+Mozambique emerged so late from civil war (1994). As of 2014, Chinese
+capital had entered into 52 countries in Africa (UNHabitat 2018), and
+though more recent data are not available, it's likely that this figure
+has grown since then. Though as of 2015, China did not figure in the top
+10 FDI countries in Mozambique, its upward trend is notable, with
+approximately 1.5 million USD invested in 2010 to nearly 100 million by
+2015 (a 60-fold increase). As of 2017, 3.2\% of Chinese FDI flows went
+to Mozambique (He, Xie, and Zhu 2015). Though much of the growth is
+fueled by centralized State actors, an increasing number of provincial
+and decentralized investors are playing a role in African FDI (Gu et al.
+2016).
 
@@
 environment and sport, avoiding activities which could either complement
-or conflict with the health sector (Kaufmann and Simons-Kaufmann 2015).
+or conflict with the health sector. Much of CSR in Mozambique is basic
+servcies only, a phenomenon Kaufmann described as ``CSR towards
+compliance'' as opposed to ``beyond compliance'' (Kaufmann and
+Simons-Kaufmann 2015).
 
@@
 abatement activities are profit-driven (Mouzin and al. 2011) or whose
-primary aim is not profit-related (Han 2015). Though the former is often
-portrayed as a ``win-win'' for business and public health, the latter
-also offers tangible benefits for private industry, and should be
-understood as operating under the same conditions and with the same
-motivations. To the extent that most CSR activity has avoided the health
-sector (including malaria) in favor of cultural, environmental,
-education and more general ``development'' activities, it is reasonable
-to assume that this may be due to the perceived costs of malaria
-control, lack of perceived PR benefits, the government and NGO's
-predominant role in the area, as well as the ``opt-in'' nature and
-privacy/legal issues generally related to engaging in health-related
-campaigns.
+primary aim is not profit-related (Han 2015). This should come as no
+surprise given that ``mega-projects'' make up a majority of FDI to
+Mozambique (albeit declining slightly in recent years) (UNCTAD 2012).
+Though the former is often portrayed as a ``win-win'' for business and
+public health, the latter also offers tangible benefits for private
+industry, and should be understood as operating under the same
+conditions and with the same motivations. To the extent that most CSR
+activity has avoided the health sector (including malaria) in favor of
+cultural, environmental, education and more general ``development''
+activities, it is reasonable to assume that this may be due to the
+perceived costs of malaria control, lack of perceived PR benefits, the
+government and NGO's predominant role in the area, as well as the
+``opt-in'' nature and privacy/legal issues generally related to engaging
+in health-related campaigns.
 
@@
 which do not necessarily carry with them the most recent scientific
-knowledge.
+knowledge. The methods by which private initiatives could contribute to
+malaria elimination and control while at the same time only
+``complementing'' (rather than ``substituting'') State activities is not
+entirely clear, and would require significant organization (Kaufmann and
+Simons-Kaufmann 2015).
 
@@
 is greatest (specifically, malaria control) could lead to better health
-and greater profits. This ``win-win'' for the public and private sectors
-represents a rare opportunity, which deserves more discourse, research
-and experimentation.
+and greater profits, and go beyond the perception of superficial CSR
+which doesn't address core social problems (ie, ``green washing'')
+(Kaufmann and Simons-Kaufmann 2015). This ``win-win'' for the public and
+private sectors represents a rare opportunity, which deserves more
+discourse, research and experimentation.
 
@@
 
+\hypertarget{ref-Asiedu_2005}{}
+Asiedu, Elizabeth. 2005. ``Foreign Direct Investment in Africa: The Role
+of Natural Resources, Market Size, Government Policy, Institutions and
+Political Instability.'' \emph{SSRN Electronic Journal}. Elsevier BV.
+doi:\href{https://doi.org/10.2139/ssrn.717361}{10.2139/ssrn.717361}.
+
 \hypertarget{ref-Azemar2009}{}
@@
 
+\hypertarget{ref-Gu_2016}{}
+Gu, Jing, Chuanhong Zhang, Alcides Vaz, and Langton Mukwereza. 2016.
+``Chinese State Capitalism? Rethinking the Role of the State and
+Business in Chinese Development Cooperation in Africa.'' \emph{World
+Development} 81 (May). Elsevier BV: 24--34.
+doi:\href{https://doi.org/10.1016/j.worlddev.2016.01.001}{10.1016/j.worlddev.2016.01.001}.
+
 \hypertarget{ref-Han}{}
@@
 \url{http://www.theguardian.com/global-development-professionals-network/2014/mar/31/malaria-control-payment-by-results}.
+
+\hypertarget{ref-He_2015}{}
+He, Canfei, Xiuzhen Xie, and Shengjun Zhu. 2015. ``Going Global:
+Understanding China's Outward Foreign Direct Investment from
+Motivational and Institutional Perspectives.'' \emph{Post-Communist
+Economies} 27 (4). Informa UK Limited: 448--71.
+doi:\href{https://doi.org/10.1080/14631377.2015.1084716}{10.1080/14631377.2015.1084716}.
 
@@
 
+\hypertarget{ref-africa2018}{}
+UNHabitat. 2018. \emph{The State of African Cities 2018, The geography
+of African investment}. United Nations.
+
 \hypertarget{ref-wbdata}{}

### 2019-06-19 paper_word.Rmd
major update

--- 2017-05-22 paper_word.Rmd
@@
 title: Foreign direct investment, corporate social responsibility, and malaria control in Mozambique - trends, risks, and opportunities
-date: "April 5, 2017"
 output: word_document
-bibliography: library.bib
+bibliography: fdi_moz.bib
 
@@
 
+# Authorship
+source('make_authors.R')
+
 # BASIC KNITR OPTIONS
-opts_chunk$set(comment = NA, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, fig.align = 'center', fig.height = 3, fig.width = 4)
+opts_chunk$set(comment = NA, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, fig.align = 'center', fig.height = 4, fig.width = 5, fig.path = "figures/")
 options(xtable.comment = FALSE)
@@
 
+\end{changemargin}
+\end{center}
+
 ## Introduction
 
-Mozabique's recent economic growth has been facilitated by the government's open policy towards foreign investment, a plentiful supply of natural resources (particularly minerals and hydrocarbons) [@Rogers] and relatively inexpensive labor.
-Improvements in health have accompanied economic expansion, but Mozambique still lags behind in basic health outcomes - particularly those related to malaria - even by regional standards.
-Most economic activities have to deal
-with malaria: particularly agricultural-based and extraction activities cannot be carried out without consistent investment towards malaria management, both
-prevention and prompt treatment. This can be done as part of the “normal activities” of private sector or can go slightly beyond what is strictly needed for the
-activities and be seen as part of the corporate social responsibilityDespite  strong evidence showing that improved health yields substantial economic benefits [@World1999, @Bloom2008], there is little evidence regarding the extent to which "corporate social responsibility" (CSR) activities of foreign firms have targeted or impacted malaria.
-Understanding the relationship between FDI, CSR and malaria is essential to identifying which incentives and barriers to private sector involvement in malaria elimination exist, as well as identifying potential opportunities and synergies for scaling up private sector involvment in malaria elimination.
+Mozambique is a Southern-East African country with a surface area of nearly 800,000 m2 (more than 3 times that of the United Kingdom) and a population of nearly 30 million inhabitants.
+Though low-income, Mozambique experienced an upward trend in gross domestic product (GDP) per capita from approximately 200 USD in the late 90s to 623 in 2014, followed by a steep decrease in 2015 down to 382 in 2016 [@doingbusiness].
 
-That said, 
-The few malaria-related CSR projects in recent years has come in large, "mega-projects" whose well-publicized malaria abatement activities are profit-driven [@Mouzin2011] or whose primary aim is not profit-related [@Han].
-Though the former is often portrayed as a "win-win" for business and public health, the latter also offers tangible benefits for private industry, and should be understood as operating under the same conditions and with the same motivations.
+Mozambique's economic growth up to 2014 was facilitated by the government's open policy towards foreign direct investment (FDI), a plentiful supply of natural resources [@Rogers], and relatively inexpensive labor.
+These factors combined to push Maputo (Mozambique's capital) into the top-20 African countries in terms of gross FDI [@africa2018].
+Its location (a large, eastward-facing coastline, proximity to rapidly developing South Africa, and integration into trade corridor and zones such as those with Zimbabwe [@africa2018] and the Southern African community as a whole [@Asiedu_2005]) are favorable to foreign investment, which is largely geared towards the export market, and heavily concentrated.
+63% of exports come from aluminum, electricity, minerals, and gas, with each sector displaying high degrees of concentration (greater than 50% of the export share of the previous four industries are attributable to one company) [@Sutton2014].
+The total number of foreign enterprises operating in Mozambique is not ascertainable, but in both quantity and diversity of sources, FDI has increased dramatically in recent years [@Sutton2014].
+Mozambique was the Sub-Saharan African country with the greatest increase in FDI (defined by the World Bank as cross-border investment to establish a lasting interest) inflows from 2006 to 2014, registering a boom between 2010 and 2014 of about 5000 million USD [@Unctad2012].
 
-Private foreign investors showed a great optimism during the FMI meeting held in Maputo in 2014 (reference needed).
+The impact of FDI on economic growth and other economic indicators has been largely assessed [@Alfaro_2006] [@Almfraji_2014] [@World1999].
+As FDI has been found to improve working conditions in low-income countries and, consequently, population ability to pay [@Bloom2008], they are likely to improve access to healthcare and, therefore, improve health [@Feenstra].
+However, there is some evidence that FDI may have adverse effects on health, through increasing the consumption of “bads” such as tobacco and alcohol and the level of harmful pollution [@Pazienza_2015] [@Shahbaz_2015].
+It has been recently assessed that FDIs are weakly associated with a marginal benefit in adults’ life expectancy in low and middle income countries (no impact has been found on children’s health), yet investments into the secondary sector (for being polluting) have been found potentially harmful [@Burns_2017].
+
+In Mozambique, improvements in health have accompanied economic expansion, but the country still lags behind in many basic health outcomes [@Williams_2015].
+For example, clusters exist in Southern Mozambique with HIV prevalence as high as 50\% [@Gonz_lez_2015].
+
+Malaria is a protozoan parasitic disease, transmitted to humans by mosquitoes.
+The _Plasmodium falciparum_ species, transmitted by the female _Anopheles_ mosquito, accounts for a large majority of deaths [@White2014].
+Estimates for the year 2015 quantify the burden of malaria in Mozambique in 8,300,000 cases and 15,000 associated deaths [@World2016].
+Despite the still high malaria burden, Mozambique is part of the "Elimination Eight (E8)" malaria initative, with the country-specific goal of malarai elimination by 2030.
+Elimination efforts have been particularly active in Southern Mozambique, in border areas contiguous with Swaziland and South Africa, where the goal is elimination by 2020 [@Moonasar_2016].
+
+Several economic activities are strictly linked to malaria.
+On the one hand, particularly agricultural-based and extraction activities are responsible for considerable increases in the malaria burden, through land and environment modifications that favor the efficiency of mosquitos in spreading the malaria parasites [@Sheela_2017].
+On the other hand, the same activities, are barely carried out without consistent investment towards malaria management, both prevention and prompt treatment: firms need to protect their workers against infections to maintain their productivity high.
+In this regards, there is little evidence on economic burden malaria places on the private sector in endemic countries: however, available reports point to profitability of founding malaria control [@Nonvignon_2016].
+Malaria control carried out by some multinational companies have been highly successful and strategies implemented were adopted as a model for the whole country [@Asante_2011].
+
+Private investments in malaria control can be done as part of the “normal actions” of private sector or can go slightly beyond what is strictly needed for having workers’ health to carry out activities and be seen as part of the corporate social responsibility (CSR).
+CSR is defined as those cases in which a company goes "beyond the legal requirements of the country in which they operate" in their work towards that country's long-term interests, in such a systematic manner that it becomes "business as usual" [@Ortega_2016].
+
+Motivations for engaging in CSR has been analyzed in the economic literature.
+Enterprises would have no incentives to engage in CSR according to Friedman [@Friedman] as firms are organized to maximize their profits only.
+The opposite view is that CSR is actually a strategy for maximizing profits.
+According to this vision, CSR helps enterprises to make their products known and, therefore, foment consumers’ demand [@Porter2002].
+CSR also helps building a positive reputation of enterprises among institutions and communities where they operate: this is particularly useful for foreign firms that have no previous contacts with the host country/communities [@Zhu_2013].
+
+A study has recently looked at the propensity and at the magnitude of CSR to the host communities in the US by comparing foreign and local enterprises [@NBERw17282].
+They found that foreign-owned enterprises are less likely to give, but that when they do give, it is substantially more in magnitude than domestic firms, everything else equal.
+Their null hypothesis was that foreign firms give as much as local ones based on the assumptions that firms decide on CSR decisions from a strict profit maximization perspective and that, therefore, foreign or local ownership does not affect such a choice.
+One alternative hypothesis was that foreign firms are less likely to give or give less than local because their products are mainly not for the local markets and because foreign managers have higher propensity to give not locally but either internationally or to their country of origin.
+A second alternative hypothesis is that foreign firms give more than the local ones.
+This may be supported by the fact that foreign companies have different culture from local, more in favor of giving or because foreign firms use CSR to overcome political and cultural barriers in the host country that would not be able to mitigate differently.
+
+The relationship between FDI and foreign aid may also be an interesting determinant of CSR: if aid from a particular country increases FDI from the same country to the same host communities, incentives to engage in CSR from foreign private companies may disappear [@Selaya_2012].
+In other words, why should foreign companies donate to the host country when the government of the country of origin is already giving?
+
+Private foreign investors showed a great optimism during the International Monetary Fund meeting held in Maputo, the Mozambican city capital, in 2014 [@Pfeiffer_2017].
 Beginning in late 2015, the flood of FDI into Mozambique (and many other subsaharan African nations) slowed to a trickle.
 The impact of this slowdown on CSR is not yet known, but it can reasonably be assumed that it will mean a reduction in CSR activities (albeit with lag).
-Almost in parallel to this, Mozambique has had serious financial problems which have
-resulted in default (reference needed http://www.economist.com/news/middleeast-and-africa/21715030-mozambique-fails-pay-its-debtsmozambiques-default).
+Almost in parallel to this, Mozambique has had serious financial problems which have resulted in default [@Economist2017].
 Given the rapidly changing economic and epidemiologic context in Mozambique, a comprehensive and current understanding of both (a) the landscape of FDI and CSR in Mozambique and (b) public health issues (namely, malaria control) which are directly affected those investments is sorely needed.
@@
 ## Quantitiative datasets pertaining to FDI and CSR in Mozambique
+
+```{r}
+# Define function for reading fdi data from imf
+if('imf.RData' %in% dir()){
+  load('imf.RData')
+} else {
+  read_imf <- function(file){
+  srcs <- readxl::read_excel(file, skip = 4)
+  # Rename columns
+  srcs <- srcs[,1:3]
+  srcs <- srcs[2:nrow(srcs),]
+  names(srcs) <- c('from_country', 'inward', 'outward')
+  # Make numberic
+  srcs <- data.frame(srcs)
+  for (j in 2:3){
+    this_column <- as.numeric(as.character(gsub('.', '', srcs[,j], fixed = TRUE)))
+    srcs[,j] <- this_column
+  }
+  srcs <- srcs %>%
+    filter(from_country != 'World')
+
+  # Get amount by inward
+  x <- srcs %>%
+    arrange(desc(inward)) %>%
+    filter(!is.na(inward) & inward > 0) %>%
+    mutate(from_country = ifelse(from_country == 'China, P.R.: Mainland',
+                                 'China',
+                                 ifelse(from_country == 'United Arab Emirates',
+                                        'U.A.E.',
+                                        ifelse(from_country == 'United Kingdom',
+                                               'U.K.',
+                                               from_country)))) #%>%
+    # mutate(from_country = factor(from_country, levels = from_country))
+  return(x)
+}
+
+  years <- 2009:2015
+  results_list <- list()
+  for (i in 1:length(years)){
+    this_year <- years[i]
+    this_data <- read_imf(paste0('data/imf_inward_direct_investment_positions/', this_year, '.xls')) %>%
+      mutate(year = this_year)
+    results_list[[i]] <- this_data
+  }
+  x <- bind_rows(results_list)
+  save(x, file = 'imf.RData')
+}
+```
+
+According to the IMF, `r length(x$from_country[x$year == 2015 & !is.na(x$inward)])` countries had foreign direct investment in Mozambique as of the end of 2015\footnote{`r paste0(x$from_country[x$year == 2015 & !is.na(x$inward)], collapse = ', ')`}.
+The below chart shows the top 10 countries ranked by amount invested in Mozambique.
+
+```{r}
+last_year <- x %>%
+  filter(year == 2015)
+last_year <- last_year %>% 
+  arrange(desc(inward)) %>%
+  mutate(from_country = factor(from_country, levels = as.character(from_country)))
+last_year <- last_year[1:10,]
+ggplot(data = last_year,
+       aes(x = from_country,
+           y = inward)) +
+  geom_bar(stat = 'identity',
+           alpha = 0.6,
+           fill = 'darkblue') +
+  labs(x = 'Source of investment',
+       y = '2015 USD, Millions',
+       title = 'Sources of FDI in Mozambique',
+       subtitle = 'By investor country') +
+  theme_bw() +
+    theme(axis.text.x = element_text(angle = 90)) 
+```
+
+What stands out is the presence of many "non-Western" countries in the top 10.
+Though reliable data only go as far back as 2009, the landscape of investment 20 years prior would have been both (a) far smaller and (b) primarily composed of western countries, especially given that Mozambique emerged so late from civil war (1994).
+As of 2014, Chinese capital had entered into 52 countries in Africa [@africa2018], and though more recent data are not available, it's likely that this figure has grown since then.
+Though as of 2015, China did not figure in the top 10 FDI countries in Mozambique, its upward trend is notable, with approximately 1.5 million USD invested in 2010 to nearly 100 million by 2015 (a 60-fold increase).
+As of 2017, 3.2% of Chinese FDI flows went to Mozambique [@He_2015].
+Though much of the growth is fueled by centralized State actors, an increasing number of provincial and decentralized investors are playing a role in African FDI [@Gu_2016].
+
+```{r, eval = FALSE}
+china <- x %>%
+  filter(from_country == 'China')
+
+cols <- colorRampPalette(brewer.pal(n = 9, name = 'Spectral'))(nrow(last_year))
+cols[3] <- 'black'
+ggplot(data = x %>%
+         filter(from_country %in% last_year$from_country),
+       aes(x = year,
+           y = inward,
+           group = from_country,
+           color = from_country)) +
+  geom_line() +
+  scale_color_manual(name = '',
+                     values = cols) +
+labs(x = 'Source of investment',
+       y = '2015 USD, Millions',
+       title = 'Sources of FDI in Mozambique over time',
+       subtitle = 'By investor country') +
+  theme_bw() +
+  theme(legend.position = 'bottom') +
+  guides(col = guide_legend(nrow = 4))
+```
 
@@
 macro_names <- names(db)
+macro_names <- gsub('...', 'X__...', macro_names, fixed = TRUE)
+macro_names <- ifelse(grepl('X__', macro_names), 'Var', macro_names)
+
 for (j in 1:length(macro_names)){
@@
     theme_bw() +
-   theme(strip.text = element_text(size=5))
+   theme(strip.text = element_text(size=5)) +
+  labs(x = 'Year', y = '%')
 
-multiplot(g1, g2)  
+multiplot(g3, g4)  
 ```
@@
 
-One notable exception is the Nando's-lead Goodbye Malaria Trust, an umbrella organization which inclues a development impact bond, the Goodbye Malaria initiative, and partnerships with several other foreign firms operating in Mozambique (more details here).
+Corporate social responsibility in Mozambique is new in nomenclature, but activities which could be classified as CSR have existed for decades.
+The number of firms actively engaging in CSR cannot be ascertained (given the large and ever-changing number of small businesses), but virtually all of the largest firms, both foreign and domestic, have a CSR component.
+
+Firms with CSR activities are often large and foreign.
+Among the largest "key players" in Mozambican CSR are Coca-Cola, British Petroleum and Colgate-Palmolive.
+State conglomerates, such as Águas de Moçambique and Electricidade de Moçambique, also participate in CSR activities [@Compact2007].
+Large communications firms, such as Vodacom and MCEL, and banks (BIM and BCI) have a CSR component, but generally focus on culture, environment and sport, avoiding activities which could either complement or conflict with the health sector.
+Much of CSR in Mozambique is basic servcies only, a phenomenon Kaufmann described as "CSR towards compliance" as opposed to "beyond compliance" [@Kaufmann_2015].
+
+Among major foreign companies, those operating in extraction - particularly mining - appear to be the most involved with malaria.
+BHP Billiton has been active in vector control support to the governments of Mozambique, Swaziland, and South Africa, particularly in the early 2000s [@bhp].
+Kropfmuhl AG and Vale, also mining companies, have carried out CSR related to both community development and malaria control in Northern Mozambique [@Kaufmann_2015].
+
+Most CSR-funded activities are focused on education, community development, women's rights and entrepreneurship.
+According to a UN poll, none of the country's largest firms invest directly in malaria-related CSR activities [@Compact2007].
+That said, a majority of the firms interviewed by the UN indicated that one of the principal reasons for investment in CSR is to complement government efforts.
+To the extent that malaria accounts for more of the loss in disability-adjusted life years in Mozambique than comparable countries [@ihme], malaria control's lack of representation among core CSR activities is a notable absence.
+One important exception is the Nando's-lead Goodbye Malaria Trust, an umbrella organization which inclues a development impact bond, the Goodbye Malaria initiative, and partnerships with several other foreign firms operating in Mozambique.
 
@@
 
+The intensity of CSR activity seems to closely track FDI, but generally avoids malaria control. 
+The few malaria-related CSR projects in recent years has come in large, "mega-projects" whose well-publicized malaria abatement activities are profit-driven [@Mouzin2011] or whose primary aim is not profit-related [@Han].
+This should come as no surprise given that "mega-projects" make up a majority of FDI to Mozambique (albeit declining slightly in recent years) [@Unctad2012].
+Though the former is often portrayed as a "win-win" for business and public health, the latter also offers tangible benefits for private industry, and should be understood as operating under the same conditions and with the same motivations.
+To the extent that most CSR activity has avoided the health sector (including malaria) in favor of cultural, environmental, education and more general "development" activities, it is reasonable to assume that this may be due to the perceived costs of malaria control, lack of perceived PR benefits, the government and NGO's predominant role in the area, as well as the "opt-in" nature and privacy/legal issues generally related to engaging in health-related campaigns.
+
 ## Scaling up malaria control through CSR: opportunity and risk

### 2020-04-18 paper.Rmd
updated

@@
 }
-pubmed_results <- rbind(csr_results$results %>% mutate(type = 'CSR'),
-                        fdi_results$results %>% mutate(type = 'FDI'),
-                        moz_csr_results$results %>% mutate(type = 'CSR MOZ'))
+pubmed_results <- bind_rows(csr_results$results %>% mutate(type = 'CSR | Malaria'),
+                        fdi_results$results %>% mutate(type = 'FDI | Malaria'),
+                        moz_csr_results$results %>% mutate(type = 'CSR | Mozambique'))
 
@@
   read_imf <- function(file){
-  srcs <- readxl::read_excel(file, skip = 4)
+  srcs <- readxl::read_excel(file, skip = 5)
   # Rename columns
   srcs <- srcs[,1:3]
-  srcs <- srcs[2:nrow(srcs),]
+  # srcs <- srcs[2:nrow(srcs),]
   names(srcs) <- c('from_country', 'inward', 'outward')
@@
   for (j in 2:3){
-    this_column <- as.numeric(as.character(gsub('.', '', srcs[,j], fixed = TRUE)))
+    # this_column <- as.numeric(as.character(gsub('.', '', srcs[,j], fixed = TRUE)))
+    this_column <- as.numeric(as.character(srcs[,j]))
+    this_column <- ifelse(is.na(this_column), 0, this_column)
     srcs[,j] <- this_column
@@
   mutate(from_country = factor(from_country, levels = as.character(from_country)))
-last_year <- last_year[1:10,]
+last_year <- last_year[1:20,]
 ggplot(data = last_year,
        aes(x = from_country,
-           y = inward)) +
+           y = inward )) +
   geom_bar(stat = 'identity',
@@
   theme_bw() +
-    theme(axis.text.x = element_text(angle = 90)) 
+    theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
 ```
@@
                          'A Critical Review of FDI Inflows and Economic Growth in Low-Income SADC Countries: Prospects and Challenges',
-                         'FDI MOZ',
+                         'FDI | Mozambique',
                          'Mahembe')
-econ_results[2,2:5] <- c(2000, 'Administrative barriers to foreign investment: Reducing red tape in Africa', 'FDI MOZ', 'Emery')
+econ_results[2,2:5] <- c(2000, 'Administrative barriers to foreign investment: Reducing red tape in Africa', 'FDI | Mozambique', 'Emery')
 econ_results[3,2:5] <- c(2012,
                          'Attracting Foreign Direct Investment in Developing Countries: Determinants and Policies-A Comparative Study between Mozambique and China',
-                         'FDI MOZ', 'Tembe')
+                         'FDI | Mozambique', 'Tembe')
 econ_results[4,2:5] <- c(2013,
                          'Contemporary Processes of Large-Scale Land Acquisition in Sub-Saharan Africa: Legal Deficiency or Elite Capture of the Rule of Law?',
-                         'FDI MOZ', 'German')
+                         'FDI | Mozambique', 'German')
 econ_results[5,2:5] <- c(2012,
                          'Corruption and Multinational Companies Entry Modes--Do Linguistic and Historical Ties Matter?',
-                         'FDI MOZ', 'Grande')
+                         'FDI | Mozambique', 'Grande')
 econ_results[6,2:5] <- c(2014,
                          'Determining the Nature and Extent of Spillovers: Empirical Assessment',
-                         'FDI MOZ', 'Winkler')
+                         'FDI | Mozambique', 'Winkler')
 econ_results[7,2:5] <- c(1999, 'Foreign Direct Investment in Africa: Policies Also Matter',
-                         'FDI MOZ', 'Morisset')
+                         'FDI | Mozambique', 'Morisset')
 econ_results[8,2:5] <- c(2002,
                          'Foreign Direct Investment in Africa--Some Case Studies',
-                         'FDI MOZ', 'Bassu')
+                         'FDI | Mozambique', 'Bassu')
 econ_results[9,2:5] <- c(2014,
                          'Growth, Capital Accumulation and Economic Porosity in Mozambique: Social Losses, Private Gains',
-                         'FDI MOZ', 'Castel-Branco')
+                         'FDI | Mozambique', 'Castel-Branco')
 econ_results[10,2:5] <- c(2007,
                           'Linkage Between foreign Direct Investment, Trade and Trade Policy: An Economic Analysis with Application to the Food Sector in OECD Countries and Case Studies in Ghana, Mozambique, Tunisia and Uganda',
-                          'FDI MOZ', 
+                          'FDI | Mozambique', 
                           'Wilson')
@@
                           'Mining FDI and Infrastructure Development on Africas East Coast: Examining the Recent Experience of Tanzania and Mozambique',
-                          'FDI MOZ',
+                          'FDI | Mozambique',
                           'Robbins')
@@
                           'Potential and actual FDI spillovers in global value chains: the role of foreign investor characteristics, absorptive capacity and transmission channels',
-                          'FDI MOZ',
+                          'FDI | Mozambique',
                           'Winkler')
@@
                           'Regional integration, FDI and competitiveness in southern Africa',
-                          'FDI MOZ',
+                          'FDI | Mozambique',
                           'Goldstein')
@@
                           'Sector Case Study: Mining',
-                          'FDI MOZ',
+                          'FDI | Mozambique',
                           'Barnard')
@@
                           'Strategic Privatisation: Rehabilitating the Mozambican Sugar Industry',
-                          'FDI MOZ',
+                          'FDI | Mozambique',
                           'Buur')
@@
                           'The Economics and Politics of Local Content in African Extractives: Lessons from Tanzania, Uganda and Mozambique',
-                          'FDI MOZ',
+                          'FDI | Mozambique',
                           'Hansen')
@@
                           'The Role of FDI in Economic Development',
-                          'FDI MOZ',
+                          'FDI | Mozambique',
                           'Bjorvatn')
@@
                           'Uncovering Trends in the Accumulation of Technological Capabilities and Skills in the Mozambican Manufacturing Sector',
-                          'FDI MOZ',
+                          'FDI | Mozambique',
                           'Warren-Rodriguez')
@@
                           'Public Governance, Health and Foreign Direct Investment in Sub-Saharan Africa',
-                          'FDI Malaria',
+                          'FDI | Malaria',
                           'Azemar')
@@
                           'Sector Case Study: Agribusiness',
-                          'FDI MOZ',
+                          'FDI | Mozambique',
                           'Barnard')
@@
 search_results$Year <- as.character(search_results$Year)
-search_results <- search_results %>% arrange(Source, Type)
+search_results <- search_results %>% arrange(Source, Type) %>% filter(!duplicated(Title))
 ```
@@
       tabular.environment='longtable')
+# save(search_results, file = 'sr.RData')
 ```

### 2020-04-18 paper_word.Rmd
updated

@@
 }
-pubmed_results <- rbind(csr_results$results %>% mutate(type = 'CSR'),
+pubmed_results <- bind_rows(csr_results$results %>% mutate(type = 'CSR'),
                         fdi_results$results %>% mutate(type = 'FDI'),

### 2020-04-18 search_pubmed.R

@@ -143,10 +143,10 @@ pubmed <- function(start_year = 2014,
     }
     
     # Bind together the results
-    results <- do.call('rbind', results_list)
+    results <- bind_rows(results_list)
     
     # Bind together the abstracts
-    abstracts <- do.call('rbind', abstracts_list)
+    abstracts <- bind_rows(abstracts_list)
     
     # Put into list
     return_object <- 
